### 2022-11-10 writeups/Observations.tex
add rules-based model

\documentclass{article}

\usepackage[a4paper, total={6in, 10in}]{geometry}

\begin{document}

{\bf \emph{Observations on the Datasets}}\\\\

{\bf Difficulty of Rules-Based Classification}

The initial takeaway I get from looking at the dataset is that is seems specifically tailored to be difficult for a rules-based system. Many of the non-antisemitic tweets have words typically used by antisemites to express antisemitism and the vocabulary overlaps between the antisemitic tweets and the tweets that simply mention Judaism. There are a few words that appear to only come up in the antisemitic tweets, such as mentions of George Soros, but those words appear in few tweets and even in the case of filtering out mentions of Soros, that would also affect innocent mentions of him. \\

The files in the counters folder can be used to examine the vocabulary differences between the antisemitic and non-antisemitic texts. The folder is split into four files, one for each type of antisemitism identified in () et al. Each word that appears in a tweet of that type is included in the CSV, next to the percentage of the time it appeared in the tweets of that type subtracted by the percentage of the time it appeared in the non-antisemitic tweets. An initial idea for a rules-based system may use the most comparatively common of these words and classify texts where they appear more than usual (with 'usual' being defined as the baseline identified with the word frequencies in the non-antisemitic tweets), possibly using some sort of hypothesis testing to see if the word frequencies appear to come from the same distribution. The problem with this is that, as exemplified by the appearance of these common words in non-antisemitic tweets in the dataset, the use of these words for classification is not very useful without the context they are used in, which is the primary advantage of a machine learning approach to text classification.

\end{document}

### 2022-11-10 make_counters.py

@@ -1,0 +1,79 @@
+import pandas as pd
+import nltk
+import collections
+import string
+import re
+import math
+
+to_zero = lambda x: 0 if math.isnan(x) else x #set missing value to 0
+delete_punctuation = lambda x: ''.join([char for char in x if char not in string.punctuation])
+
+#put text counter into df
+def to_df(text_counter, column1, column2):
+    keys = [key for key in text_counter]
+    values = [text_counter[key] for key in text_counter]
+    output_df = pd.DataFrame([keys, values], [column1, column2]).transpose()
+    output_df = output_df.sort_values(column2, ascending=False) #sort
+    return output_df
+
+#get a counter of the percentage of a text taken up by the words in that text.
+#if a baseline counter is provided to compare, subtract it.
+def get_counter(type, df, baseline=pd.DataFrame()):
+
+    #put the text of the rows of a certain type in a string, and get rid of stopwords
+    df = df[df['type'] == type]
+    text = ' '.join(list(df['text']))
+    text = text.lower()
+    for stopword in nltk.corpus.stopwords.words('english'):
+        text = re.sub(f' {stopword} ', ' ', text)
+    
+    #get a counter of the percentage of the text taken up by each word in the text
+    split_text = text.split()
+    text_counter = collections.Counter(split_text)
+    length = len(split_text)
+    for key in text_counter:
+        text_counter[key] = text_counter[key] * 100 / length
+    
+    #if a baseline is provided, subtract the value from each word in the baseline
+    #from the same word in the current text counter
+    if not baseline.empty:
+        for key in text_counter:
+            value = 0 if key not in baseline else baseline[key]
+            text_counter[key] = text_counter[key] - value
+    
+    text_counter = collections.OrderedDict(text_counter.most_common())
+    
+    return text_counter
+
+def main():
+
+    df = pd.read_csv('antisemitism_dataset.csv')
+    
+    #reconfigure/reorganize dataframe
+    df['type_of_antisemitism'] = df['type_of_antisemitism'].apply(to_zero) #replace missing values with zero
+    df['type_of_antisemitism'] = df['type_of_antisemitism'].astype(int)
+    df['classification'] = df['classification'].astype(int)
+    df['type'] = df['classification'] + df['type_of_antisemitism'] #type and classification are seperate columns, combine them
+    df['text'] = df['text'].astype(str)
+    df['text'] = df['text'].apply(delete_punctuation) #delete punctuation from the text
+    
[…]

### 2022-11-10 nlp_analysis.py

@@ -1,53 +1,0 @@
-import pandas as pd
-import nltk
-import collections
-import string
-import re
-import math
-
-to_zero = lambda x: 0 if math.isnan(x) else x
-delete_punctuation = lambda x: ''.join([char for char in x if char not in string.punctuation])
-
-def get_counter(type, df, baseline=None):
-    df = df[df['type'] == type]
-    text = ' '.join(list(df['text']))
-    text = text.lower()
-    for stopword in nltk.corpus.stopwords.words('english'):
-        text = re.sub(f' {stopword} ', ' ', text)
-    
-    split_text = text.split()
-    text_counter = collections.Counter(split_text)
-    length = len(split_text)
-    for key in text_counter:
-        text_counter[key] = text_counter[key] * 100 / length
-    
-    if baseline:
-        for key in text_counter:
-            value = 0 if key not in baseline else baseline[key]
-            text_counter[key] = text_counter[key] - value
-    
-    text_counter = collections.OrderedDict(text_counter.most_common())
-    
-    return text_counter
-
-def main():
-
-    df = pd.read_csv('antisemitism_dataset.csv')
-    
-    df['type_of_antisemitism'] = df['type_of_antisemitism'].apply(to_zero)
-    df['type_of_antisemitism'] = df['type_of_antisemitism'].astype(int)
-    df['classification'] = df['classification'].astype(int)
-    df['type'] = df['classification'] + df['type_of_antisemitism']
-    df['text'] = df['text'].astype(str) #convert column to strings
-    df['text'] = df['text'].apply(delete_punctuation)
-    
-    baseline = get_counter(0, df)
-    types = range(1, 5)
-    for type in types:
-        type_counter = get_counter(type, df, baseline)
-        formatted_string = '\n'.join([f'{key}: {type_counter[key]:.3f}' for key in type_counter])
-        with open(f'counters\\{type}.txt', 'w', encoding='utf8') as f:
-            f.write(formatted_string)
-
-if __name__ == '__main__':
-    main()

### 2022-11-10 rulesbased.py

@@ -1,0 +1,102 @@
+import pandas as pd
+import nltk
+import math
+import re
+import collections
+
+vector_distance = lambda v1, v2: math.sqrt(sum([(v1[i]-v2[i])**2 for i, _ in enumerate(v1)])) #distance between two vectors equation
+
+#put text counter into df
+def to_df(text_counter, column1, column2):
+    keys = [key for key in text_counter]
+    values = [text_counter[key] for key in text_counter]
+    output_df = pd.DataFrame([keys, values], [column1, column2]).transpose()
+    return output_df
+    
+def get_text_counter(text):
+    text = text.lower()
+    for stopword in nltk.corpus.stopwords.words('english'):
+        text = re.sub(f' {stopword} ', ' ', text)
+    
+    #get a counter of the percentage of the text taken up by each word in the text
+    split_text = text.split()
+    text_counter = collections.Counter(split_text)
+    length = len(split_text)
+    for key in text_counter:
+        text_counter[key] = text_counter[key] * 100 / length
+    
+    text_counter = collections.OrderedDict(text_counter.most_common())
+    
+    return text_counter
+
+#get a counter of the percentage of a text taken up by the words in that text.
+def get_counter(type, df):
+
+    #put the text of the rows of a certain type in a string, and get rid of stopwords
+    df = df[df['type'] == type]
+    text = ' '.join(list(df['text']))
+    text = text.lower()
+    for stopword in nltk.corpus.stopwords.words('english'):
+        text = re.sub(f' {stopword} ', ' ', text)
+    
+    #get a counter of the percentage of the text taken up by each word in the text
+    split_text = text.split()
+    text_counter = collections.Counter(split_text)
+    length = len(split_text)
+    for key in text_counter:
+        text_counter[key] = text_counter[key] * 100 / length
+    
+    text_counter = collections.OrderedDict(text_counter.most_common())
+    
+    return text_counter
+
+#input frequency df for tweet, list of comparison dfs for dataset,
+#and names of column1 (words) and column2 (frequencies),
+#output the vector of the text and list of vectors for each comparison df
+def to_vector(text_df, comparison_dict, column1, column2):
+    comparison_vector_dict = dict()
+    text_df = text_df.sort_values(column1) #sort alphabetically (so they're in the same order)
+    text_vector = list(text_df[column2])
+    words = list(text_df[column1]) #words in tweet
+
+    for key in comparison_dict:
+        comparison_df = comparison_dict[key]
[…]

### 2022-11-11 make_counters.py
add high-level functionality to rules-based classifier

@@ -18,11 +18,11 @@ def to_df(text_counter, column1, column2):
 
 #get a counter of the percentage of a text taken up by the words in that text.
 #if a baseline counter is provided to compare, subtract it.
-def get_counter(type, df, baseline=pd.DataFrame()):
+def get_counter(type, df, typename='type', baseline=pd.DataFrame()):
 
     #put the text of the rows of a certain type in a string, and get rid of stopwords
-    df = df[df['type'] == type]
-    text = ' '.join(list(df['text']))
+    df = df[df[typename] == type]
+    text = ' '.join(list(df['text'].astype(str)))
     text = text.lower()
     for stopword in nltk.corpus.stopwords.words('english'):
         text = re.sub(f' {stopword} ', ' ', text)
@@ -50,18 +50,25 @@ def main():
     df = pd.read_csv('antisemitism_dataset.csv')
     
     #reconfigure/reorganize dataframe
-    df['type_of_antisemitism'] = df['type_of_antisemitism'].apply(to_zero) #replace missing values with zero
-    df['type_of_antisemitism'] = df['type_of_antisemitism'].astype(int)
     df['classification'] = df['classification'].astype(int)
-    df['type'] = df['classification'] + df['type_of_antisemitism'] #type and classification are seperate columns, combine them
     df['text'] = df['text'].astype(str)
     df['text'] = df['text'].apply(delete_punctuation) #delete punctuation from the text
     
-    #get counters and write to file
-    baseline = get_counter(0, df) #get a baseline percentage counter (this is the counter for the non-antisemitic text)
+    #get counters and write to file for non-antisemitic text
+    baseline = get_counter(0, df, typename='classification') #get a baseline percentage counter (this is the counter for the non-antisemitic text)
     base_df = to_df(baseline, 'Words', 'Percent Appearance')
     base_df.to_csv('frequencies\\baseline.csv', index=False)
     
+    #get counters and write to file for antisemitic text
+    all_antisemitic = get_counter(1, df, typename='classification') #get a baseline percentage counter (this is the counter for the non-antisemitic text)
+    all_df = to_df(all_antisemitic, 'Words', 'Percent Appearance')
+    all_df.to_csv('frequencies\\antisemitic.csv', index=False)
+    
+    #combine type of antisemitism and classification columns
+    df['type_of_antisemitism'] = df['type_of_antisemitism'].apply(to_zero) #replace missing values with zero
+    df['type_of_antisemitism'] = df['type_of_antisemitism'].astype(int)
[…]

### 2022-11-11 rulesbased.py

@@ -4,6 +4,9 @@ import math
 import re
 import collections
 
+WORD_COLUMN = 'Words'
+PERCENT_COLUMN = 'Percent Appearance'
+
 vector_distance = lambda v1, v2: math.sqrt(sum([(v1[i]-v2[i])**2 for i, _ in enumerate(v1)])) #distance between two vectors equation
 
 #put text counter into df
@@ -90,13 +93,25 @@ def compare_vectors(text_vector, comparison_vector_dict):
     return closest
 
 #put it all together
-def compare_text(text, df, column1, column2):
-    keys = ['baseline', '1', '2', '3', '4']
+def compare_text(text, keys, df_dict=None, column1=WORD_COLUMN, column2=PERCENT_COLUMN):
     text_counter = get_text_counter(text)
     text_df = to_df(text_counter, column1, column2)
-    df_dict = dict()
-    for key in keys:
-        df_dict[key] = pd.read_csv(f'frequencies\\{key}.csv')
+    
+    #make dictionary with given keys
+    if not df_dict:
+        df_dict = dict()
+        for key in keys:
+            df_dict[key] = pd.read_csv(f'frequencies\\{key}.csv')
+    
     text_vector, comparison_vector_dict = to_vector(text_df, df_dict, column1, column2)
     closest = compare_vectors(text_vector, comparison_vector_dict)
     return closest
+
+#returns True if classified as antisemitic, False otherwise
+def is_antisemitic(text, df_dict=None):
+    return 'antisemitic' == compare_text(text, ['antisemitic', 'baseline'], df_dict=df_dict)
+
+#returns type of antisemitism predicted
+def type(text, df_dict=None):
+    types = {'1': 'political', '2': 'economic', '3': 'religious', '4': 'racial'}
+    return types[compare_text(text, ['1', '2', '3', '4'], df_dict=df_dict)]

### 2022-11-11 reorg_scripts/consolidate.py
reorganize into directories

@@ -1,23 +1,25 @@
 import re
 import pandas as pd
 
+BASIC = '..\\data'
+
 select = lambda value_list, selection_list, minimum_value: [value for (i, value) in enumerate(value_list) if selection_list[i] > minimum_value]
 
 def main():
-    with open('raw.txt', 'r', encoding='utf8') as f:
+    with open(f'{BASIC}\\raw.txt', 'r', encoding='utf8') as f:
         text = f.read()
         text = re.sub('\n[A-Z]\n', '', text) #delete lines with only single uppercase letter (these lines denote the beginning of a new letter, not useful)
         words = re.findall('(?<=\n).+?(?= \– )', text) #get words, in format (\nWORD - ).
         definitions = re.findall('(?<= \– ).+?(?=\n)', text) #get definitions, in format ( - DEFINITION\n).
     
-    with open('4chanwords.txt', 'r', encoding='utf8') as f:
+    with open(f'{BASIC}\\4chanwords.txt', 'r', encoding='utf8') as f:
         text = f.read()
         text = re.sub('\n.(?=\n)', '', text) #delete single-character lines
         temp_words = re.findall('(?<=\n).+?(?=\n)', text) #get words
         words += temp_words #add to total words
         definitions += ['None'] * len(temp_words) #add 'None' definitions to words
     
-    df = pd.read_csv('cleaned_data.csv') #read data
+    df = pd.read_csv(f'{BASIC}\\cleaned_data.csv') #read data
     
     appearances_list = [] #list of appearances of each word
     averaged_values = {'hateful': [], 'offensive': [],
@@ -44,7 +46,7 @@ def main():
     #sort descending by # of appearances
     word_df = word_df.sort_values('appearances', ascending=False)
     
-    word_df.to_csv('words.csv', index=False)
+    word_df.to_csv(f'{BASIC}\\words.csv', index=False)
 
 if __name__ == '__main__':
     main()

### 2022-11-11 reorg_scripts/make_counters.py

@@ -57,12 +57,12 @@ def main():
     #get counters and write to file for non-antisemitic text
     baseline = get_counter(0, df, typename='classification') #get a baseline percentage counter (this is the counter for the non-antisemitic text)
     base_df = to_df(baseline, 'Words', 'Percent Appearance')
-    base_df.to_csv('frequencies\\baseline.csv', index=False)
+    base_df.to_csv('..\\rulesbased_model\\frequencies\\baseline.csv', index=False)
     
     #get counters and write to file for antisemitic text
     all_antisemitic = get_counter(1, df, typename='classification') #get a baseline percentage counter (this is the counter for the non-antisemitic text)
     all_df = to_df(all_antisemitic, 'Words', 'Percent Appearance')
-    all_df.to_csv('frequencies\\antisemitic.csv', index=False)
+    all_df.to_csv('..\\rulesbased_model\\frequencies\\antisemitic.csv', index=False)
     
     #combine type of antisemitism and classification columns
     df['type_of_antisemitism'] = df['type_of_antisemitism'].apply(to_zero) #replace missing values with zero
@@ -75,12 +75,12 @@ def main():
         #non-antisemitic text
         text_counter = get_counter(type, df, base_df) #make compared counter
         temp_df = to_df(text_counter, 'Words', 'Percent Increase')
-        temp_df.to_csv(f'counters\\{type}.csv', index=False)
+        temp_df.to_csv(f'..\\data\\counters\\{type}.csv', index=False)
         
         #make raw counters to be used for the rules-based model
         frequency_counter = get_counter(type, df) #make frequency counter
         temp_df = to_df(frequency_counter, 'Words', 'Percent Appearance')
-        temp_df.to_csv(f'frequencies\\{type}.csv', index=False)
+        temp_df.to_csv(f'..\\rulesbased_model\\frequencies\\{type}.csv', index=False)
 
 if __name__ == '__main__':
     main()

### 2022-11-12 reorg_scripts/gather_tweets.py
rework baseline frequencies

@@ -1,0 +1,45 @@
+import snscrape.modules.twitter as sntwitter
+import pandas as pd
+
+LIMIT = 1000
+
+#transpose a matrix
+def transpose(matrix):
+    num_rows = len(matrix[0])
+    num_columns = len(matrix)
+    Final = [[0]*num_columns for x in range(num_rows)]
+    for x in range(num_rows):
+        for i in range(num_columns):
+            Final[x][i] = matrix[i][x]
+    return Final
+    
+#change an int to a string. if it's only made up of one character,
+#add a zero to the beginning. (this is because of the way twitter 
+#handles dates)
+def add_zero(date):
+    date = str(date)
+    if len(date) == 1:
+        date = '0' + date
+    return date
+
+def main():
+    
+    start_dates = [x+1 for x in range(11)]
+    
+    tweets = []
+    for date_index, start_date in enumerate(start_dates):
+        end_date = start_date + 1
+        start_date = add_zero(start_date) #convert to str and add 0 to start if one digit
+        end_date = add_zero(end_date)
+        query = f" lang:en until:2021-{end_date}-01 since:2021-{start_date}-01" #the query submitted to twitter
+        for index, tweet in enumerate(sntwitter.TwitterSearchScraper(query).get_items()):
+            tweets.append([date_index+1, tweet.content]) #appends each tweet we get to the list with the month it's from
+            if index == LIMIT-1: #stop once the limit has been reached
+                break
+    tweets = transpose(tweets)
+    tweet_df = pd.DataFrame(tweets, ['Month', 'text'])
+    tweet_df = tweet_df.transpose()
+    tweet_df.to_csv('..\\data\\normal_tweets.csv', index=False)
+
+if __name__ == '__main__':
+    main() 

### 2022-11-12 reorg_scripts/make_counters.py

@@ -5,6 +5,9 @@ import string
 import re
 import math
 
+COLUMN1 = 'Words'
+COLUMN2 = 'Percent Appearance'
+
 to_zero = lambda x: 0 if math.isnan(x) else x #set missing value to 0
 delete_punctuation = lambda x: ''.join([char for char in x if char not in string.punctuation])
 
@@ -15,6 +18,23 @@ def to_df(text_counter, column1, column2):
     output_df = pd.DataFrame([keys, values], [column1, column2]).transpose()
     output_df = output_df.sort_values(column2, ascending=False) #sort
     return output_df
+
+#get counter from only text
+def get_text_counter(text):
+    text = text.lower()
+    for stopword in nltk.corpus.stopwords.words('english'):
+        text = re.sub(f' {stopword} ', ' ', text)
+    
+    #get a counter of the percentage of the text taken up by each word in the text
+    split_text = text.split()
+    text_counter = collections.Counter(split_text)
+    length = len(split_text)
+    for key in text_counter:
+        text_counter[key] = text_counter[key] * 100 / length
+    
+    text_counter = collections.OrderedDict(text_counter.most_common())
+    
+    return text_counter
 
 #get a counter of the percentage of a text taken up by the words in that text.
 #if a baseline counter is provided to compare, subtract it.
@@ -47,22 +67,24 @@ def get_counter(type, df, typename='type', baseline=pd.DataFrame()):
 
 def main():
 
-    df = pd.read_csv('antisemitism_dataset.csv')
+    df = pd.read_csv('..\\data\\antisemitism_dataset.csv')
     
     #reconfigure/reorganize dataframe
     df['classification'] = df['classification'].astype(int)
     df['text'] = df['text'].astype(str)
     df['text'] = df['text'].apply(delete_punctuation) #delete punctuation from the text
     
-    #get counters and write to file for non-antisemitic text
-    baseline = get_counter(0, df, typename='classification') #get a baseline percentage counter (this is the counter for the non-antisemitic text)
-    base_df = to_df(baseline, 'Words', 'Percent Appearance')
-    base_df.to_csv('..\\rulesbased_model\\frequencies\\baseline.csv', index=False)
-    
     #get counters and write to file for antisemitic text
     all_antisemitic = get_counter(1, df, typename='classification') #get a baseline percentage counter (this is the counter for the non-antisemitic text)
[…]

### 2022-11-12 rulesbased_model/rulesbased.py

@@ -109,7 +109,7 @@ def compare_text(text, keys, df_dict=None, column1=WORD_COLUMN, column2=PERCENT_
 
 #returns True if classified as antisemitic, False otherwise
 def is_antisemitic(text, df_dict=None):
-    return 'antisemitic' == compare_text(text, ['antisemitic', 'baseline'], df_dict=df_dict)
+    return 'antisemitic' == compare_text(text, ['baseline', 'antisemitic'], df_dict=df_dict)
 
 #returns type of antisemitism predicted
 def type(text, df_dict=None):

### 2022-11-13 writeups/Observations.tex
add tex explanation doc and test

@@
 \usepackage[a4paper, total={6in, 10in}]{geometry}
+\usepackage{amssymb}
 
@@
 
-The initial takeaway I get from looking at the dataset is that is seems specifically tailored to be difficult for a rules-based system.
+The initial takeaway I get from looking at the dataset is that is seems specifically tailored to be difficult for a rules-based system to classify.
 Many of the non-antisemitic tweets have words typically used by antisemites to express antisemitism and the vocabulary overlaps between the antisemitic tweets and the tweets that simply mention Judaism.
@@
 
-The files in the counters folder can be used to examine the vocabulary differences between the antisemitic and non-antisemitic texts.
-The folder is split into four files, one for each type of antisemitism identified in () et al.
-Each word that appears in a tweet of that type is included in the CSV, next to the percentage of the time it appeared in the tweets of that type subtracted by the percentage of the time it appeared in the non-antisemitic tweets.
-An initial idea for a rules-based system may use the most comparatively common of these words and classify texts where they appear more than usual (with 'usual' being defined as the baseline identified with the word frequencies in the non-antisemitic tweets), possibly using some sort of hypothesis testing to see if the word frequencies appear to come from the same distribution.
-The problem with this is that, as exemplified by the appearance of these common words in non-antisemitic tweets in the dataset, the use of these words for classification is not very useful without the context they are used in, which is the primary advantage of a machine learning approach to text classification.
+In the directory labeled ``data'', there is a subdirectory labeled ``counters'' that have files that coule be useful for comparing the frequency of different words in tweets that were identified as being different types of antisemitic.
+The file's title gives the type of antisemitism (a number 1 through 4, corresponding to political, economic, religious, or racial antisemitism), and in the file is a list of words next to a number.
+This number is the frequency of that word in the dataset of that type of antisemitism subtracted by the frequency of the same word in the non-antisemitic tweets, to give a view of which words were more or less common in these datasets.
+Looking at the words, it seems as if none of those that are significantly more or less common appear to be necessarily antisemitic words, so just using the appearance of these words and nothing else would likely not be very useful for classification.
+Instead, I decided to take a more holistic approach to the rules-based classifier.
+\\
+
+{\bf Rules-Based Classifier Design}
+
+I begin by taking in an input text, putting it all in lowercase, removing stopwords, removing punctuation, and reducing it to a percentage frequency counter of each unique word in the text.
+For example, \texttt{"The big, big brown fox"} would be represented as \texttt{\{"big": 50, "brown": 25, "fox": 25\}}.
+This process is done for the combined text of all antisemitic tweets and non-antisemitic tweets, as well as individually for the tweets of each of the four types of antisemitism.
+I stored the resultant frequencies in a subdirectory of the "rulesbased\_model" directory entitled "frequencies." In order to determine if a text is antisemitic, I do this same basic process to represent it as a series of frequencies.
+I then read the frequencies of the antisemitic and non-antisemitic datasets, removing any words from those datasets that do not appear in the input text and adding a new row with a frequency of 0 for words that appear in the input text but not the dataset.
+After that, I order the words alphabetically in all three \texttt{DataFrame} objects so they are all in the same order.
+Then, I stored the second column of the \texttt{DataFrame} objects (the one containing the frequency percentages) in three vectors, one for the input, and the other two for the two datasets.
+I then got the distance between the input vector and each of the two dataset vectors, and chose whichever dataset had the smaller distance as the one that more closely represented the input data.\\
+
+For example, for the counter \texttt{\{"big": 50, "brown": 25, "fox": 25\}}, the antisemitic dataset would have a counter \texttt{\{"big": 0.056159, "brown": 0, "fox": 0\}}, and the baseline would have \texttt{\{"big": 0.125219, "brown": 0, "fox": 0.008348\}}.
+The next step would be to sort these words alphabetically, but luckily the order they come in in the text is already alphabetical.
+Next, they would be represented as vectors and the distance would be found between them and the input text, as shown:\\
+\[
+\left[ {\begin{array}{c}
+    50\\
+    25\\
+    25\\
+  \end{array} } \right] - \left[ {\begin{array}{c}
+    0.056159\\
+    0\\
+    0\\
+  \end{array} } \right] = \left[ {\begin{array}{c}
+    49.943841\\
+    25\\
+    25\
+  \end{array} } \right] , \left| {\begin{array}{c}
+    49.943841\\
+    25\\
+    25\
+  \end{array} } \right| \approx 61.19
+\]
+
+\[
+\left[ {\begin{array}{c}
+    50\\
+    25\\
+    25\\
+  \end{array} } \right] - \left[ {\begin{array}{c}
+    0.125219\\
+    0\\
+    0.008348\\
+  \end{array} } \right] = \left[ {\begin{array}{c}
+    49.874781\\
+    25\\
+    24.991652\\
+  \end{array} } \right] , \left| {\begin{array}{c}
+    49.874781\\
+    25\\
+    24.991652\\
+  \end{array} } \right| \approx 61.13
+\]
+
+Since the vector representing the non-antisemitic dataset is slightly closer to this input text, it would be classified as not antisemitic.
+The distances between these vectors becomes smaller as larger texts are inputted with frequency distributions that more closely match those of the datasets.
 

### 2022-11-13 reorg_scripts/make_counters.py

@@ -19,23 +19,6 @@ def to_df(text_counter, column1, column2):
     output_df = output_df.sort_values(column2, ascending=False) #sort
     return output_df
 
-#get counter from only text
-def get_text_counter(text):
-    text = text.lower()
-    for stopword in nltk.corpus.stopwords.words('english'):
-        text = re.sub(f' {stopword} ', ' ', text)
-    
-    #get a counter of the percentage of the text taken up by each word in the text
-    split_text = text.split()
-    text_counter = collections.Counter(split_text)
-    length = len(split_text)
-    for key in text_counter:
-        text_counter[key] = text_counter[key] * 100 / length
-    
-    text_counter = collections.OrderedDict(text_counter.most_common())
-    
-    return text_counter
-
 #get a counter of the percentage of a text taken up by the words in that text.
 #if a baseline counter is provided to compare, subtract it.
 def get_counter(type, df, typename='type', baseline=pd.DataFrame()):
@@ -43,9 +26,10 @@ def get_counter(type, df, typename='type', baseline=pd.DataFrame()):
     #put the text of the rows of a certain type in a string, and get rid of stopwords
     df = df[df[typename] == type]
     text = ' '.join(list(df['text'].astype(str)))
-    text = text.lower()
+    text = f' {text} '
+    text = delete_punctuation(text.lower())
     for stopword in nltk.corpus.stopwords.words('english'):
-        text = re.sub(f' {stopword} ', ' ', text)
+        text = text.replace(f' {stopword} ', ' ')
     
     #get a counter of the percentage of the text taken up by each word in the text
     split_text = text.split()
@@ -80,9 +64,7 @@ def main():
     all_df.to_csv('..\\rulesbased_model\\frequencies\\antisemitic.csv', index=False)
     
     #get counters and write to file for non-antisemitic text
-    non_antisemitic = pd.read_csv('..\\data\\normal_tweets.csv')
-    non_antisemitic_text = ' '.join(list(non_antisemitic['text']))
-    non_antisemitic_counter = get_text_counter(non_antisemitic_text)
+    non_antisemitic_counter = get_counter(0, df, typename='classification')
     base_df = to_df(non_antisemitic_counter, COLUMN1, COLUMN2)
     base_df.to_csv('..\\rulesbased_model\\frequencies\\baseline.csv', index=False)
     

### 2022-11-13 rulesbased_model/dataset.py

@@ -1,0 +1,28 @@
+import pandas as pd
+import rulesbased as rb
+
+DATA_DIR = '..\\data'
+FREQ_DIR = 'frequencies'
+
+def main():
+    types = {'political': 0,
+             'economic': 1,
+             'religious': 2,
+             'racial': 3}
+
+    is_antisemitic_dict = {'baseline': pd.read_csv(f'{FREQ_DIR}\\baseline.csv'),
+                           'antisemitic': pd.read_csv(f'{FREQ_DIR}\\antisemitic.csv')}
+    
+    type_dict = dict()
+    for x in range(1, 4):
+        type_dict[str(x)] = pd.read_csv(f'{FREQ_DIR}\\{x}.csv')
+    
+    files = ['normal_tweets', 'cleaned_data', 'antisemitism_dataset']
+    for filename in files:
+        df = pd.read_csv(f'{DATA_DIR}\\{filename}.csv')
+        df['is_antisemitic'] = df['text'].apply(lambda x: int(rb.is_antisemitic(x, df_dict=is_antisemitic_dict)))
+        df['predicted_type'] = df['text'].apply(lambda x: types[rb.type(x, df_dict=type_dict)])
+        df.to_csv(f'{DATA_DIR}\\predicted\\{filename}-predicted.csv', index=False)
+
+if __name__ == '__main__':
+    main()

### 2022-11-13 rulesbased_model/rulesbased.py

@@ -3,11 +3,16 @@ import nltk
 import math
 import re
 import collections
+import string
 
 WORD_COLUMN = 'Words'
 PERCENT_COLUMN = 'Percent Appearance'
 
-vector_distance = lambda v1, v2: math.sqrt(sum([(v1[i]-v2[i])**2 for i, _ in enumerate(v1)])) #distance between two vectors equation
+vector_distance = lambda v1, v2: math.sqrt(sum([(v1[i]-v2[i])**2 for (i, _) in enumerate(v1)])) #distance between two vectors equation
+#dot = lambda v1, v2: sum([v1[i]*v2[i] for (i, _) in enumerate(v1)])
+#mag = lambda v: math.sqrt(sum([item**2 for item in v]))
+#cosine_similarity = lambda v1, v2: dot(v1, v2)/mag(v1)*mag(v2)
+delete_punctuation = lambda x: ''.join([char for char in x if char not in string.punctuation])
 
 #put text counter into df
 def to_df(text_counter, column1, column2):
@@ -17,9 +22,10 @@ def to_df(text_counter, column1, column2):
     return output_df
     
 def get_text_counter(text):
-    text = text.lower()
+    text = f' {text} '
+    text = delete_punctuation(text.lower())
     for stopword in nltk.corpus.stopwords.words('english'):
-        text = re.sub(f' {stopword} ', ' ', text)
+        text = text.replace(f' {stopword} ', ' ')
     
     #get a counter of the percentage of the text taken up by each word in the text
     split_text = text.split()
@@ -38,9 +44,10 @@ def get_counter(type, df):
     #put the text of the rows of a certain type in a string, and get rid of stopwords
     df = df[df['type'] == type]
     text = ' '.join(list(df['text']))
-    text = text.lower()
+    text = f' {text} '
+    text = delete_punctuation(text.lower())
     for stopword in nltk.corpus.stopwords.words('english'):
-        text = re.sub(f' {stopword} ', ' ', text)
+        text = text.replace(f' {stopword} ', ' ')
     
     #get a counter of the percentage of the text taken up by each word in the text
     split_text = text.split()

### 2022-11-14 rulesbased_model/dataset.py
evaluate model

@@ -10,18 +10,23 @@ def main():
              'religious': 2,
              'racial': 3}
 
+    #get dataframes for bool is_antisemitic
     is_antisemitic_dict = {'baseline': pd.read_csv(f'{FREQ_DIR}\\baseline.csv'),
                            'antisemitic': pd.read_csv(f'{FREQ_DIR}\\antisemitic.csv')}
     
+    #get dataframes for types
     type_dict = dict()
     for x in range(1, 4):
         type_dict[str(x)] = pd.read_csv(f'{FREQ_DIR}\\{x}.csv')
     
+    #apply to files
     files = ['normal_tweets', 'cleaned_data', 'antisemitism_dataset']
     for filename in files:
         df = pd.read_csv(f'{DATA_DIR}\\{filename}.csv')
         df['is_antisemitic'] = df['text'].apply(lambda x: int(rb.is_antisemitic(x, df_dict=is_antisemitic_dict)))
         df['predicted_type'] = df['text'].apply(lambda x: types[rb.type(x, df_dict=type_dict)])
+        df['cos_is_antisemitic'] = df['text'].apply(lambda x: int(rb.cosine_is_antisemitic(x, df_dict=is_antisemitic_dict)))
+        df['cos_predicted_type'] = df['text'].apply(lambda x: types[rb.cosine_type(x, df_dict=type_dict)])
         df.to_csv(f'{DATA_DIR}\\predicted\\{filename}-predicted.csv', index=False)
 
 if __name__ == '__main__':

### 2022-11-14 rulesbased_model/rulesbased.py

@@ -9,9 +9,9 @@ WORD_COLUMN = 'Words'
 PERCENT_COLUMN = 'Percent Appearance'
 
 vector_distance = lambda v1, v2: math.sqrt(sum([(v1[i]-v2[i])**2 for (i, _) in enumerate(v1)])) #distance between two vectors equation
-#dot = lambda v1, v2: sum([v1[i]*v2[i] for (i, _) in enumerate(v1)])
-#mag = lambda v: math.sqrt(sum([item**2 for item in v]))
-#cosine_similarity = lambda v1, v2: dot(v1, v2)/mag(v1)*mag(v2)
+dot = lambda v1, v2: sum([v1[i]*v2[i] for (i, _) in enumerate(v1)])
+mag = lambda v: math.sqrt(sum([item**2 for item in v]))
+cosine_similarity = lambda v1, v2: dot(v1, v2)/mag(v1)*mag(v2)
 delete_punctuation = lambda x: ''.join([char for char in x if char not in string.punctuation])
 
 #put text counter into df
@@ -99,6 +99,16 @@ def compare_vectors(text_vector, comparison_vector_dict):
     closest = min(distances)[1]
     return closest
 
+#compare vectors using cosine similarity
+def cosine_compare_vectors(text_vector, comparison_vector_dict):
+    similarities = [] #list of tuples in format (distance, key)
+    for key in comparison_vector_dict:
+        compare_vector = comparison_vector_dict[key]
+        similarity = cosine_similarity(text_vector, compare_vector)
+        similarities.append((similarity, key))
+    closest = max(similarities)[1]
+    return closest
+
 #put it all together
 def compare_text(text, keys, df_dict=None, column1=WORD_COLUMN, column2=PERCENT_COLUMN):
     text_counter = get_text_counter(text)
@@ -113,12 +123,36 @@ def compare_text(text, keys, df_dict=None, column1=WORD_COLUMN, column2=PERCENT_
     text_vector, comparison_vector_dict = to_vector(text_df, df_dict, column1, column2)
     closest = compare_vectors(text_vector, comparison_vector_dict)
     return closest
+    
+def cosine_compare_text(text, keys, df_dict=None, column1=WORD_COLUMN, column2=PERCENT_COLUMN):
+    text_counter = get_text_counter(text)
+    text_df = to_df(text_counter, column1, column2)
+    
+    #make dictionary with given keys
+    if not df_dict:
+        df_dict = dict()
+        for key in keys:
[…]

### 2022-11-14 writeups/Observations.tex
evaluate model
fix model avoidance of class 4

@@
 
-{\bf \emph{Observations on the Datasets}}\\\\
+{\bf \emph{Observations on the Datasets}} \hfill Nicolas Cloutier\\\\
 
@@
 The file's title gives the type of antisemitism (a number 1 through 4, corresponding to political, economic, religious, or racial antisemitism), and in the file is a list of words next to a number.
-This number is the frequency of that word in the dataset of that type of antisemitism subtracted by the frequency of the same word in the non-antisemitic tweets, to give a view of which words were more or less common in these datasets.
-Looking at the words, it seems as if none of those that are significantly more or less common appear to be necessarily antisemitic words, so just using the appearance of these words and nothing else would likely not be very useful for classification.
+This number is the percent frequency of that word in the dataset of that type of antisemitism subtracted by the percent frequency of the same word in the non-antisemitic tweets, to give a view of which words were more or less common in these datasets.
+Looking at the words, it seems as if none of those that are significantly more or less common appear to be necessarily antisemitic words.
+As a result, only using the appearance of these words would likely not be very useful for classification.
 Instead, I decided to take a more holistic approach to the rules-based classifier.
@@
 This process is done for the combined text of all antisemitic tweets and non-antisemitic tweets, as well as individually for the tweets of each of the four types of antisemitism.
-I stored the resultant frequencies in a subdirectory of the "rulesbased\_model" directory entitled "frequencies." In order to determine if a text is antisemitic, I do this same basic process to represent it as a series of frequencies.
+I stored the resultant frequencies in a subdirectory of the ``rulesbased\texttt{\char`_}model'' directory entitled ``frequencies.'' In order to determine if a text is antisemitic, I do this same basic process to represent it as a series of frequencies.
 I then read the frequencies of the antisemitic and non-antisemitic datasets, removing any words from those datasets that do not appear in the input text and adding a new row with a frequency of 0 for words that appear in the input text but not the dataset.
 After that, I order the words alphabetically in all three \texttt{DataFrame} objects so they are all in the same order.
-Then, I stored the second column of the \texttt{DataFrame} objects (the one containing the frequency percentages) in three vectors, one for the input, and the other two for the two datasets.
-I then got the distance between the input vector and each of the two dataset vectors, and chose whichever dataset had the smaller distance as the one that more closely represented the input data.\\
+Then, I store the second column of the \texttt{DataFrame} objects (the one containing the frequency percentages) in three vectors, one for the input, and the other two for the two datasets.
+I then find the distance between the input vector and each of the two dataset vectors, and choose whichever dataset has the smaller distance as the one that more closely represents the input data.\\
 
@@
     25\
-  \end{array} } \right] , \left| {\begin{array}{c}
-    49.943841\\
-    25\\
-    25\
-  \end{array} } \right| \approx 61.19
+  \end{array} } \right] = \vec{a} , \left| \vec{a} \right| \approx 61.19
 \]
@@
     24.991652\\
-  \end{array} } \right] , \left| {\begin{array}{c}
-    49.874781\\
-    25\\
-    24.991652\\
-  \end{array} } \right| \approx 61.13
+  \end{array} } \right] = \vec{b},\hfill \left| \vec{b} \right| \approx 61.13
 \]
@@
 The distances between these vectors becomes smaller as larger texts are inputted with frequency distributions that more closely match those of the datasets.
+An alternative method would be finding the cosine similarity between the vectors and choosing whichever has the highest as the closest distribution.\\
 
+{\bf Model Evaluation}
+
+Since the model used distributions from the antisemitic tweet dataset, using these tweets for evaluation is not very useful.
+That being said, we can use it as a preliminary tool to make sure the model is identifying differences in word distributions.
+Testing the model on the all of the dataset tweets for the Boolean antisemitic classifications and the type classifications evaluated on only the antisemitic tweets for which classifications were provided, the following accuracy ratings (out of 1) were achieved.
+\begin{center}
+Table 1: Evaluation on training data\\
+\begin{tabular}{ |c|c|c|c| } 
+ \hline
+{\bf Method} & \bf{Task} & \bf{Accuracy} & \bf{Chance Accuracy}\\
+\hline
+ Vector distance & Boolean classification & 0.76 & 0.5\\ 
+ Vector distance & Type classification & 0.57 & 0.25\\ 
+ Cosine similarity & Boolean classification & 0.76 & 0.5\\ 
+ Cosine similarity & Type classification & 0.57 & 0.25\\ 
+ \hline
+\end{tabular}
+\end{center}
+
+It would appear as if the method used has little effect on the accuracy, and that the models, scoring significantly above chance, have at the very least identified distribution differences in the tweets.
+The model can be used for binary classification on the 4Chan posts and normal tweets (a series of several thousand tweets randomly gathered from Twitter tweeted over the course of 2021) to check that they correctly identify that the 4Chan posts are mostly antisemitic and the normal tweets are mostly not antisemitic.
+Results of the model on those datasets are shown below.
+In Table 2, the ``Bool.
+avg.'' column is the arithmetic mean of all of the Boolean classifications of each dataset, the ``Cos.
+Bool.
+avg.'' column is the same but using cosine similarity, and the columns ending in ``\#'' give the raw number of occurences of each type classification for each post in the dataset (using vector distance).
+\begin{center}
+Table 2: Data summaries\\
+\begin{tabular}{ |c|c|c|c|c|c|c|}
+\hline {\bf Dataset} & {\bf Bool. avg.} & {\bf Cos. Bool. avg.} & {\bf 1 \#} & {\bf 2 \#} & {\bf 3 \#} & {\bf 4 \#}\\
+\hline 
+4Chan data & 0.85 & 0.85 & 504 & 575 & 1474 & 793\\
+Normal tweets & 0.60 & 0.60 & 3675 & 1357 & 2419 & 3549\\
+Labeled data & 0.44 & 0.44 & 515 & 133 & 277 & 296\\
+\hline
+\end{tabular}
+\end{center}
+
+Again, the different methods for calculating similarity appear to not affect outcomes at all.
+The occurence columns are given with vector distance, but these data are exactly the same as those calculated using cosine similarity.
+Looking at the Boolean averages for the normal data, it appears to be higher than may have been anticipated or hoped for, given that the tweets were collected from a random sample of tweets across 2021.
+Interestingly, the Boolean average of the normal tweets is higher than the Boolean average of the labeled data, which contained antisemitic tweets.
 \end{document}

### 2022-11-14 rulesbased_model/dataset.py

@@ -16,7 +16,7 @@ def main():
     
     #get dataframes for types
     type_dict = dict()
-    for x in range(1, 4):
+    for x in range(1, 5):
         type_dict[str(x)] = pd.read_csv(f'{FREQ_DIR}\\{x}.csv')
     
     #apply to files

### 2022-11-25 ml_models/ml_models.py
fit ml models

@@ -1,0 +1,121 @@
+import pandas as pd
+from sklearn import tree, svm, naive_bayes
+import string
+import nltk
+import collections
+import math
+import pickle
+
+BASIC = '..\\data'
+MODEL_DIR = '.\\models'
+
+value_or_zero = lambda x, dict: 0 if not x in dict else dict[x] #return dictionary value if key in dictionary, otherwise return 0
+delete_punctuation = lambda x: ''.join([char for char in x if char not in string.punctuation]) #delete punctuation from string
+
+#remove stopwords from text, put in lowercase, and delete punctuation
+def clean(text):
+    text = f' {text} '
+    text = delete_punctuation(text.lower())
+    for stopword in nltk.corpus.stopwords.words('english'):
+        text = text.replace(f' {stopword} ', ' ')
+    return text
+
+#test single-variable output models,
+#output of 1 means every prediction correct,
+#0 means every prediction incorrect
+def test(model, test_input, test_output):
+    predictions = model.predict(test_input)
+    wrong = [int(prediction == test_output[i]) for (i, prediction) in enumerate(predictions)]
+    avg = sum(wrong) / len(wrong)
+    return avg
+    
+#fit a model, save to file, and test
+def fit_save_test(model, model_name, train_input, train_output, test_input, test_output):
+    model.fit(train_input, train_output)
+    filename = f'{MODEL_DIR}\\{model_name}.sav'
+    pickle.dump(model, open(filename, 'wb'))
+    avg = test(model, test_input, test_output)
+    return avg
+    
+#WORD REPRESENTATIONS
+
+#represent text with bag of words
+def bow(text, wordlist):
+    return [int(word in text) for word in wordlist]
+
+#represent text with frequencies
+def freq(text, wordlist):
+    text = text.split()
+    counter = collections.Counter(text)
+    length = len(text)
+    return [value_or_zero(word, counter)/length for word in wordlist]
+
+#represent text with tf-idf
+def tfidf(text, word_dict):
+    text = text.split()
+    counter = collections.Counter(text)
+    length = len(text)
+    frequencies = [value_or_zero(word, counter)/length for word in word_dict]
+    return [word_dict[word] * frequencies[i] for (i, word) in enumerate(word_dict)]
+    
+#/WORD REPRESENTATIONS
+
+def main():
+
+    #Data preprocessing
+    train_df = pd.read_csv(f'{BASIC}\\train.csv')
+    train_df['text'] = train_df['text'].apply(clean).astype(str)
[…]

### 2022-11-25 reorg_scripts/train_test.py

@@ -1,0 +1,15 @@
+import pandas as pd
+
+BASIC = '..\\data'
+
+def main():
+    df = pd.read_csv(f'{BASIC}\\antisemitism_dataset.csv')
+    df = df.dropna(subset=['text'])
+    train_df = df.sample(frac=0.9)
+    test_df = df.drop(index=train_df.index)
+    
+    train_df.to_csv(f'{BASIC}\\train.csv', index=False)
+    test_df.to_csv(f'{BASIC}\\test.csv', index=False)
+
+if __name__ == '__main__':
+    main()

### 2022-11-26 writeups/Observations.tex
fit type classifiers

@@
 
-{\bf \emph{Observations on the Datasets}} \hfill Nicolas Cloutier\\\\
+{\bf \emph{Data Science Techniques for Automatic Antisemitism Detection}} \hfill Nicolas Cloutier
+
+Last edited \today.\\
 
@@
 {\bf Method} & \bf{Task} & \bf{Accuracy} & \bf{Chance Accuracy}\\
-\hline
- Vector distance & Boolean classification & 0.76 & 0.5\\ 
- Vector distance & Type classification & 0.57 & 0.25\\ 
- Cosine similarity & Boolean classification & 0.76 & 0.5\\ 
- Cosine similarity & Type classification & 0.57 & 0.25\\ 
+\hline Vector distance & Boolean classification & 0.76 & 0.5\\ 
+\hline Vector distance & Type classification & 0.57 & 0.25\\ 
+\hline Cosine similarity & Boolean classification & 0.76 & 0.5\\ 
+\hline Cosine similarity & Type classification & 0.57 & 0.25\\ 
  \hline
@@
 \hline {\bf Dataset} & {\bf Bool. avg.} & {\bf Cos. Bool. avg.} & {\bf 1 \#} & {\bf 2 \#} & {\bf 3 \#} & {\bf 4 \#}\\
-\hline 
-4Chan data & 0.85 & 0.85 & 504 & 575 & 1474 & 793\\
-Normal tweets & 0.60 & 0.60 & 3675 & 1357 & 2419 & 3549\\
-Labeled data & 0.44 & 0.44 & 515 & 133 & 277 & 296\\
+\hline 4Chan data & 0.85 & 0.85 & 504 & 575 & 1474 & 793\\
+\hline Normal tweets & 0.60 & 0.60 & 3675 & 1357 & 2419 & 3549\\
+\hline Labeled data & 0.44 & 0.44 & 515 & 133 & 277 & 296\\
 \hline
@@
 Looking at the Boolean averages for the normal data, it appears to be higher than may have been anticipated or hoped for, given that the tweets were collected from a random sample of tweets across 2021.
-Interestingly, the Boolean average of the normal tweets is higher than the Boolean average of the labeled data, which contained antisemitic tweets.
+Interestingly, the Boolean average of the normal tweets is higher than the Boolean average of the labeled data, which contained antisemitic tweets.\\
+
+{\bf Machine Learning Models}
+
+Three machine learning models were used: SVMs, Decision Trees, and Naïve Bayes classifiers.
+Each of those models was used for Boolean and type classification and with three seperate text representations (Bag-of-Words, raw frequency, and TF-IDF), making a total of 12 models.
+For the sake of evaluation, the rules-based model was retrained on only the training data, then tested on the test data.
+Models labeled RB are rules-based.
+The scores represent the proportion of the test dataset that the model correctly predicted, 1 means perfect predictions, 0 means entirely incorrect predictions.
+\begin{center}
+Table 3: Machine learning model evaluations and comparisons to rules-based model
+\begin{tabular}{ |c|c|c|c| }
+\hline {\bf Model type} & {\bf Text representation} & {\bf Boolean score} & {\bf Type score} \\
+\hline Decision Tree & BoW & 0.653 &  0.513\\
+\hline Decision Tree & Frequency & 0.695  &  0.692\\
+\hline Decision Tree & TF-IDF & 0.703 & 0.487\\
+\hline SVM & BoW & 0.703 & 0.692\\
+\hline SVM & Frequency & 0.703 & 0.564\\
+\hline SVM & TF-IDF & 0.678 & 0.462\\
+\hline Naïve Bayes & BoW & 0.669 & 0.769\\
+\hline Naïve Bayes & Frequency & 0.602 & 0.487\\
+\hline Naïve Bayes & TF-IDF & 0.602 & 0.462\\
+\hline Cosine Similarity (RB) & Frequency & 0.602 & 0.462\\
+\hline Vector Distance (RB) & Frequency & 0.602 & 0.462\\
+\hline
+\end{tabular}
+\end{center}
 \end{document}

### 2022-11-26 ml_models/ml_models.py

@@ -60,20 +60,11 @@ def tfidf(text, word_dict):
     
 #/WORD REPRESENTATIONS
 
-def main():
-
-    #Data preprocessing
-    train_df = pd.read_csv(f'{BASIC}\\train.csv')
-    train_df['text'] = train_df['text'].apply(clean).astype(str)
-    train_df = train_df[train_df['text'].apply(lambda x: len(x.split()) >= 1)]
-    test_df = pd.read_csv(f'{BASIC}\\test.csv')
-    test_df['text'] = test_df['text'].apply(clean).astype(str)
-    test_df = test_df[test_df['text'].apply(lambda x: len(x.split()) >= 1)]
-    
+def process_fit_test(train_df, test_df, output):
     train_input = train_df['text']
-    train_output = train_df['classification']
+    train_output = train_df[output]
     test_input = test_df['text']
-    test_output = test_df['classification']
+    test_output = test_df[output]
     
     #get wordlist
     all_text = ' '.join(list(train_df['text']) + list(test_df['text']))
@@ -111,11 +102,29 @@ def main():
             algo_name = algorithm['name']
             algo_model = algorithm['model']
         
-            score = fit_save_test(algo_model, f'{algo_name}-{rep_name}', 
+            score = fit_save_test(algo_model, f'{algo_name}-{rep_name}-{output}', 
                                   temp_train_input, temp_train_output,
                                   temp_test_input, temp_test_output)
             
-            print(f'Finished training {algo_name} model with {rep_name} representation. Scored {score:.3f}.')
+            print(f'Finished training {algo_name} model with {rep_name} representation, predicting {output}. Scored {score:.3f}.')
+    
+
+def main():
+
+    #Data preprocessing
+    train_df = pd.read_csv(f'{BASIC}\\train.csv')
+    train_df['text'] = train_df['text'].apply(clean).astype(str)
+    train_df = train_df[train_df['text'].apply(lambda x: len(x.split()) >= 1)]
+    test_df = pd.read_csv(f'{BASIC}\\test.csv')
+    test_df['text'] = test_df['text'].apply(clean).astype(str)
+    test_df = test_df[test_df['text'].apply(lambda x: len(x.split()) >= 1)]
+    
+    process_fit_test(train_df, test_df, 'classification')
+    
+    train_df = train_df[train_df['classification'] == 1]
+    test_df = test_df[test_df['classification'] == 1]
+    
+    process_fit_test(train_df, test_df, 'type_of_antisemitism')
 
 if __name__ == '__main__':
     main()

### 2022-11-26 reorg_scripts/make_counters.py

@@ -51,7 +51,7 @@ def get_counter(type, df, typename='type', baseline=pd.DataFrame()):
 
 def main():
 
-    df = pd.read_csv('..\\data\\antisemitism_dataset.csv')
+    df = pd.read_csv('..\\data\\train.csv')
     
     #reconfigure/reorganize dataframe
     df['classification'] = df['classification'].astype(int)

### 2022-11-26 rulesbased_model/dataset.py

@@ -20,7 +20,7 @@ def main():
         type_dict[str(x)] = pd.read_csv(f'{FREQ_DIR}\\{x}.csv')
     
     #apply to files
-    files = ['normal_tweets', 'cleaned_data', 'antisemitism_dataset']
+    files = ['normal_tweets', 'cleaned_data', 'test']
     for filename in files:
         df = pd.read_csv(f'{DATA_DIR}\\{filename}.csv')
         df['is_antisemitic'] = df['text'].apply(lambda x: int(rb.is_antisemitic(x, df_dict=is_antisemitic_dict)))

### 2022-11-29 writeups/Observations.tex
add tf-idf rules-based model

@@
 Three machine learning models were used: SVMs, Decision Trees, and Naïve Bayes classifiers.
-Each of those models was used for Boolean and type classification and with three seperate text representations (Bag-of-Words, raw frequency, and TF-IDF), making a total of 12 models.
+Each of those models was used for Boolean and type classification and with three seperate text representations (Bag-of-Words, raw frequency, and TF-IDF), making a total of 18 models.
 For the sake of evaluation, the rules-based model was retrained on only the training data, then tested on the test data.
@@
 \hline Vector Distance (RB) & Frequency & 0.602 & 0.462\\
+\hline Vector Distance (RB) & TF-IDF & 0.602 & 0.462\\
 \hline

### 2022-11-29 ml_models/ml_models.py

@@ -80,6 +80,7 @@ def process_fit_test(train_df, test_df, output):
     rep_methods = [{'name': 'bow', 'function': bow, 'wordlist': wordlist}, #methods for representing text
                    {'name': 'freq', 'function': freq, 'wordlist': wordlist},
                    {'name': 'tfidf', 'function': tfidf, 'wordlist': word_dict}]
+    
     algorithms = [{'name': 'decision-tree', 'model': tree.DecisionTreeClassifier()}, #algorithms for classification
                   {'name': 'svm', 'model': svm.SVC()},
                   {'name': 'naive-bayes', 'model': naive_bayes.GaussianNB()}]

### 2022-11-29 reorg_scripts/make_counters.py

@@ -19,12 +19,20 @@ def to_df(text_counter, column1, column2):
     output_df = output_df.sort_values(column2, ascending=False) #sort
     return output_df
 
+#get idf part of tfidf
+def get_idf(text_counter, length):
+    for word in text_counter:
+        count = text_counter[word]
+        value = math.log(length/count)
+        text_counter[word] = value
+    return text_counter
+
 #get a counter of the percentage of a text taken up by the words in that text.
 #if a baseline counter is provided to compare, subtract it.
-def get_counter(type, df, typename='type', baseline=pd.DataFrame()):
+def get_counter(df, type=None, typename='type', baseline=pd.DataFrame()):
 
     #put the text of the rows of a certain type in a string, and get rid of stopwords
-    df = df[df[typename] == type]
+    if type: df = df[df[typename] == type]
     text = ' '.join(list(df['text'].astype(str)))
     text = f' {text} '
     text = delete_punctuation(text.lower())
@@ -36,7 +44,7 @@ def get_counter(type, df, typename='type', baseline=pd.DataFrame()):
     text_counter = collections.Counter(split_text)
     length = len(split_text)
     for key in text_counter:
-        text_counter[key] = text_counter[key] * 100 / length
+        text_counter[key] = text_counter[key] / length
     
     #if a baseline is provided, subtract the value from each word in the baseline
     #from the same word in the current text counter
@@ -49,6 +57,7 @@ def get_counter(type, df, typename='type', baseline=pd.DataFrame()):
     
     return text_counter
 
+
 def main():
 
     df = pd.read_csv('..\\data\\train.csv')
@@ -58,13 +67,18 @@ def main():
     df['text'] = df['text'].astype(str)
     df['text'] = df['text'].apply(delete_punctuation) #delete punctuation from the text
     
+    all_text = ' '.join(list(df['text'])).split()
+    all_counter = get_idf(get_counter(df), len(all_text))
+    all_counter_df = to_df(all_counter, COLUMN1, COLUMN2)
+    all_counter_df.to_csv('..\\rulesbased_model\\frequencies\\all.csv', index=False)
+    
     #get counters and write to file for antisemitic text
-    all_antisemitic = get_counter(1, df, typename='classification') #get a baseline percentage counter (this is the counter for the non-antisemitic text)
[…]

### 2022-11-29 rulesbased_model/dataset.py

@@ -1,10 +1,14 @@
 import pandas as pd
 import rulesbased as rb
+import time
+
+#This script tests the rules-based models on the testing data.
 
 DATA_DIR = '..\\data'
 FREQ_DIR = 'frequencies'
 
 def main():
+    start = time.time()
     types = {'political': 0,
              'economic': 1,
              'religious': 2,
@@ -14,20 +18,43 @@ def main():
     is_antisemitic_dict = {'baseline': pd.read_csv(f'{FREQ_DIR}\\baseline.csv'),
                            'antisemitic': pd.read_csv(f'{FREQ_DIR}\\antisemitic.csv')}
     
+    tfidf_df = pd.read_csv('.\\frequencies\\all.csv')
+    
     #get dataframes for types
     type_dict = dict()
     for x in range(1, 5):
         type_dict[str(x)] = pd.read_csv(f'{FREQ_DIR}\\{x}.csv')
     
+    end = time.time()
+    print(f'Setup completed in {end - start:.3f} seconds.')
+    
     #apply to files
-    files = ['normal_tweets', 'cleaned_data', 'test']
+    files = ['test', 'cleaned_data', 'normal_tweets']
     for filename in files:
+        file_start = time.time()
         df = pd.read_csv(f'{DATA_DIR}\\{filename}.csv')
+        
+        start = time.time()
         df['is_antisemitic'] = df['text'].apply(lambda x: int(rb.is_antisemitic(x, df_dict=is_antisemitic_dict)))
+        end = time.time()
+        print(f'Binary freq completed in {int((end - start)//60)} minutes and {(end-start)-(end-start)//60*60:.2f} seconds.')
+        start = time.time()
         df['predicted_type'] = df['text'].apply(lambda x: types[rb.type(x, df_dict=type_dict)])
-        df['cos_is_antisemitic'] = df['text'].apply(lambda x: int(rb.cosine_is_antisemitic(x, df_dict=is_antisemitic_dict)))
-        df['cos_predicted_type'] = df['text'].apply(lambda x: types[rb.cosine_type(x, df_dict=type_dict)])
+        end = time.time()
+        print(f'Type freq completed in {int((end - start)//60)} minutes and {(end-start)-(end-start)//60*60:.2f} seconds.')
+        
+        start = time.time()
[…]

### 2022-11-29 rulesbased_model/rulesbased.py

@@ -7,6 +7,7 @@ import string
 
 WORD_COLUMN = 'Words'
 PERCENT_COLUMN = 'Percent Appearance'
+TFIDF = '.\\frequencies\\all.csv'
 
 vector_distance = lambda v1, v2: math.sqrt(sum([(v1[i]-v2[i])**2 for (i, _) in enumerate(v1)])) #distance between two vectors equation
 dot = lambda v1, v2: sum([v1[i]*v2[i] for (i, _) in enumerate(v1)])
@@ -32,7 +33,7 @@ def get_text_counter(text):
     text_counter = collections.Counter(split_text)
     length = len(split_text)
     for key in text_counter:
-        text_counter[key] = text_counter[key] * 100 / length
+        text_counter[key] = text_counter[key] / length
     
     text_counter = collections.OrderedDict(text_counter.most_common())
     
@@ -54,7 +55,7 @@ def get_counter(type, df):
     text_counter = collections.Counter(split_text)
     length = len(split_text)
     for key in text_counter:
-        text_counter[key] = text_counter[key] * 100 / length
+        text_counter[key] = text_counter[key] / length
     
     text_counter = collections.OrderedDict(text_counter.most_common())
     
@@ -86,6 +87,49 @@ def to_vector(text_df, comparison_dict, column1, column2):
         
         comparison_vector = list(comparison_df[column2])
         comparison_vector_dict[key] = comparison_vector
+    
+    return text_vector, comparison_vector_dict
+    
+#do same as above, but with tfidf
+def to_vector_tfidf(text_df, comparison_dict, tfidf_df, column1, column2):
+    comparison_vector_dict = dict()
+    text_df = text_df.sort_values(column1) #sort alphabetically (so they're in the same order)
+    text_vector = list(text_df[column2])
+    words = list(text_df[column1]) #words in tweet
+    
+    tfidf_words = list(tfidf[column1])
+    non_overlap_words = [word for word in words if word not in tfidf_words] #words in tweet, not in dataset
+    zeroes = [0 for word in non_overlap_words]
+    non_overlap_df = pd.DataFrame([non_overlap_words, zeroes], [column1, column2]).transpose()
+    comparison_df = pd.concat([comparison_df, non_overlap_df])
+    
+    #delete values in dataset that aren't in text
+    comparison_df = comparison_df.loc[comparison_df[column1].isin(words)]
[…]

### 2022-11-30 reorg_scripts/train_test.py
increase test dataset size

@@ -5,7 +5,7 @@ BASIC = '..\\data'
 def main():
     df = pd.read_csv(f'{BASIC}\\antisemitism_dataset.csv')
     df = df.dropna(subset=['text'])
-    train_df = df.sample(frac=0.9)
+    train_df = df.sample(frac=0.75)
     test_df = df.drop(index=train_df.index)
     
     train_df.to_csv(f'{BASIC}\\train.csv', index=False)

### 2022-11-30 rulesbased_model/dataset.py

@@ -29,7 +29,7 @@ def main():
     print(f'Setup completed in {end - start:.3f} seconds.')
     
     #apply to files
-    files = ['test', 'cleaned_data', 'normal_tweets']
+    files = ['test']
     for filename in files:
         file_start = time.time()
         df = pd.read_csv(f'{DATA_DIR}\\{filename}.csv')

### 2022-12-08 ml_models/ml_models.py
add k-fold xvalidation

@@ -8,6 +8,9 @@ import pickle
 
 BASIC = '..\\data'
 MODEL_DIR = '.\\models'
+NUMBER_PARTITIONS = 10
+COLUMNS = ['Algorithm', 'Representation', 'Task', 'Score', 'Partition']
+report_df = pd.DataFrame(columns=COLUMNS) #df that will be written to a report file with info on model performance
 
 value_or_zero = lambda x, dict: 0 if not x in dict else dict[x] #return dictionary value if key in dictionary, otherwise return 0
 delete_punctuation = lambda x: ''.join([char for char in x if char not in string.punctuation]) #delete punctuation from string
@@ -59,15 +62,41 @@ def tfidf(text, word_dict):
     return [word_dict[word] * frequencies[i] for (i, word) in enumerate(word_dict)]
     
 #/WORD REPRESENTATIONS
-
-def process_fit_test(train_df, test_df, output):
-    train_input = train_df['text']
-    train_output = train_df[output]
-    test_input = test_df['text']
-    test_output = test_df[output]
     
+#train all algorithms on a single representation method and single partition,
+#update report df
+def train_representations(train_input, train_output, test_input, 
+                          test_output, rep_method, algorithms, partition, task):
+    global report_df
+    
+    rep_name = rep_method['name']
+    rep_func = rep_method['function']
+    rep_list = rep_method['wordlist']
+    
+    #represent text with that method
+    temp_train_input = train_input.apply(lambda x: rep_func(x, rep_list)).values.tolist()
+    temp_train_output = train_output.values.tolist()
+    temp_test_input = test_input.apply(lambda x: rep_func(x, rep_list)).values.tolist()
+    temp_test_output = test_output.values.tolist()
+    
+    for algorithm in algorithms:
+        
+        algo_name = algorithm['name']
+        algo_model = algorithm['model']
+    
+        score = fit_save_test(algo_model, f'{algo_name}-{rep_name}-{task}', 
+                              temp_train_input, temp_train_output,
+                              temp_test_input, temp_test_output)
+        
+        print(f'Finished training {algo_name} model with {rep_name} representation, predicting {task}. Scored {score:.3f}.')
+        temp_report = pd.DataFrame([algo_name, rep_name, task, score, partition], COLUMNS).transpose()
+        report_df = pd.concat([report_df, temp_report])
+    
+#given a single partition, train and test each model with each rep method
+def train_and_test(train_input, train_output, test_input, 
[…]

### 2022-12-08 ml_models/summarize.py

@@ -1,0 +1,36 @@
+import pandas as pd
+
+#Summarize the model results
+
+def main():
+    df = pd.read_csv('model_report.csv')
+    
+    algorithms = ['svm', 'decision-tree', 'naive-bayes']
+    representations = ['bow', 'freq', 'tfidf']
+    tasks = ['classification', 'type_of_antisemitism']
+    
+    algo_rep_tasks = []
+    
+    for algorithm in algorithms:
+        for representation in representations:
+            algo_rep_tasks += [{'Algorithm': algorithm,
+                               'Representation': representation,
+                               'Task': task} for task in tasks]
+    
+    for algo_rep_task in algo_rep_tasks:
+        algo = algo_rep_task['Algorithm']
+        representation = algo_rep_task['Representation']
+        task = algo_rep_task['Task']
+        
+        temp_df = df[df['Algorithm'] == algo]
+        temp_df = temp_df[temp_df['Representation'] == representation]
+        temp_df = temp_df[temp_df['Task'] == task]
+        
+        score = temp_df['Score'].mean()
+        algo_rep_task['Score'] = score
+    
+    output_df = pd.DataFrame(algo_rep_tasks)
+    output_df.to_csv('summarize.csv', index=False)
+    
+if __name__ == '__main__':
+    main()

### 2022-12-20 ml_models/ml_models.py
add predictions dataframe

@@ -10,10 +10,24 @@ BASIC = '..\\data'
 MODEL_DIR = '.\\models'
 NUMBER_PARTITIONS = 10
 COLUMNS = ['Algorithm', 'Representation', 'Task', 'Score', 'Partition']
-report_df = pd.DataFrame(columns=COLUMNS) #df that will be written to a report file with info on model performance
 
 value_or_zero = lambda x, dict: 0 if not x in dict else dict[x] #return dictionary value if key in dictionary, otherwise return 0
 delete_punctuation = lambda x: ''.join([char for char in x if char not in string.punctuation]) #delete punctuation from string
+
+#concatenate two dictionaries
+def dict_concat(dict1, dict2):
+    for key in dict2:
+        dict1[key] = dict2[key]
+    return dict1
+
+#given two dictionaries with identical keys and list values,
+#concatenate the corresponding values lists from the second to
+#the first where the keys are the same
+def append_dict_values(dict1, dict2):
+    for key in dict1:
+        if key in dict2:
+            dict1[key] = dict1[key] + dict2[key]
+    return dict1
 
 #remove stopwords from text, put in lowercase, and delete punctuation
 def clean(text):
@@ -22,23 +36,14 @@ def clean(text):
     for stopword in nltk.corpus.stopwords.words('english'):
         text = text.replace(f' {stopword} ', ' ')
     return text
-
-#test single-variable output models,
-#output of 1 means every prediction correct,
-#0 means every prediction incorrect
-def test(model, test_input, test_output):
-    predictions = model.predict(test_input)
-    wrong = [int(prediction == test_output[i]) for (i, prediction) in enumerate(predictions)]
-    avg = sum(wrong) / len(wrong)
-    return avg
     
 #fit a model, save to file, and test
-def fit_save_test(model, model_name, train_input, train_output, test_input, test_output):
+def fit_save_test(model, model_name, train_input, train_output, test_input):
     model.fit(train_input, train_output)
     filename = f'{MODEL_DIR}\\{model_name}.sav'
     pickle.dump(model, open(filename, 'wb'))
-    avg = test(model, test_input, test_output)
-    return avg
+    predictions = model.predict(test_input)
+    return predictions
     
 #WORD REPRESENTATIONS
 
@@ -67,7 +72,7 @@ def tfidf(text, word_dict):
 #update report df
 def train_representations(train_input, train_output, test_input, 
                           test_output, rep_method, algorithms, partition, task):
[…]

### 2023-01-02 writeups/Observations.tex
increase test dataset size
add k-fold xvalidation
edit observations
run friedman tests on model outputs

@@
     25\
-  \end{array} } \right] = \vec{a} , \left| \vec{a} \right| \approx 61.19
+  \end{array} } \right] = \vec{a} , \left| \left| \vec{a} \right| \right| \approx 61.19
 \]
@@
     24.991652\\
-  \end{array} } \right] = \vec{b},\hfill \left| \vec{b} \right| \approx 61.13
+  \end{array} } \right] = \vec{b},\hfill \left| \left| \vec{b} \right| \right| \approx 61.13
 \]
@@
 The scores represent the proportion of the test dataset that the model correctly predicted, 1 means perfect predictions, 0 means entirely incorrect predictions.
+The models were evaluated using 10-fold cross-validation.
+Tables with rankings for the specific algorithms and text representations are given after.
+
 \begin{center}
-Table 3: Machine learning model evaluations and comparisons to rules-based model
-\begin{tabular}{ |c|c|c|c| }
-\hline {\bf Model type} & {\bf Text representation} & {\bf Boolean score} & {\bf Type score} \\
-\hline Decision Tree & BoW & 0.653 &  0.513\\
-\hline Decision Tree & Frequency & 0.695  &  0.692\\
-\hline Decision Tree & TF-IDF & 0.703 & 0.487\\
-\hline SVM & BoW & 0.703 & 0.692\\
-\hline SVM & Frequency & 0.703 & 0.564\\
-\hline SVM & TF-IDF & 0.678 & 0.462\\
-\hline Naïve Bayes & BoW & 0.669 & 0.769\\
-\hline Naïve Bayes & Frequency & 0.602 & 0.487\\
-\hline Naïve Bayes & TF-IDF & 0.602 & 0.462\\
-\hline Cosine Similarity (RB) & Frequency & 0.602 & 0.462\\
-\hline Vector Distance (RB) & Frequency & 0.602 & 0.462\\
-\hline Vector Distance (RB) & TF-IDF & 0.602 & 0.462\\
+Table 3: Machine learning model evaluations --- binary classification
+\begin{tabular}{ |c|c|c| }
+\hline {\bf Model type} & {\bf Text representation} & {\bf Score}\\
+\hline Decision Tree & BoW & 0.722\\
+\hline Decision Tree & Frequency & 0.756\\
+\hline Decision Tree & TF-IDF & 0.760\\
+\hline SVM & BoW & 0.769\\
+\hline SVM & Frequency & 0.757\\
+\hline SVM & TF-IDF & 0.732\\
+\hline Naïve Bayes & BoW & 0.638\\
+\hline Naïve Bayes & Frequency & 0.591\\
+\hline Naïve Bayes & TF-IDF & 0.599\\
 \hline
@@
 \end{center}
+
+\begin{center}
+Table 4: Algorithm comparison --- binary classification\\
+\begin{tabular}{ |c|c| }
+\hline {\bf Model type}& {\bf Score}\\
+\hline Support Vector Machine & 0.714\\
+\hline Decision Tree & 0.568\\
+\hline Naïve Bayes & 0.466\\
+\hline
+\end{tabular}
+\end{center}
+
+\begin{center}
+Table 5: Representation comparison --- binary classification\\
+\begin{tabular}{ |c|c| }
+\hline {\bf Representation}& {\bf Score}\\
+\hline TF-IDF & 0.550\\
+\hline Frequency & 0.516\\
+\hline Bag of Words & 0.507\\
+\hline
+\end{tabular}
+\end{center}
+
+Using a Friedman Chi Squared test, the p-value for table 3 is $6.30 \times 10^{-89}$, the p-value for the different algorithms is $4.39 \times 10^{-73}$, and the p-value for the different representations is $5.21 \times 10^{-10}$.
+
+\begin{center}
+Table 6: Machine learning model evaluations --- type classification\\
+\begin{tabular}{ |c|c|c| }
+\hline {\bf Model type} & {\bf Text representation} & {\bf Score} \\
+\hline Decision Tree & BoW & 0.576\\
+\hline Decision Tree & Frequency & 0.567\\
+\hline Decision Tree & TF-IDF & 0.555\\
+\hline SVM & BoW & 0.582\\
+\hline SVM & Frequency & 0.521\\
+\hline SVM & TF-IDF & 0.476\\
+\hline Naïve Bayes & BoW & 0.515\\
+\hline Naïve Bayes & Frequency & 0.538\\
+\hline Naïve Bayes & TF-IDF & 0.526\\
+\hline
+\end{tabular}
+\end{center}
+
+\begin{center}
+Table 7: Algorithm comparison --- type classification\\
+\begin{tabular}{ |c|c| }
+\hline {\bf Model type}& {\bf Score}\\
+\hline Support Vector Machine & 0.479\\
+\hline Decision Tree & 0.321\\
+\hline Naïve Bayes & 0.300\\
+\hline
+\end{tabular}
+\end{center}
+
+\begin{center}
+Table 8: Representation comparison --- type classification\\
+\begin{tabular}{ |c|c| }
+\hline {\bf Representation}& {\bf Score}\\
+\hline TF-IDF & 0.397\\
+\hline Bag of Words & 0.347\\
+\hline Frequency & 0.338\\
+\hline
+\end{tabular}
+\end{center}
+
 \end{document}

### 2023-01-02 ml_models/ml_models.py

@@ -5,6 +5,7 @@ import nltk
 import collections
 import math
 import pickle
+from scipy import stats
 
 BASIC = '..\\data'
 MODEL_DIR = '.\\models'
@@ -170,9 +171,37 @@ def main():
     binary_predictions_df = process_fit_test(df, 'classification')
     binary_predictions_df.to_csv('binary_predictions.csv', index=False)
     
+    friedman_result = stats.friedmanchisquare(
+                         binary_predictions_df['decision-tree-tfidf-classification'],
+                         binary_predictions_df['svm-tfidf-classification'],
+                         binary_predictions_df['naive-bayes-tfidf-classification'],
+                         binary_predictions_df['decision-tree-freq-classification'],
+                         binary_predictions_df['svm-freq-classification'],
+                         binary_predictions_df['naive-bayes-freq-classification'],
+                         binary_predictions_df['decision-tree-bow-classification'],
+                         binary_predictions_df['svm-bow-classification'],
+                         binary_predictions_df['naive-bayes-bow-classification'],
+                         )
+    
+    print(f'Binary result: {friedman_result}')
+    
     df = df[df['classification'] == 1]
     type_predictions_df = process_fit_test(df, 'type_of_antisemitism')
     type_predictions_df.to_csv('type_predictions.csv', index=False)
+    
+    friedman_result = stats.friedmanchisquare(
+                         type_predictions_df['decision-tree-tfidf-type_of_antisemitism'],
+                         type_predictions_df['svm-tfidf-type_of_antisemitism'],
+                         type_predictions_df['naive-bayes-tfidf-type_of_antisemitism'],
+                         type_predictions_df['decision-tree-freq-type_of_antisemitism'],
+                         type_predictions_df['svm-freq-type_of_antisemitism'],
+                         type_predictions_df['naive-bayes-freq-type_of_antisemitism'],
+                         type_predictions_df['decision-tree-bow-type_of_antisemitism'],
+                         type_predictions_df['svm-bow-type_of_antisemitism'],
+                         type_predictions_df['naive-bayes-bow-type_of_antisemitism'],
+                         )
+    
+    print(f'Type result: {friedman_result}')
     
 
 if __name__ == '__main__':

### 2023-01-03 writeups/Observations.tex
fix model evaluation

@@
 The models were evaluated using 10-fold cross-validation.
-Tables with rankings for the specific algorithms and text representations are given after.
+Tables with rankings for the specific algorithms and text representations are given after, these were taken by getting the accuracy scores of the most common responses given by the models that use that representation or algorithm.
 
@@
 \hline {\bf Model type}& {\bf Score}\\
-\hline Support Vector Machine & 0.714\\
-\hline Decision Tree & 0.568\\
-\hline Naïve Bayes & 0.466\\
+\hline Decision Tree & 0.767\\
+\hline Support Vector Machine & 0.754\\
+\hline Naïve Bayes & 0.600\\
 \hline
@@
 \hline {\bf Representation}& {\bf Score}\\
-\hline TF-IDF & 0.550\\
-\hline Frequency & 0.516\\
-\hline Bag of Words & 0.507\\
+\hline Bag of Words & 0.766\\
+\hline Frequency & 0.761\\
+\hline TF-IDF & 0.749\\
 \hline
@@
 
-Using a Friedman Chi Squared test, the p-value for table 3 is $6.30 \times 10^{-89}$, the p-value for the different algorithms is $4.39 \times 10^{-73}$, and the p-value for the different representations is $5.21 \times 10^{-10}$.
+Using a Friedman Chi Squared test, the \emph{p}-value for table 3 is $6.30 \times 10^{-89}$, the \emph{p}-value for the different algorithms is $1.98 \times 10^{-59}$, and the \emph{p}-value for the different representations is $1.18 \times 10^{-16}$.
 
@@
 \hline {\bf Model type}& {\bf Score}\\
-\hline Support Vector Machine & 0.479\\
-\hline Decision Tree & 0.321\\
-\hline Naïve Bayes & 0.300\\
+\hline Decision Tree & 0.582\\
+\hline Naïve Bayes & 0.535\\
+\hline Support Vector Machine & 0.524\\
 \hline
@@
 \hline {\bf Representation}& {\bf Score}\\
-\hline TF-IDF & 0.397\\
-\hline Bag of Words & 0.347\\
-\hline Frequency & 0.338\\
+\hline Frequency & 0.588\\
+\hline Bag of Words & 0.585\\
+\hline TF-IDF & 0.541\\
 \hline
@@
 
+Using a Friedman Chi Squared test, the \emph{p}-value for table 6 is $2.52 \times 10^{-29}$, the \emph{p}-value for the different algorithms is $1.00 \times 10^{-11}$, and the \emph{p}-value for the different representations is $0.0014$.
+
 \end{document}

### 2023-01-04 writeups/Observations.tex
add multinomial naive bayes

@@
 \hline {\bf Model type} & {\bf Text representation} & {\bf Score}\\
-\hline Decision Tree & BoW & 0.722\\
-\hline Decision Tree & Frequency & 0.756\\
-\hline Decision Tree & TF-IDF & 0.760\\
-\hline SVM & BoW & 0.769\\
-\hline SVM & Frequency & 0.757\\
-\hline SVM & TF-IDF & 0.732\\
-\hline Naïve Bayes & BoW & 0.638\\
-\hline Naïve Bayes & Frequency & 0.591\\
-\hline Naïve Bayes & TF-IDF & 0.599\\
+\hline SVM & BoW & 0.774\\
+\hline SVM & Frequency & 0.760\\
+\hline Decision Tree & Frequency & 0.747\\
+\hline Decision Tree & TF-IDF & 0.739\\
+\hline Multinomial NB & BoW & 0.732\\
+\hline SVM & TF-IDF & 0.729\\
+\hline Decision Tree & BoW & 0.720\\
+\hline Multinomial NB & Frequency & 0.718\\
+\hline Multinomial NB & TF-IDF & 0.630\\
+\hline Gaussian NB & BoW & 0.630\\
+\hline Gaussian NB & TF-IDF & 0.602\\
+\hline Gaussian NB & Frequency & 0.600\\
 \hline
@@
 \hline {\bf Model type}& {\bf Score}\\
-\hline Decision Tree & 0.767\\
-\hline Support Vector Machine & 0.754\\
-\hline Naïve Bayes & 0.600\\
+\hline Support Vector Machine & 0.757\\
+\hline Decision Tree & 0.749\\
+\hline Multinomial NB & 0.728\\
+\hline Gaussian NB & 0.707\\
 \hline
@@
 \hline {\bf Representation}& {\bf Score}\\
-\hline Bag of Words & 0.766\\
-\hline Frequency & 0.761\\
-\hline TF-IDF & 0.749\\
+\hline Bag of Words & 0.767\\
+\hline TF-IDF & 0.742\\
+\hline Frequency & 0.736\\
 \hline
@@
 
-Using a Friedman Chi Squared test, the \emph{p}-value for table 3 is $6.30 \times 10^{-89}$, the \emph{p}-value for the different algorithms is $1.98 \times 10^{-59}$, and the \emph{p}-value for the different representations is $1.18 \times 10^{-16}$.
+Using a Friedman Chi Squared test, the \emph{p}-value for table 3 is approximately 0, the \emph{p}-value for the different algorithms is $6.42 \times 10^{-33}$, and the \emph{p}-value for the different representations is $3.12 \times 10^{-29}$.
 
@@
 \hline {\bf Model type} & {\bf Text representation} & {\bf Score} \\
-\hline Decision Tree & BoW & 0.576\\
-\hline Decision Tree & Frequency & 0.567\\
-\hline Decision Tree & TF-IDF & 0.555\\
-\hline SVM & BoW & 0.582\\
-\hline SVM & Frequency & 0.521\\
-\hline SVM & TF-IDF & 0.476\\
-\hline Naïve Bayes & BoW & 0.515\\
-\hline Naïve Bayes & Frequency & 0.538\\
-\hline Naïve Bayes & TF-IDF & 0.526\\
+\hline Multinomial NB & TF-IDF & 0.612\\
+\hline Multinomial NB & BoW & 0.597\\
+\hline SVM & BoW & 0.594\\
+\hline Decision Tree & TF-IDF & 0.588\\
+\hline Decision Tree & Frequency & 0.576\\
+\hline SVM & Frequency & 0.556\\
+\hline Decision Tree & BoW & 0.550\\
+\hline Gaussian NB & BoW & 0.541\\
+\hline Gaussian NB & Frequency & 0.538\\
+\hline Gaussian NB & TF-IDF & 0.517\\
+\hline SVM & TF-IDF & 0.509\\
+\hline Multinomial NB & Frequency & 0.494\\
 \hline
@@
 \hline {\bf Model type}& {\bf Score}\\
-\hline Decision Tree & 0.582\\
-\hline Naïve Bayes & 0.535\\
-\hline Support Vector Machine & 0.524\\
+\hline Multinomial NB & 0.606\\
+\hline Decision Tree & 0.585\\
+\hline Gaussian NB & 0.579\\
+\hline Support Vector Machine & 0.550\\
 \hline
@@
 \hline {\bf Representation}& {\bf Score}\\
-\hline Frequency & 0.588\\
-\hline Bag of Words & 0.585\\
-\hline TF-IDF & 0.541\\
+\hline Bag of Words & 0.600\\
+\hline TF-IDF & 0.576\\
+\hline Frequency & 0.547\\
 \hline
@@
 
-Using a Friedman Chi Squared test, the \emph{p}-value for table 6 is $2.52 \times 10^{-29}$, the \emph{p}-value for the different algorithms is $1.00 \times 10^{-11}$, and the \emph{p}-value for the different representations is $0.0014$.
+Using a Friedman Chi Squared test, the \emph{p}-value for table 6 is $4.29 \times 10^{-112}$, the \emph{p}-value for the different algorithms is $3.08 \times 10^{-28}$, and the \emph{p}-value for the different representations is $9.20 \times 10^{-16}$.
 

### 2023-01-04 ml_models/ml_models.py

@@ -5,7 +5,6 @@ import nltk
 import collections
 import math
 import pickle
-from scipy import stats
 
 BASIC = '..\\data'
 MODEL_DIR = '.\\models'
@@ -124,7 +123,8 @@ def train_and_test(train_input, train_output, test_input,
     
     algorithms = [{'name': 'decision-tree', 'model': tree.DecisionTreeClassifier()}, #algorithms for classification
                   {'name': 'svm', 'model': svm.SVC()},
-                  {'name': 'naive-bayes', 'model': naive_bayes.GaussianNB()}]
+                  {'name': 'naive-bayes', 'model': naive_bayes.GaussianNB()},
+                  {'name': 'naive-bayes-multinomial', 'model': naive_bayes.MultinomialNB()}]
     
     for rep_method in rep_methods:
         temp_predictions = dict_concat(train_representations(train_input, train_output, test_input, test_output, 
@@ -171,38 +171,9 @@ def main():
     binary_predictions_df = process_fit_test(df, 'classification')
     binary_predictions_df.to_csv('binary_predictions.csv', index=False)
     
-    friedman_result = stats.friedmanchisquare(
-                         binary_predictions_df['decision-tree-tfidf-classification'],
-                         binary_predictions_df['svm-tfidf-classification'],
-                         binary_predictions_df['naive-bayes-tfidf-classification'],
-                         binary_predictions_df['decision-tree-freq-classification'],
-                         binary_predictions_df['svm-freq-classification'],
-                         binary_predictions_df['naive-bayes-freq-classification'],
-                         binary_predictions_df['decision-tree-bow-classification'],
-                         binary_predictions_df['svm-bow-classification'],
-                         binary_predictions_df['naive-bayes-bow-classification'],
-                         )
-    
-    print(f'Binary result: {friedman_result}')
-    
     df = df[df['classification'] == 1]
     type_predictions_df = process_fit_test(df, 'type_of_antisemitism')
     type_predictions_df.to_csv('type_predictions.csv', index=False)
-    
-    friedman_result = stats.friedmanchisquare(
-                         type_predictions_df['decision-tree-tfidf-type_of_antisemitism'],
-                         type_predictions_df['svm-tfidf-type_of_antisemitism'],
-                         type_predictions_df['naive-bayes-tfidf-type_of_antisemitism'],
-                         type_predictions_df['decision-tree-freq-type_of_antisemitism'],
[…]

### 2023-01-04 ml_models/summarize.py

@@ -1,36 +1,0 @@
-import pandas as pd
-
-#Summarize the model results
-
-def main():
-    df = pd.read_csv('model_report.csv')
-    
-    algorithms = ['svm', 'decision-tree', 'naive-bayes']
-    representations = ['bow', 'freq', 'tfidf']
-    tasks = ['classification', 'type_of_antisemitism']
-    
-    algo_rep_tasks = []
-    
-    for algorithm in algorithms:
-        for representation in representations:
-            algo_rep_tasks += [{'Algorithm': algorithm,
-                               'Representation': representation,
-                               'Task': task} for task in tasks]
-    
-    for algo_rep_task in algo_rep_tasks:
-        algo = algo_rep_task['Algorithm']
-        representation = algo_rep_task['Representation']
-        task = algo_rep_task['Task']
-        
-        temp_df = df[df['Algorithm'] == algo]
-        temp_df = temp_df[temp_df['Representation'] == representation]
-        temp_df = temp_df[temp_df['Task'] == task]
-        
-        score = temp_df['Score'].mean()
-        algo_rep_task['Score'] = score
-    
-    output_df = pd.DataFrame(algo_rep_tasks)
-    output_df.to_csv('summarize.csv', index=False)
-    
-if __name__ == '__main__':
-    main()

### 2023-02-13 writeups/Observations.tex
add evaluations for specific classifications
perform post-hoc dunn tests

--- 2023-01-02 writeups/Observations.tex
@@
 This process is done for the combined text of all antisemitic tweets and non-antisemitic tweets, as well as individually for the tweets of each of the four types of antisemitism.
-I stored the resultant frequencies in a subdirectory of the ``rulesbased\texttt{\char`_}model'' directory entitled ``frequencies.'' In order to determine if a text is antisemitic, I do this same basic process to represent it as a series of frequencies.
+I stored the resultant frequencies in a subdirectory of the ``rulesbased-model'' directory entitled ``frequencies.'' In order to determine if a text is antisemitic, I do this same basic process to represent it as a series of frequencies.
 I then read the frequencies of the antisemitic and non-antisemitic datasets, removing any words from those datasets that do not appear in the input text and adding a new row with a frequency of 0 for words that appear in the input text but not the dataset.
@@
 Each of those models was used for Boolean and type classification and with three seperate text representations (Bag-of-Words, raw frequency, and TF-IDF), making a total of 18 models.
-For the sake of evaluation, the rules-based model was retrained on only the training data, then tested on the test data.
-Models labeled RB are rules-based.
-The scores represent the proportion of the test dataset that the model correctly predicted, 1 means perfect predictions, 0 means entirely incorrect predictions.
-The models were evaluated using 10-fold cross-validation.
-Tables with rankings for the specific algorithms and text representations are given after.
-
-\begin{center}
-Table 3: Machine learning model evaluations --- binary classification
-\begin{tabular}{ |c|c|c| }
-\hline {\bf Model type} & {\bf Text representation} & {\bf Score}\\
-\hline Decision Tree & BoW & 0.722\\
-\hline Decision Tree & Frequency & 0.756\\
-\hline Decision Tree & TF-IDF & 0.760\\
-\hline SVM & BoW & 0.769\\
-\hline SVM & Frequency & 0.757\\
-\hline SVM & TF-IDF & 0.732\\
-\hline Naïve Bayes & BoW & 0.638\\
-\hline Naïve Bayes & Frequency & 0.591\\
-\hline Naïve Bayes & TF-IDF & 0.599\\
-\hline
-\end{tabular}
-\end{center}
-
-\begin{center}
-Table 4: Algorithm comparison --- binary classification\\
-\begin{tabular}{ |c|c| }
-\hline {\bf Model type}& {\bf Score}\\
-\hline Support Vector Machine & 0.714\\
-\hline Decision Tree & 0.568\\
-\hline Naïve Bayes & 0.466\\
-\hline
-\end{tabular}
-\end{center}
-
-\begin{center}
-Table 5: Representation comparison --- binary classification\\
-\begin{tabular}{ |c|c| }
-\hline {\bf Representation}& {\bf Score}\\
-\hline TF-IDF & 0.550\\
-\hline Frequency & 0.516\\
-\hline Bag of Words & 0.507\\
-\hline
-\end{tabular}
-\end{center}
-
-Using a Friedman Chi Squared test, the p-value for table 3 is $6.30 \times 10^{-89}$, the p-value for the different algorithms is $4.39 \times 10^{-73}$, and the p-value for the different representations is $5.21 \times 10^{-10}$.
-
-\begin{center}
-Table 6: Machine learning model evaluations --- type classification\\
-\begin{tabular}{ |c|c|c| }
-\hline {\bf Model type} & {\bf Text representation} & {\bf Score} \\
-\hline Decision Tree & BoW & 0.576\\
-\hline Decision Tree & Frequency & 0.567\\
-\hline Decision Tree & TF-IDF & 0.555\\
-\hline SVM & BoW & 0.582\\
-\hline SVM & Frequency & 0.521\\
-\hline SVM & TF-IDF & 0.476\\
-\hline Naïve Bayes & BoW & 0.515\\
-\hline Naïve Bayes & Frequency & 0.538\\
-\hline Naïve Bayes & TF-IDF & 0.526\\
-\hline
-\end{tabular}
-\end{center}
-
-\begin{center}
-Table 7: Algorithm comparison --- type classification\\
-\begin{tabular}{ |c|c| }
-\hline {\bf Model type}& {\bf Score}\\
-\hline Support Vector Machine & 0.479\\
-\hline Decision Tree & 0.321\\
-\hline Naïve Bayes & 0.300\\
-\hline
-\end{tabular}
-\end{center}
-
-\begin{center}
-Table 8: Representation comparison --- type classification\\
-\begin{tabular}{ |c|c| }
-\hline {\bf Representation}& {\bf Score}\\
-\hline TF-IDF & 0.397\\
-\hline Bag of Words & 0.347\\
-\hline Frequency & 0.338\\
-\hline
-\end{tabular}
-\end{center}
+The evaluation metric used was the arithmetic mean of the recall for each individual classification in order to account for large class imbalances in the initial dataset.
+The models were then grouped by algorithm type and text representation and the most common answer was taken to find the differences between the classes.
+After this, a Cochran Q test was performed on the results, and, if they were significant, a post-hoc Dunn test was used to find the precise differences between the model types.
+This process was done for each task.
+The result of these tests can be found in the \texttt{cochran.ipynb} file in the \texttt{ml-models} directory.
 

### 2023-04-28 paper/paper.tex
add initial paper

\documentclass[runningheads]{llncs}
\usepackage{graphicx}

\begin{document}
\title{Generative LLM resampling for the class imbalance problem with hate speech detection}
\author{Nicolas Antonio Cloutier \and Nathalie Japkowicz\inst{1}}

\institute{American University, Washington, D.C., USA}

%research along the way
%4/26: outline
%4/27: introduction
%4/28: previous work
%4/29: methodology
%4/30: results
%5/1: discussion
%5/2: conclusions
%5/3: abstract and finalize bibliography

\maketitle

\begin{abstract}

\end{abstract}

\section{Introduction}
%P1: Introduce the issue of hate speech online
In recent years, hate speech has become increasingly mainstream and common within social media sites \cite{siegel}. This rise has not only caused online spaces to become less hospitable, but also had several offline effects. On top of having severe psychological effects on the recipient \cite{siegel}, online hate speech also played a roll in disseminating extremist anti-Rohingya voices leading to violence in Myanmar \cite{green}, and has provided motivation for several perpetrators of offline violent hate crimes \cite{siegel}. These events have motivated responses from numerous parties, including the social media sites themselves, that are increasingly looking to stop the spread of these messages \cite{ullmann}, and governmental bodies, that are seeking to regulate or prevent the spread of hate speech \cite{banks}. At the same time as hate speech has been becoming more common, social media sites have been generating more and more content, with popular social media site Twitter generating an average of 500 million tweets per day in 2019 \cite{pereira}.

%P2: Introduce the need for automated methods of hate speech detection
With these developments and the large amount of content on social media sites, these sites have been increasingly looking to automatic detection methods for hate speech \cite{ullmann}. These methods use Machine Learning (ML) to automatically detect and classify hateful speech, removing some of the work done by moderators, whose primary job is to respond reactively to user reports of hate speech \cite{ullmann}. Chandra et al. (2021) used a combination of image and text processing and classification algorithms to classify images and text on social media sites Twitter and Gab \cite{chandra}. We use their Twitter dataset to further analyze the presence of hate speech on Twitter and investigate new algorithms for classification.

%P3: Introduce the imbalanced data issue
One difficulty with this dataset is that it is imbalanced, meaning one classification is far more common in the dataset than another. Imbalanced data can negatively impact the performance of ML classification algorithms \cite{sun}, affecting numerous domains that use ML classification. Many ML algorithms are inadequately prepared to handle the class imbalance problem \cite{sun}, leading many to look to other solutions, including resampling methods, that in some way change the training data in order to allay the effects of the class imbalance problem \cite{japkowicz}.

%P4: Wrap up the research question and context
With this in mind, there are two crucial research questions this paper seeks to answer. First: how can automatic ML methods be improved in the domain of text classification for hate speech detection? Second: how can the class imbalance problem be dealt with for text data? These are the questions we seek to provide answers to, with the hope that they may inform future research and hate speech detection systems.

\section{Previous work}
%P1: mention some papers on antisemitism detection

%P2: mention big paper that you use

%P3: introduce previous work on class imbalance problem

%P4: talk about generative models used in vision, tabular data

%P5: focus on resampling methods

%P6: talk about LLMs and their applications, find whether there is research done with LLMs on class imbalance

\section{Methodology}
%P1: talk about the different tasks

%P2: talk about the different types of models trained

%P3: talk about the resampling methods used

%P4: talk about the augmented dataset specifically and testing

%P5: talk about evaluation (mean recall) and methods used (cochran, etc)

\section{Results}
%P1: introduce, discuss difference in results in binary and 4-class tasks

%P2: make diluted 4-class table, show Dunn test array (also for binary?)

%P3: directly compare aug to randomundersampling

%P4: talk about how many dimensions saw improvements with aug

%P5: summarize?

\section{Discussion}
%P1: talk about difference between binary and 4-class tasks

%P2: suggest reasons for this (authenticity of text samples)

%P3: suggest further lines of inquiry (testing of this on images, tabular, testing more texts)

\section{Conclusions}
%P1: talk about implications for antisemitism classification

%P2: talk about implications for the class imbalance problem

%P3: talk about implications for text data

\begin{thebibliography}{6}
\bibitem{banks}
Banks, J.: Regulating hate speech online. International Review of Law, Computers and Technology \textbf{24}(3), 233--239 (2010)

\bibitem{chandra}
Chandra, M., Pailla, D., Bhatia, H., Sanchawala, A., Gupta, M., Shrivastava, M., Kumaraguru, P.: ``Subverting the Jewtocracy'': Online Antisemitism Detection Using Multimodal Deep Learning.  ACM WEB SCIENCE CONFERENCE 2021, pp. 148-157.

\bibitem{green}
Green, P., MacManus, T., De la Cour Venning, A.: Countdown to Annihilation: Genocide in Myanmar. International State of Crime Initiative, London (2015)

\bibitem{japkowicz}
Japkowicz, N., Shaju, S.: The class imbalance problem: A systematic study. Intelligent Data Analysis \textbf{6}(5), 429--449 (2002)

\bibitem{pereira}
Pereira-Kohatsu, J. C., Quijano-Sánchez, L., Liberatore, F., Camacho-Collados, M.: Detecting and Monitoring Hate Speech in Twitter. Sensors \textbf{19}(21), 4654--4691 (2019)

\bibitem{siegel}
Siegel, A.: Online Hate Speech. Social media and democracy: The state of the field, prospects for reform, 56--88 (2020).

\bibitem{sun}
Sun, Y., Wong, A., Kamel, M.: Classification of Imbalanced Data: A Review. International Journal of Pattern Recognition and Artificial Intelligence \textbf{23}(4), 687--719 (2009)

\bibitem{ullmann}
Ullmann, S., Tomalin, M.: Quarantining online hate speech: technical and ethical perspectives. Ethics and Information Technology \textbf{22}, 69--80 (2020)

\end{thebibliography}
\end{document}

### 2023-04-28 paper/paper.tex
add previous work section

@@
 \begin{document}
-\title{Generative LLM resampling for the class imbalance problem with hate speech detection}
+\title{Generative LLM oversampling for the class imbalance problem in hate speech detection}
 \author{Nicolas Antonio Cloutier \and Nathalie Japkowicz\inst{1}}
 
-\institute{American University, Washington, D.C., USA}
+\institute{1. American University, Washington, D.C., USA}
 
@@
 Chandra et al.
-(2021) used a combination of image and text processing and classification algorithms to classify images and text on social media sites Twitter and Gab \cite{chandra}.
+\cite{chandra} used a combination of image and text processing and classification algorithms to classify images and text on social media sites Twitter and Gab \cite{chandra}.
 We use their Twitter dataset to further analyze the presence of hate speech on Twitter and investigate new algorithms for classification.
@@
 Imbalanced data can negatively impact the performance of ML classification algorithms \cite{sun}, affecting numerous domains that use ML classification.
-Many ML algorithms are inadequately prepared to handle the class imbalance problem \cite{sun}, leading many to look to other solutions, including resampling methods, that in some way change the training data in order to allay the effects of the class imbalance problem \cite{japkowicz}.
+Many ML algorithms are inadequately prepared to handle the class imbalance problem \cite{sun}, leading many to look to other solutions, including resampling methods, that in some way change the training data by adding or removing data points in order to allay the effects of the class imbalance problem \cite{japkowicz}.
 
@@
 \section{Previous work}
-%P1: mention some papers on antisemitism detection
+%P1: mention big paper that you use
+Antisemitism detection, as well as hate speech detection generally, has been the subject of much research.
+Martins et al.
+\cite{martins} used ML models to analyze and classify antisemitism in posts on social media sites Gab and 4chan, and González-Pizarro and Zanettou \cite{gonzalez} used large language models (LLMs) to do the same.
+Chandra et al.
+\cite{chandra} introduced a new dataset of labeled antisemitic posts, including text and images.
+These posts were labeled not only for whether or not they are antisemitic, but also their type of antisemitism, with the researchers grouping antisemitic posts into political, economic, religious, and racial antisemitism, and trained models to classify both antisemitic status and type of antisemitism.
+This Twitter dataset is imbalanced, with the non-antisemitic class far outnumbering the antisemitic class, and with political and racial antisemitism being more common than religious and economic antisemitism.
 
-%P2: mention big paper that you use
+%P2: introduce previous work on class imbalance problem
+Imbalance is a common issue in ML classification problems.
+The class imbalance problem can severely negatively affect model predictive power, with the less common class (termed the ``minority class'') is often misclassified due to its low prevalence in the dataset relative to the larger class (the ``majority class'') \cite{abdelrahman}.
+This problem has affected fields as distinct as medicine, fraudulent call detection, and risk management \cite{sun}.
+Due to the severity with which this problem can limit model performance and its widespread nature, being seen in numerous distinct fields, it has become a focus of researchers as an area of improvement \cite{abdelrahman}, with numerous techniques being created to allay the effects of the problem.
 
-%P3: introduce previous work on class imbalance problem
+%P3: focus on resampling methods
+One such technique is resampling, altering the training data by adding new data points or taking existing data points away in order to improve model performance.
+Generally, there are two types of resampling: oversampling, which describes the process of adding synthetic data points to a dataset, and undersampling, which describes the process of removing existing data points from a dataset \cite{shelke}.
+Resampling techniques can improve the performance of ML models when trained on imbalanced data \cite{lee} \cite{khushi}, advancing the state of the field of ML in problems with unbalanced datasets.
 
 %P4: talk about generative models used in vision, tabular data
+One method for oversampling is the use of generative models, including generative adversarial networks (GANs) and autoencoders to generate synthetic data.
+These methods seek to match the distribution of the original dataset and create synthetic datapoints in accordance with that distribution \cite{hao}, with the goal of creating authentic synthetic examples for model training.
+They have achieved success in their applications to primarily computer vision and tabular data \cite{hao} \cite{engelmann} \cite{bellinger} \cite{dai}.
+Applications of these generative models to NLP do occur \cite{phung}, they are generally less common than applications to other areas.
+Natural language processing (NLP) tends to use more traditional statistical methods for oversampling such as SMOTE and random oversampling \cite{wijaya} \cite{glazkova}.
 
-%P5: focus on resampling methods
-
-%P6: talk about LLMs and their applications, find whether there is research done with LLMs on class imbalance
+%P5: talk about LLMs and their applications, find whether there is research done with LLMs on class imbalance
+While these advances in resampling have been occuring, similar advances have been made in generative LLMs, such as the GPT series of models from OpenAI, which have the ability to generate human-like text \cite{floridi} and been applied to patent claim generation \cite{hsiang}, healthcare education \cite{sallam}.
+These LLMs, despite their ability of producing authentic examples given prompts, have not been widely investigated as a method for oversampling text data.
 
@@
 
-\begin{thebibliography}{6}
+\begin{thebibliography}{24}
+\bibitem{abdelrahman}
+Abd Elrahman, S. M., Abraham, A: A review of class imbalance problem. Journal of Network and Innovative Computing \textbf{1}, 332--340 (2013)
+
 \bibitem{banks}
 Banks, J.: Regulating hate speech online. International Review of Law, Computers and Technology \textbf{24}(3), 233--239 (2010)
+
+\bibitem{bellinger}
+Bellinger, C., Japkowicz, N., Drummond, C.: Synthetic oversampling for advanced radioactive threat detection. IEEE CONFERENCE ON MACHINE LEARNING AN APPLICATIONS 2015.
 
@@
 
+\bibitem{dai}
+Dai, W., Ng, K., Severson, K., Huan, W., Anderson, F., Stultz, C.: Generative oversampling with a contrastive variational autoencoder. IEEE INTERNATIONAL CONFERENCE ON DATA MINING 2019, pp. 101-109.
+
+\bibitem{engelmann}
+Engelmann, J., Lessmann, S.: Conditional Wasserstein GAN-based oversampling of tabular data for imbalanced learning. Expert Systems with Applications \textbf{174}, (2021)
+
+\bibitem{floridi}
+Floridi, L., Chiriatti, M.: GPT-3: Its nature, scope, limits, and consequences. Minds and Machines \textbf{30}, 681--694 (2020)
+
+\bibitem{glazkova}
+Glazkova, A.: A comparison of synthetic oversampling methods for multi-class text classification. \emph{arXiv preprint arXiv:2008.04636t}, (2020)
+
+\bibitem{gonzalez}
+González-Pizarro, F., Zannettou, S.: Understanding and detecting hateful content using contrastive learning. \emph{arXiv preprint arXiv:2201.08387}, (2022)
+
 \bibitem{green}
 Green, P., MacManus, T., De la Cour Venning, A.: Countdown to Annihilation: Genocide in Myanmar. International State of Crime Initiative, London (2015)
+
+\bibitem{hao}
+Hao, J., Wang, C., Zhang, H., Yang, G.: Annealing genetic GAN for minority oversampling. \emph{arXiv preprint}, (2020)
 
@@
 
+\bibitem{khushi}
+Khushi, M., Shaukat, K., Alam, T.
+M., Hammed, I.
+A., Uddin, S., Luo, S., Yang, X., Consuelo Reyes, M.: A Comparitive Performance Analysis of Data Resampling Methods on Imbalance Medical Data.
+IEEE Access \textbf{9}, 109960--109975 (2021)
+
+\bibitem{hsiang} %the main author is also named lee, this is the second author
+Lee, J. S., Hsiang, J.: Patent Claim Generation by Fine-Tuning OpenAI GPT-2. World Patent Information \textbf{62}, (2020)
+
+\bibitem{lee}
+Lee, P. H.: Resampling methods improve the predictive power of modeling in class-imbalanced datasets. International Journal of Environmental Research and Public Health \textbf{11}(9), 9776--9789.
+
+\bibitem{martins}
+Martins, R., Gomes, M., Almeida, J. J., Novais, P., Henriques, P.: Hate speech classification in social media using emotional analysis. BRAZILIAN CONFERENCE ON INTELLIGENT SYSTEMS 2018, pp. 61-66
+
 \bibitem{pereira}
 Pereira-Kohatsu, J. C., Quijano-Sánchez, L., Liberatore, F., Camacho-Collados, M.: Detecting and Monitoring Hate Speech in Twitter. Sensors \textbf{19}(21), 4654--4691 (2019)
+
+\bibitem{phung}
+Phung, N. M., Mimura, M.: Evaluation of a cGAN Model and Random Seed Oversampling on Imbalanced JavaScript Datasets. Journal of Information Processing \textbf{30}, 591--600 (2022)
+
+\bibitem{sallam}
+Sallam, M.: The utility of ChatGPT as an example of large language models in healthcare education, research and practice: Systematic review on the future perspectives and potential limitations.
+\emph{medRxiv preprint}, (2023)
+
+\bibitem{shelke}
+Shelke, M.
+S., Deshmukh, P.
+R., Shandilya, V.
+K.: A review on imbalanced data handling using undersampling and oversampling technique.
+International Journal of Recent Trends in Engineering Research \textbf{3}(4), 444--449 (2017).
 
@@
 
+\bibitem{wijaya}
+Wijaya, I.
+D., Putrada, A.
+G., Oktaria, D.: Overcoming Data Imbalance Problems in Sexual Harassment Classification with SMOTE.
+International Journal on Information and Communication Technology \textbf{8}(1), 20--29 (2022).
+
 \end{thebibliography}

### 2023-04-30 paper/paper.tex
finalize first draft of paper

@@
 
+\newenvironment{nscenter}
+ {\parskip=0pt\par\nopagebreak\centering}
+ {\par\noindent\ignorespacesafterend}
+
 \begin{document}
+\makeatletter
+\renewcommand\section{\@startsection{section}{1}{\z@}%
+                       {-10\p@}%
+                       {8\p@ }%
+                       {\normalfont\large\bfseries\boldmath
+                        \rightskip=\z@ \@plus 8em\pretolerance=10000 }}
+\renewcommand\subsection{\@startsection{subsection}{2}{\z@}%
+                       {-9\p@ }%
+                       {7\p@ }%
+                       {\normalfont\normalsize\bfseries\boldmath
+                        \rightskip=\z@ \@plus 8em\pretolerance=10000 }}
+\renewcommand\subsubsection{\@startsection{subsubsection}{3}{\z@}%
+                       {-5\p@ }%
+                       {-2em }%
+                       {\normalfont\normalsize\bfseries\boldmath}}
+\makeatother
+
+\setlength{\textfloatsep}{0pt plus 0pt minus 0pt}
+\setlength{\intextsep}{0pt plus 0pt minus 0pt}
+ \setlength{\parskip}{0pt plus 0pt minus 0pt}
+
 \title{Generative LLM oversampling for the class imbalance problem in hate speech detection}
 \author{Nicolas Antonio Cloutier \and Nathalie Japkowicz\inst{1}}
+\authorrunning{Cloutier \and Japkowicz}
+\titlerunning{Generative LLM oversampling for the class imbalance problem}
 
@@
 \begin{abstract}
-
+Online hate speech has become increasingly prevalent with the rise of social media.
+As such, methods for automatically detecting and classifying hate speech have become the subject of much research.
+A common challenge in this and other domains is the class imbalance problem, where one class in a dataset is far more common than another.
+We propose the use of a generative large language model (LLM), specifically OpenAI's GPT-2, as a method of oversampling text data in order to account for this imbalance, comparing it to other resampling methods on three tasks: binary classification of tweets, classification of antisemitic tweets into types of antisemitism, and a combination of the two.
+We find that generative LLM resampling does not produce better results for binary classification than other resampling methods, but does improve performance on the other two tasks.
 \end{abstract}
@@
 In recent years, hate speech has become increasingly mainstream and common within social media sites \cite{siegel}.
-This rise has not only caused online spaces to become less hospitable, but also had several offline effects.
-On top of having severe psychological effects on the recipient \cite{siegel}, online hate speech also played a roll in disseminating extremist anti-Rohingya voices leading to violence in Myanmar \cite{green}, and has provided motivation for several perpetrators of offline violent hate crimes \cite{siegel}.
+This rise has not only caused online spaces to become less hospitable, but has also had several offline effects.
+On top of having severe psychological effects on the recipient \cite{siegel}, online hate speech also played a role in disseminating extremist anti-Rohingya voices leading to violence in Myanmar \cite{green}, and has provided motivation for several perpetrators of offline violent hate crimes \cite{siegel}.
 These events have motivated responses from numerous parties, including the social media sites themselves, that are increasingly looking to stop the spread of these messages \cite{ullmann}, and governmental bodies, that are seeking to regulate or prevent the spread of hate speech \cite{banks}.
@@
 Imbalance is a common issue in ML classification problems.
-The class imbalance problem can severely negatively affect model predictive power, with the less common class (termed the ``minority class'') is often misclassified due to its low prevalence in the dataset relative to the larger class (the ``majority class'') \cite{abdelrahman}.
+The class imbalance problem can severely negatively affect model predictive power, with the less common class (termed the ``minority class'') often being misclassified due to its low prevalence in the dataset relative to the larger class (the ``majority class'') \cite{abdelrahman}.
 This problem has affected fields as distinct as medicine, fraudulent call detection, and risk management \cite{sun}.
@@
 %P3: focus on resampling methods
-One such technique is resampling, altering the training data by adding new data points or taking existing data points away in order to improve model performance.
+One such group of techniques is resampling, altering the training data by adding new data points or taking existing data points away in order to improve model performance.
 Generally, there are two types of resampling: oversampling, which describes the process of adding synthetic data points to a dataset, and undersampling, which describes the process of removing existing data points from a dataset \cite{shelke}.
-Resampling techniques can improve the performance of ML models when trained on imbalanced data \cite{lee} \cite{khushi}, advancing the state of the field of ML in problems with unbalanced datasets.
+Resampling techniques can improve the performance of ML models when trained on imbalanced data \cite{lee} \cite{khushi}.
 
@@
 They have achieved success in their applications to primarily computer vision and tabular data \cite{hao} \cite{engelmann} \cite{bellinger} \cite{dai}.
-Applications of these generative models to NLP do occur \cite{phung}, they are generally less common than applications to other areas.
-Natural language processing (NLP) tends to use more traditional statistical methods for oversampling such as SMOTE and random oversampling \cite{wijaya} \cite{glazkova}.
+Applications of these generative models to natural language processing (NLP) do occur \cite{phung}, but they are generally less common than applications to other areas.
+NLP tends to use more traditional statistical methods for oversampling such as SMOTE and random oversampling \cite{wijaya} \cite{glazkova}.
 
 %P5: talk about LLMs and their applications, find whether there is research done with LLMs on class imbalance
-While these advances in resampling have been occuring, similar advances have been made in generative LLMs, such as the GPT series of models from OpenAI, which have the ability to generate human-like text \cite{floridi} and been applied to patent claim generation \cite{hsiang}, healthcare education \cite{sallam}.
+While these advances in resampling have been occurring, similar advances have been made in generative LLMs, such as the GPT series of models from OpenAI, which have the ability to generate human-like text \cite{floridi} and have been applied to fields such as patent claim generation \cite{hsiang} and healthcare education \cite{sallam}.
 These LLMs, despite their ability of producing authentic examples given prompts, have not been widely investigated as a method for oversampling text data.
@@
 %P1: talk about the different tasks
+Models were trained on three different tasks.
+The first was a simple binary classification task, where the models would attempt to classify the text of a tweet as either antisemitic or not antisemitic.
+The second was the 4-class type classification, where the dataset would be limited to only antisemitic samples, that would then be grouped into four classes: political, economic, religious, and racial antisemitism.
+Finally, the models were trained on a 5-class type classification task, where the dataset included both antisemitic and non-antisemitic samples, and the model would attempt to classify the samples into one of the four groups of antisemitism or classify them as not antisemitic, creating five classes total.
+Every model was trained and tested separately for all three tasks.
 
 %P2: talk about the different types of models trained
+In order to best represent a variety of ML architectures, different algorithms and methods of text representation were used.
+For algorithms, we trained classifiers utilizing the Naïve Bayes, Extreme Gradient Boosting, Decision Tree, and Support Vector Machine algorithms.
+We also used three methods of representing text: term frequency-inverse document frequency (TF-IDF), raw frequency, and Bag of Words (BoW).
+Each model was trained with each method of text representation, creating a total of twelve models that were trained for each task.
+In order to reduce the dimensionality of the dataset, representations for training and testing data were limited to words that had a frequency score of at least 0.5\%, meaning that the word had to have a frequency of at least 0.5\% of the original, full length of the text in order to be considered by the models.
+
+The dataset is imbalanced.
+The following figures show the distribution of different classes in the dataset.
+The first is for the entire dataset, and the second, containing classifications on types of antisemitism, only contains datapoints that are antisemitic.\\
+
+\begin{center}
+\includegraphics[scale=0.6]{binary\_pie.pdf}\\
+
+{\bf Figure 1.} Binary classification distributions
+\end{center}
+
+\begin{nscenter}
+\includegraphics[scale=0.6]{type\_pie.pdf}\\
+{\bf Figure 2.} Type classification distributions
+\end{nscenter}
 
 %P3: talk about the resampling methods used
+Several methods of resampling were used to reduce the impact of the class imbalance.
+One set of models was trained with no resampling, with additional sets being trained using random undersampling, random oversampling, SMOTE with Tomek Links, ADASYN, and a final set being trained on the augmented dataset generated using the LLM.
+GPT-2 was used to generate the samples because it is easily available and callable programmatically with the HuggingFace API.
+The models that were trained using the augmented dataset also used random undersampling on the augmented data.
 
 %P4: talk about the augmented dataset specifically and testing
+In regards to testing, each model was tested using 10-fold cross-validation on the original dataset.
+It was ensured that no samples generated from oversampling methods were used during testing, and additionally that, when testing the augmented model, if a sample that was used to generate more samples appeared in the testing split, the samples generated with it would not appear in the training split.
+This ensures that the models were not unfairly advantaged, and that each was tested on a large amount of genuine, unseen data.
 
 %P5: talk about evaluation (mean recall) and methods used (cochran, etc)
+For model evaluation, the main metric used was the mean of the recall scores across each of the classes the model had to classify.
+This was used in lieu of accuracy in order to account for the class imbalance in the data, but accuracy was also tracked for informative purposes.
+Once the models were evaluated, their answers to the testing samples were converted to a binary matrix with each column representing a model and each row representing a sample.
+We then used the Cochran's Q-test to test for significant difference in the models, then the Dunn test for post-hoc analysis.
+When testing the resampling methods against each other, the data were turned into another matrix with each column representing a resampling method and each row representing a model trained with that method, with the value in the cell being the mean recall of that model.
+A Friedman's $\chi^2$ test was then performed with a post-hoc Nemenyi test to analyze the results.
 
+{\parskip=0pt
 \section{Results}
-%P1: introduce, discuss difference in results in binary and 4-class tasks
+Overall, the models trained on the LLM-augmented data showed no improvement on the binary task, but they did show improvement on the 5-class classification and especially the 4-class classification tasks.
+The following table contains information for the binary classification task with each resampling method, as well as the best-performing model performance and mean model performance for that resampling method, with ``performance'' meaning the mean recall score the model achieved.
+The data used to create the below table received a Friedman $\chi^2$ test statistic of 28.19 and \emph{p}-value of $3.34 \times 10^{-5}$.
+After the table, a matrix is provided with the results of the post-hoc Nemenyi test on the resampling methods, with \emph{p}-values displayed in the array.
+A higher \emph{p}-value and lighter color means a greater degree of similarity between two models' answers and performance, while a lower \emph{p}-value and darker color means a lower degree of similarity.
+In this array, ``RU'' stands for random undersampling, ``ST'' stands for SMOTE with Tomek links, ``RO'' stands for random oversampling, ``AD'' stands for ADASYN, ``NO'' stands for no resampling method, and ``LA'' stands for LLM-augmented.}
+\begin{nscenter}
+{\bf Table 1.} Binary task resampling method performance\\
 
-%P2: make diluted 4-class table, show Dunn test array (also for binary?)
+\begin{tabular}{|l|l|l|}
+\hline
+Resampling Method & Best Score & Mean Score \\\hline
+Random Undersampling & 0.696 & 0.640 \\\hline
+SMOTE-Tomek & 0.670 & 0.612 \\\hline
+Random Oversampling & 0.686 & 0.624 \\\hline
+ADASYN & 0.660 & 0.606 \\\hline
+None & 0.667 & 0.613 \\\hline
+LLM-Augmented & 0.679 & 0.617 \\\hline
+\end{tabular}
 
-%P3: directly compare aug to randomundersampling
+\includegraphics[scale=0.6]{binary.pdf}\\
+{\bf Figure 3.} Binary Nemenyi test results
+\end{nscenter}
+In the binary task, the best performing resampling method both in mean score and best score was random undersampling, which was also found through the Nemenyi analysis to be the least similar to the other models.
+Overall, it seems as if the augmented data had no positive impact on the binary classification task, and that in this case simple and traditional statistical methods are better than generative LLM oversampling.
+The generative LLM did beat some oversampling methods (SMOTE and ADASYN), and was better than no resampling at all, but was beaten by simple random oversampling, suggesting an inappropriateness of the method to this task.
 
-%P4: talk about how many dimensions saw improvements with aug
+The results of the 5-class classification task are shown below.
+The data used to create the below table received a Friedman $\chi^2$ test statistic of 29.19 and \emph{p}-value of $2.13 \times 10^{-5}$.
+\begin{nscenter}
+{\bf Table 2.} 5-class task resampling method performance\\
+\begin{tabular}{|l|l|l|}
+\hline
+Resampling Method & Best Score & Mean Score \\\hline
+Random Undersampling & 0.300 & 0.260 \\\hline
+SMOTE-Tomek & 0.306 & 0.276 \\\hline
+Random Oversampling & 0.371 & 0.298 \\\hline
+ADASYN & 0.316 & 0.251 \\\hline
+None & 0.312 & 0.264 \\\hline
+LLM-Augmented & 0.402 & 0.347 \\\hline
+\end{tabular}
 
-%P5: summarize?
+\includegraphics[scale=0.6]{5type.pdf}\\
+{\bf Figure 4.} 5-class Nemenyi test results
+\end{nscenter}
+These 5-type results paint a better picture of the potential of generative LLM oversampling.
+This time, it is the LLM augmented models that are most different from the other models, with these models also achieving the highest best and mean scores.
+In the 5-class problem, the random undersampling that was previously the best choice ends up having the second worst mean score and the single worst best score, overall producing worse models than no resampling at all.
+Oversampling methods such as SMOTE, ADASYN, and especially random oversampling, fared better, but they were still beaten by the generative LLM.
+
+The results of the 4-class classification task are shown below.
+The data used to create the below table received a Friedman $\chi^2$ test statistic of 23.14 and \emph{p}-value of $3.16 \times 10^{-4}$.
+\begin{nscenter}
+{\bf Table 3.} 4-class task resampling method performance\\
+
+\begin{tabular}{|l|l|l|}
+\hline
+Resampling Method & Best Score & Mean Score \\\hline
+Random Undersampling & 0.426 & 0.339 \\\hline
+SMOTE-Tomek & 0.438 & 0.366 \\\hline
+Random Oversampling & 0.405 & 0.351 \\\hline
+ADASYN & 0.471 & 0.319 \\\hline
+None & 0.348 & 0.323 \\\hline
+LLM-Augmented & 0.509 &  0.433 \\\hline
+\end{tabular}
+
+\includegraphics[scale=0.6]{4type.pdf}
+
+{\bf Figure 5.} 4-class Nemenyi test results
+\end{nscenter}
+Again in the 4-class classification, the models trained on LLM-augmented data showed the best performance, and by a wider margin than in the 5-class task.
+In the 5-class task, there was a difference of 0.049 in the score of the LLM augmentation and the second place model, and that margin increases to 0.067, more than a 33\% increase in margin.
+Again, looking at the Nemenyi analysis results, the models trained with LLM-augmented data appear to be the odd men out, having the least in common with other models.
+An interesting portion of the results given is the stark contrast between the best and mean scores for the ADASYN models.
+This is due to a single model, the Naïve Bayes model trained with the Bag-of-Words representation, significantly outperforming the other ADASYN models.
+
+One initial possible explanation for the high performance of the models with LLM-augmented data relative to other models on the 4- and 5-class tasks is that the LLM augmentation provides performance improvements with only one particular class that other resampling methods did poorly on, artificially raising the mean recall even though there was only one class affected significantly.
+Looking further at the data, however, while some classes were more affected than others, overall, the models trained on LLM-augmented data show improvements in multiple classes.
+On the 4-type classification task, the following table shows the mean recall of each method on each particular class.
+\begin{nscenter}
+{\bf Table 4.} 4-class task class-wise resampling method recall\\
+
+\begin{tabular}{|l|l|l|l|l|}
+\hline
+Resampling Method & Political & Economic & Religious & Racial \\\hline
+Random Undersampling & 0.355 & 0.250 & 0.450 & 0.196  \\\hline
+SMOTE-Tomek & 0.511 & 0.135 & 0.460 & 0.246 \\\hline
+Random Oversampling & 0.482 & 0.134 & 0.448 & 0.232  \\\hline
+ADASYN & 0.568 & 0.125 & 0.219 & 0.264 \\\hline
+None & 0.622 & 0.096 & 0.193 & 0.280  \\\hline
+Non-LLM mean & 0.508 & 0.148 & 0.354 & 0.245 \\\hline
+LLM-Augmented & 0.492 & 0.230 & 0.471 & 0.405  \\\hline
+\end{tabular}
+\end{nscenter}
+The performance of the models trained on LLM-augmented data is roughly consistent with the performance of other models on political antisemitism, although it is slightly lower than the average, but it outperforms the other models in every other class.
+It is expected for a resampling method that the smallest classes be the most affected, as its purpose is to account for the imbalance of classes.
 
@@
 %P1: talk about difference between binary and 4-class tasks
+The most striking result of these trials is the difference in relative performance among the different tasks.
+Given that a class imbalance exists in all tasks, it may be expected that the generative LLM oversampling would produce similar results for each of these imbalances, but this is not the case.
+Furthermore, the difference in performance of the generative LLM-trained models to the models trained on data from other resampling methods grows from the 5-class to the 4-class task, suggesting that there is something particular to the binary classification problem that the generative LLM is not as good at accounting for as other resampling methods.
 
 %P2: suggest reasons for this (authenticity of text samples)
+One possible reason for this is the relative performance of undersampling and oversampling methods in general.
+Generally, oversampling methods tend to, at best, perform as well as undersampling methods \cite{drummond}, explaining why the only pure undersampling method scored best on the binary classification task.
+That being said, in tasks where undersampling would leave a dataset very small, this pattern may reverse.
+In the case of the 4-class classification task, one of the classes has only 8 observations, meaning, during undersampling, all other classes would also be reduced to 8 samples, making a very small dataset, that is perhaps inadequate for classification, providing one possible explanation for why oversampling outperformed undersampling on the 4-class and 5-class problems.
+As for why the generative LLM specifically outperformed other oversampling methods, it may simply be that generating text and then converting it into a vectorized form produces more authentic results than generating new samples from those vectorized forms with methods such as ADASYN and SMOTE.
+This would explain the augmented LLM's outperformance of other oversampling methods on the 4-class and 5-class tasks, but would leave unexplained why it was outperformed by random oversampling in the classification task.
 
 %P3: suggest further lines of inquiry (testing of this on images, tabular, testing more texts)
+In order to reach a better understanding of the applications of generative LLM oversampling, more research should be done on the abilities of the method to account for class imbalances in text data.
+Hopefully, future research that applies the method to datasets with differing sizes, levels of severity of imbalance, and number of classes to classify will help pinpoint the exact situations where generative LLM oversampling can be most helpful, informing future systems that may use the method.
 
 \section{Conclusions}
-%P1: talk about implications for antisemitism classification
+This paper proposes the use of generative LLMs to oversample imbalanced text datasets in order to account for the imbalance.
+Training models with different resampling methods, we find that results are unequal among tasks.
+The models trained on generative LLM data showed no improvement over other models in the binary classification of social media posts into an antisemitic/not antisemitic dichotomy, but they showed great improvement on the classification of tweets into types of antisemitism.
+Further research on imbalanced datasets with different characteristics is required to understand why the difference in relative performance was so stark and what factors most contribute to the utility of this method to accounting for the imbalance problem.
 
-%P2: talk about implications for the class imbalance problem
-
-%P3: talk about implications for text data
-
-\begin{thebibliography}{24}
+\begin{thebibliography}{25}
 \bibitem{abdelrahman}
@@
 \bibitem{bellinger}
-Bellinger, C., Japkowicz, N., Drummond, C.: Synthetic oversampling for advanced radioactive threat detection. IEEE CONFERENCE ON MACHINE LEARNING AN APPLICATIONS 2015.
+Bellinger, C., Japkowicz, N., Drummond, C.: Synthetic oversampling for advanced radioactive threat detection. IEEE CONFERENCE ON MACHINE LEARNING AND APPLICATIONS 2015.
 
@@
 Dai, W., Ng, K., Severson, K., Huan, W., Anderson, F., Stultz, C.: Generative oversampling with a contrastive variational autoencoder. IEEE INTERNATIONAL CONFERENCE ON DATA MINING 2019, pp. 101-109.
+
+\bibitem{drummond}
+Drummond, C., Holte, R. C.: C4.5, Class Imbalance, and Cost Sensitivity: Why Under-Sampling beats Over-Sampling. WORKSHOP ON LEARNING FROM IMBALANCED DATASETS 2003.
 
@@
 M., Hammed, I.
-A., Uddin, S., Luo, S., Yang, X., Consuelo Reyes, M.: A Comparitive Performance Analysis of Data Resampling Methods on Imbalance Medical Data.
+A., Uddin, S., Luo, S., Yang, X., Consuelo Reyes, M.: A Comparative Performance Analysis of Data Resampling Methods on Imbalance Medical Data.
 IEEE Access \textbf{9}, 109960--109975 (2021)

### 2023-05-03 paper/paper.tex
revise paper

@@
 \title{Generative LLM oversampling for the class imbalance problem in hate speech detection}
-\author{Nicolas Antonio Cloutier \and Nathalie Japkowicz\inst{1}}
+\author{Nicolas Antonio Cloutier\inst{1} \and Nathalie Japkowicz\inst{2}}
 \authorrunning{Cloutier \and Japkowicz}
@@
 
-\institute{1. American University, Washington, D.C., USA}
-
-%research along the way
-%4/26: outline
-%4/27: introduction
-%4/28: previous work
-%4/29: methodology
-%4/30: results
-%5/1: discussion
-%5/2: conclusions
-%5/3: abstract and finalize bibliography
+\institute{Jackson-Reed High School, Washington, D.C., USA\\\email{nicocloutier1@gmail.com} \and American University, Washington, D.C., USA\\\email{japkowicz@american.edu}}
 
@@
 A common challenge in this and other domains is the class imbalance problem, where one class in a dataset is far more common than another.
-We propose the use of a generative large language model (LLM), specifically OpenAI's GPT-2, as a method of oversampling text data in order to account for this imbalance, comparing it to other resampling methods on three tasks: binary classification of tweets, classification of antisemitic tweets into types of antisemitism, and a combination of the two.
+We propose the use of a generative large language model (LLM), specifically OpenAI's GPT-2, as a method of oversampling text data in order to account for this imbalance, comparing it to other resampling methods on three tasks: binary classification of tweets as antisemitic or not, multi-class classification of antisemitic tweets into subtypes, and a combination of the two.
 We find that generative LLM resampling does not produce better results for binary classification than other resampling methods, but does improve performance on the other two tasks.
@@
 Chandra et al.
-\cite{chandra} used a combination of image and text processing and classification algorithms to classify images and text on social media sites Twitter and Gab \cite{chandra}.
-We use their Twitter dataset to further analyze the presence of hate speech on Twitter and investigate new algorithms for classification.
+\cite{chandra} used a combination of image and text processing and classification algorithms to classify images and text on social media sites Twitter and Gab as antisemitic or not \cite{chandra}.
+We use the text component of their dataset to further analyze the presence of hate speech on Twitter and investigate new algorithms for classification.
 
@@
 Imbalanced data can negatively impact the performance of ML classification algorithms \cite{sun}, affecting numerous domains that use ML classification.
-Many ML algorithms are inadequately prepared to handle the class imbalance problem \cite{sun}, leading many to look to other solutions, including resampling methods, that in some way change the training data by adding or removing data points in order to allay the effects of the class imbalance problem \cite{japkowicz}.
+Many ML algorithms are inadequately prepared to handle the class imbalance problem \cite{sun}, leading many to look to other solutions, including resampling methods, that in some way change the training data by adding or removing samples in order to allay the effects of the class imbalance problem \cite{japkowicz}.
 
@@
 %P3: focus on resampling methods
-One such group of techniques is resampling, altering the training data by adding new data points or taking existing data points away in order to improve model performance.
-Generally, there are two types of resampling: oversampling, which describes the process of adding synthetic data points to a dataset, and undersampling, which describes the process of removing existing data points from a dataset \cite{shelke}.
+One such group of techniques is resampling, altering the training data by adding new samples or taking existing samples away in order to improve model performance.
+Generally, there are two types of resampling: oversampling, which describes the process of adding synthetic samples to a dataset, and undersampling, which describes the process of removing existing samples from a dataset \cite{shelke}.
 Resampling techniques can improve the performance of ML models when trained on imbalanced data \cite{lee} \cite{khushi}.
@@
 One method for oversampling is the use of generative models, including generative adversarial networks (GANs) and autoencoders to generate synthetic data.
-These methods seek to match the distribution of the original dataset and create synthetic datapoints in accordance with that distribution \cite{hao}, with the goal of creating authentic synthetic examples for model training.
-They have achieved success in their applications to primarily computer vision and tabular data \cite{hao} \cite{engelmann} \cite{bellinger} \cite{dai}.
+These methods seek to match the distribution of the original dataset and create synthetic samples in accordance with that distribution \cite{hao}, with the goal of creating authentic synthetic examples for model training.
+They have achieved success in their applications, but primarily in computer vision and tabular data \cite{hao} \cite{engelmann} \cite{bellinger} \cite{dai}.
 Applications of these generative models to natural language processing (NLP) do occur \cite{phung}, but they are generally less common than applications to other areas.
@@
 While these advances in resampling have been occurring, similar advances have been made in generative LLMs, such as the GPT series of models from OpenAI, which have the ability to generate human-like text \cite{floridi} and have been applied to fields such as patent claim generation \cite{hsiang} and healthcare education \cite{sallam}.
-These LLMs, despite their ability of producing authentic examples given prompts, have not been widely investigated as a method for oversampling text data.
+These LLMs have been investigated as a method for oversampling text data to account for the class imbalance problem, with relatively consistent success \cite{edwards} \cite{usuga} \cite{shaikh}, but their use is not widespread and they have not to our knowledge been applied to hate speech detection.
 
@@
 %P1: talk about the different tasks
-Models were trained on three different tasks.
+The purpose of our paper is to study the effects of various types of resampling on hate speech detection and classification tasks.
+Our models were trained on three different tasks.
 The first was a simple binary classification task, where the models would attempt to classify the text of a tweet as either antisemitic or not antisemitic.
@@
 For algorithms, we trained classifiers utilizing the Naïve Bayes, Extreme Gradient Boosting, Decision Tree, and Support Vector Machine algorithms.
-We also used three methods of representing text: term frequency-inverse document frequency (TF-IDF), raw frequency, and Bag of Words (BoW).
-Each model was trained with each method of text representation, creating a total of twelve models that were trained for each task.
+We also used three methods of representing text: term frequency-inverse document frequency (TF-IDF), raw frequency, and Bag of Words (BoW).\footnote{We initially included a BERT-based representation method, but the models trained with this method produced results no better than chance.} Each model was trained with each method of text representation, creating a total of twelve models that were trained for each task.
 In order to reduce the dimensionality of the dataset, representations for training and testing data were limited to words that had a frequency score of at least 0.5\%, meaning that the word had to have a frequency of at least 0.5\% of the original, full length of the text in order to be considered by the models.
@@
 The dataset is imbalanced.
-The following figures show the distribution of different classes in the dataset.
-The first is for the entire dataset, and the second, containing classifications on types of antisemitism, only contains datapoints that are antisemitic.\\
-
-\begin{center}
-\includegraphics[scale=0.6]{binary\_pie.pdf}\\
-
-{\bf Figure 1.} Binary classification distributions
-\end{center}
+Figure 1 shows the distribution of different classes in the dataset.
+The first and third charts are for the entire dataset, and the second, containing classifications on types of antisemitism, only contains samples that are antisemitic.\\
 
 \begin{nscenter}
-\includegraphics[scale=0.6]{type\_pie.pdf}\\
-{\bf Figure 2.} Type classification distributions
+\includegraphics[width=.3\textwidth]{binary\_pie.pdf}
+\includegraphics[width=.3\textwidth]{type\_pie.pdf}
+\includegraphics[width=.3\textwidth]{type5\_pie.pdf}
+
+{\bf Figure 1.} Classification distributions
 \end{nscenter}
@@
 Several methods of resampling were used to reduce the impact of the class imbalance.
-One set of models was trained with no resampling, with additional sets being trained using random undersampling, random oversampling, SMOTE with Tomek Links, ADASYN, and a final set being trained on the augmented dataset generated using the LLM.
-GPT-2 was used to generate the samples because it is easily available and callable programmatically with the HuggingFace API.
-The models that were trained using the augmented dataset also used random undersampling on the augmented data.
+One set of models was trained with no resampling, with additional sets being trained using random undersampling, random oversampling, SMOTE with Tomek Links, AdaSyn, and a final set being trained on the augmented dataset generated using the LLM.
+GPT-2 was used to generate the samples because it is easily available and callable programmatically with the HuggingFace API, unlike more recently released GPT models.
+The augmented dataset was created by generating 20 samples from each sample in the original dataset, creating a dataset that was larger, but had the same percentages of class imbalance.
+These models then used random undersampling on this larger dataset to account for this.
 
@@
 %P5: talk about evaluation (mean recall) and methods used (cochran, etc)
-For model evaluation, the main metric used was the mean of the recall scores across each of the classes the model had to classify.
-This was used in lieu of accuracy in order to account for the class imbalance in the data, but accuracy was also tracked for informative purposes.
+For model evaluation, the two main metrics used were the means of the recall and precision scores across each of the classes the model had to classify.
+This was used in lieu of accuracy in order to account for the class imbalance in the data.
 Once the models were evaluated, their answers to the testing samples were converted to a binary matrix with each column representing a model and each row representing a sample.
-We then used the Cochran's Q-test to test for significant difference in the models, then the Dunn test for post-hoc analysis.
-When testing the resampling methods against each other, the data were turned into another matrix with each column representing a resampling method and each row representing a model trained with that method, with the value in the cell being the mean recall of that model.
+We then used the Cochran's Q-test, an extension of McNemar's test appropriate for the case of comparing multiple algorithms on a single domain, to test for significant difference in the models, then the Dunn test for post-hoc analysis.\footnote{See \cite{boukouvalas}} When testing the resampling methods against each other, the data were turned into another matrix with each column representing a resampling method and each row representing a model trained with that method, with the value in the cell being the mean recall of that model.
 A Friedman's $\chi^2$ test was then performed with a post-hoc Nemenyi test to analyze the results.
@@
 Overall, the models trained on the LLM-augmented data showed no improvement on the binary task, but they did show improvement on the 5-class classification and especially the 4-class classification tasks.
-The following table contains information for the binary classification task with each resampling method, as well as the best-performing model performance and mean model performance for that resampling method, with ``performance'' meaning the mean recall score the model achieved.
-The data used to create the below table received a Friedman $\chi^2$ test statistic of 28.19 and \emph{p}-value of $3.34 \times 10^{-5}$.
-After the table, a matrix is provided with the results of the post-hoc Nemenyi test on the resampling methods, with \emph{p}-values displayed in the array.
+Tables 1 \& 2 contain information for the binary classification task with each resampling method, as well as the best-performing model performance and mean model performance for each resampling method, with ``performance'' meaning the mean recall or precision score the model achieved.
+The best model was the best performing model of all of the models trained with that particular resampling method, not taking into account the performance on other resampling methods.
+The data used to create Table 1 received a Friedman $\chi^2$ test statistic of 13.09 and \emph{p}-value of $2.2 \times 10^{-2}$, while Table 2 received a statistic of 19.43 and a \emph{p}-value of $1.6 \times 10^{-3}$.
+After each table, an array is provided with the results of the post-hoc Nemenyi test on the resampling methods, with \emph{p}-values displayed in the array.
 A higher \emph{p}-value and lighter color means a greater degree of similarity between two models' answers and performance, while a lower \emph{p}-value and darker color means a lower degree of similarity.
-In this array, ``RU'' stands for random undersampling, ``ST'' stands for SMOTE with Tomek links, ``RO'' stands for random oversampling, ``AD'' stands for ADASYN, ``NO'' stands for no resampling method, and ``LA'' stands for LLM-augmented.}
+In these arrays, ``RU'' stands for random undersampling, ``ST'' stands for SMOTE with Tomek links, ``RO'' stands for random oversampling, ``AD'' stands for AdaSyn, ``NO'' stands for no resampling method, and ``LA'' stands for LLM-augmented.}
 \begin{nscenter}
-{\bf Table 1.} Binary task resampling method performance\\
+{\bf Table 1.} Binary task resampling method recall\\
 
@@
 \hline
-Resampling Method & Best Score & Mean Score \\\hline
-Random Undersampling & 0.696 & 0.640 \\\hline
-SMOTE-Tomek & 0.670 & 0.612 \\\hline
-Random Oversampling & 0.686 & 0.624 \\\hline
-ADASYN & 0.660 & 0.606 \\\hline
-None & 0.667 & 0.613 \\\hline
-LLM-Augmented & 0.679 & 0.617 \\\hline
+Resampling Method & Best score (\% over none) & Mean score (\% over none) \\\hline
+None & 0.667 (0\%) & 0.613 (0\%) \\\hline
+Random Undersampling & 0.688 (3.1\%) & {\bf 0.638 (4.1\%)} \\\hline
+SMOTE-Tomek & 0.667 (0\%) & 0.614 (0.2\%) \\\hline
+Random Oversampling & {\bf 0.690 (3.4\%)} & 0.623 (1.6\%) \\\hline
+AdaSyn & 0.661 (-0.9\%) & 0.608 (-0.8\%) \\\hline
+LLM-Augmented & 0.679 (1.8\%) & 0.617 (0.7\%) \\\hline
 \end{tabular}
 
-\includegraphics[scale=0.6]{binary.pdf}\\
-{\bf Figure 3.} Binary Nemenyi test results
-\end{nscenter}
-In the binary task, the best performing resampling method both in mean score and best score was random undersampling, which was also found through the Nemenyi analysis to be the least similar to the other models.
-Overall, it seems as if the augmented data had no positive impact on the binary classification task, and that in this case simple and traditional statistical methods are better than generative LLM oversampling.
-The generative LLM did beat some oversampling methods (SMOTE and ADASYN), and was better than no resampling at all, but was beaten by simple random oversampling, suggesting an inappropriateness of the method to this task.
-
-The results of the 5-class classification task are shown below.
-The data used to create the below table received a Friedman $\chi^2$ test statistic of 29.19 and \emph{p}-value of $2.13 \times 10^{-5}$.
-\begin{nscenter}
-{\bf Table 2.} 5-class task resampling method performance\\
-\begin{tabular}{|l|l|l|}
-\hline
-Resampling Method & Best Score & Mean Score \\\hline
-Random Undersampling & 0.300 & 0.260 \\\hline
-SMOTE-Tomek & 0.306 & 0.276 \\\hline
-Random Oversampling & 0.371 & 0.298 \\\hline
-ADASYN & 0.316 & 0.251 \\\hline
-None & 0.312 & 0.264 \\\hline
-LLM-Augmented & 0.402 & 0.347 \\\hline
-\end{tabular}
-
-\includegraphics[scale=0.6]{5type.pdf}\\
-{\bf Figure 4.} 5-class Nemenyi test results
-\end{nscenter}
-These 5-type results paint a better picture of the potential of generative LLM oversampling.
-This time, it is the LLM augmented models that are most different from the other models, with these models also achieving the highest best and mean scores.
-In the 5-class problem, the random undersampling that was previously the best choice ends up having the second worst mean score and the single worst best score, overall producing worse models than no resampling at all.
-Oversampling methods such as SMOTE, ADASYN, and especially random oversampling, fared better, but they were still beaten by the generative LLM.
-
-The results of the 4-class classification task are shown below.
-The data used to create the below table received a Friedman $\chi^2$ test statistic of 23.14 and \emph{p}-value of $3.16 \times 10^{-4}$.
-\begin{nscenter}
-{\bf Table 3.} 4-class task resampling method performance\\
+\includegraphics[scale=0.6]{binary-recall.pdf}\\
+{\bf Figure 2.} Binary recall Nemenyi test results\\
+\vspace{\baselineskip}
+{\bf Table 2.} Binary task resampling method precision\\
 
@@
 \hline
-Resampling Method & Best Score & Mean Score \\\hline
-Random Undersampling & 0.426 & 0.339 \\\hline
-SMOTE-Tomek & 0.438 & 0.366 \\\hline
-Random Oversampling & 0.405 & 0.351 \\\hline
-ADASYN & 0.471 & 0.319 \\\hline
-None & 0.348 & 0.323 \\\hline
-LLM-Augmented & 0.509 &  0.433 \\\hline
+Resampling Method & Best score (\% over none) & Mean score (\% over none) \\\hline
+None & {\bf 0.734 (0\%)} & {\bf 0.644 (0\%)} \\\hline
+Random Undersampling & 0.666 (-9.3\%) & 0.621 (-3.6\%) \\\hline
+SMOTE-Tomek & 0.670 (-8.7\%) & 0.612 (-5.0\%) \\\hline
+Random Oversampling & 0.674 (-8.2\%) & 0.617 (-4.2\%) \\\hline
+AdaSyn & 0.646 (-12.0\%) & 0.600 (-6.8\%) \\\hline
+LLM-Augmented & 0.668 (-9.0\%) & 0.613 (-4.8\%) \\\hline
 \end{tabular}
 
-\includegraphics[scale=0.6]{4type.pdf}
+\includegraphics[scale=0.6]{binary-precision.pdf}\\
+{\bf Figure 3.} Binary precision Nemenyi test results
+\end{nscenter}
+In the binary task, the LLM-augmented models show no performance improvement over the other models.
+Generally, the resampling methods tend to do better than the models with no resampling on recall, but have worse precision performance, with this pattern being sustained by the models trained on LLM-augmented data.
 
-{\bf Figure 5.} 4-class Nemenyi test results
+The results of the 5-class classification task are shown below.
+The data used to create Table 3 received a Friedman $\chi^2$ test statistic of 33.24 and \emph{p}-value of $3.4 \times 10^{-6}$, while Table 4 received a statistic of 11.57 and a \emph{p}-value of $4.1 \times 10^{-2}$.
+\begin{nscenter}
+{\bf Table 3.} 5-class task resampling method recall\\
+
+\begin{tabular}{|l|l|l|}
+\hline
+Resampling Method & Best score (\% above none) & Mean score (\% above none) \\\hline
+None & 0.312 (0\%) & 0.265 (0\%) \\\hline
+Random Undersampling & 0.321 (2.9\%) & 0.265 (0\%) \\\hline
+SMOTE-Tomek & 0.314 (0.6\%) & 0.271 (2.3\%) \\\hline
+Random Oversampling & 0.369 (18.2\%) & 0.298 (12.4\%) \\\hline
+AdaSyn & 0.316 (1.3\%) & 0.251 (-5.3\%) \\\hline
+LLM-Augmented & {\bf 0.402 (28.8\%)} & {\bf 0.347 (30.9\%)} \\\hline
+\end{tabular}
+
+\includegraphics[scale=0.6]{5type-recall.pdf}\\
+{\bf Figure 4.} 5-class recall Nemenyi test results\\
+\vspace{\baselineskip}
+{\bf Table 4.} 5-class task resampling method precision\\
+
+\begin{tabular}{|l|l|l|}
+\hline
+Resampling Method & Best score (\% above none) & Mean score (\% above none) \\\hline
+None & {\bf 0.524 (0\%)} & {\bf 0.325 (0\%)} \\\hline
+Random Undersampling & 0.274 (-47.7\%) & 0.249 (-23.4\%) \\\hline
+SMOTE-Tomek & 0.295 (-43.7\%) & 0.264 (-18.8\%) \\\hline
+Random Oversampling & 0.319 (-39.1\%) & 0.271 (-16.6\%) \\\hline
+AdaSyn & 0.436 (-16.8\%) & 0.286 (-12\%) \\\hline
+LLM-Augmented & 0.302 (-42.4\%) &  0.270 (-16.9\%) \\\hline
+\end{tabular}
+
+\includegraphics[scale=0.6]{5type-precision.pdf}\\
+{\bf Figure 5.} 5-class precision Nemenyi test results
 \end{nscenter}
-Again in the 4-class classification, the models trained on LLM-augmented data showed the best performance, and by a wider margin than in the 5-class task.
-In the 5-class task, there was a difference of 0.049 in the score of the LLM augmentation and the second place model, and that margin increases to 0.067, more than a 33\% increase in margin.
-Again, looking at the Nemenyi analysis results, the models trained with LLM-augmented data appear to be the odd men out, having the least in common with other models.
-An interesting portion of the results given is the stark contrast between the best and mean scores for the ADASYN models.
-This is due to a single model, the Naïve Bayes model trained with the Bag-of-Words representation, significantly outperforming the other ADASYN models.
+The results on the 5-type task begin to show the benefits of LLM augmentation.
+The pattern of no resampling outperforming resampling on precision while the inverse happens on recall remains, but now LLM augmentation improves over no resampling by a far wider margin than any other resampling method, while retaining similar levels of drops in precision to other methods.
+Now, its gain in mean recall is larger than its loss in mean precision, which is not the case for any other resampling method in the 5-class task, and is only true for random undersampling in the binary task.
 
-One initial possible explanation for the high performance of the models with LLM-augmented data relative to other models on the 4- and 5-class tasks is that the LLM augmentation provides performance improvements with only one particular class that other resampling methods did poorly on, artificially raising the mean recall even though there was only one class affected significantly.
+The results of the 4-class classification task are shown below.
+The data used to create Table 5 received a Friedman $\chi^2$ test statistic of 21.95 and \emph{p}-value of $5.3 \times 10^{-4}$, while Table 6 received a statistic of 27.7 and a \emph{p}-value of $4.2 \times 10^{-5}$.
+\begin{nscenter}
+{\bf Table 5.} 4-class task resampling method recall\\
+
+\begin{tabular}{|l|l|l|}
+\hline
+Resampling Method & Best score (\% above none) & Mean score (\% above none) \\\hline
+None &  0.382 (0\%) & 0.328 (0\%) \\\hline
+Random Undersampling & 0.414 (8.4\%) & 0.339 (3.4\%) \\\hline
+SMOTE-Tomek & 0.421 (10.2\%) & 0.366 (11.6\%) \\\hline
+Random Oversampling & 0.400 (4.7\%) & 0.358 (9.1\%) \\\hline
+AdaSyn & 0.471 (23.3\%) & 0.322 (-1.8\%) \\\hline
+LLM-Augmented &  {\bf 0.538 (40.8\%)} & {\bf 0.433 (32.0\%)} \\\hline
+\end{tabular}
+
+\includegraphics[scale=0.6]{4type-recall.pdf}\\
+{\bf Figure 6.} 4-class recall Nemenyi test results\\
+\vspace{\baselineskip}
+{\bf Table 6.} 4-class task resampling method precision\\
+
+\begin{tabular}{|l|l|l|}
+\hline
+Resampling Method & Best score (\% above none) & Mean score (\% above none) \\\hline
+None & 0.452 (0\%) & 0.372 (0\%) \\\hline
+Random Undersampling & 0.361 (-20.1\%) & 0.323 (-13.2\%) \\\hline
+SMOTE-Tomek & 0.381 (-15.7\%) & 0.347 (-6.7\%) \\\hline
+Random Oversampling & 0.374 (-17.3\%) & 0.351 (-5.6\%) \\\hline
+AdaSyn & 0.378 (-16.4\%) & 0.344 (-7.5\%) \\\hline
+LLM-Augmented & {\bf 0.504 (11.5\%)} & {\bf 0.409 (9.9\%)} \\\hline
+\end{tabular}
+
+\includegraphics[scale=0.6]{4type-precision.pdf}\\
+{\bf Figure 5.} 4-class precision Nemenyi test results
+\end{nscenter}
+Now, the pattern of gains on recall and losses on precision continue for all methods except LLM augmentation.
+LLM-augmented models not only increase the recall by a greater margin than any other method, but also increase the precision of the model on the 4-class task.
+This makes it the only resampling method in any of the three tasks to receive a higher precision score either on the best model or on mean score than no resampling, on top of widening the margin by which it beats no resampling compared to the 5-class task.
+
+One initial possible explanation for the high performance of the models with LLM-augmented data relative to other models on the 5-class and especially 4-class tasks is that the LLM augmentation provides performance improvements with only one particular class that other resampling methods did poorly on, artificially raising the mean recall even though there was only one class affected significantly.
 Looking further at the data, however, while some classes were more affected than others, overall, the models trained on LLM-augmented data show improvements in multiple classes.
-On the 4-type classification task, the following table shows the mean recall of each method on each particular class.
+On the 4-type classification task, Tables 7 \& 8 show the mean recall and precision of each method on each particular class, respectively.
 \begin{nscenter}
-{\bf Table 4.} 4-class task class-wise resampling method recall\\
+{\bf Table 7.} 4-class task class-wise resampling method recall\\
 
@@
 Resampling Method & Political & Economic & Religious & Racial \\\hline
-Random Undersampling & 0.355 & 0.250 & 0.450 & 0.196  \\\hline
-SMOTE-Tomek & 0.511 & 0.135 & 0.460 & 0.246 \\\hline
-Random Oversampling & 0.482 & 0.134 & 0.448 & 0.232  \\\hline
-ADASYN & 0.568 & 0.125 & 0.219 & 0.264 \\\hline
-None & 0.622 & 0.096 & 0.193 & 0.280  \\\hline
-Non-LLM mean & 0.508 & 0.148 & 0.354 & 0.245 \\\hline
-LLM-Augmented & 0.492 & 0.230 & 0.471 & 0.405  \\\hline
+None & {\bf 0.618} & 0.106 & 0.203 & 0.283  \\\hline
+Random Undersampling & 0.379 & {\bf 0.240} & 0.434 &  0.200 \\\hline
+SMOTE-Tomek & 0.506 & 0.135 & 0.465 & 0.245 \\\hline
+Random Oversampling & 0.487 & 0.163 & 0.442 & 0.230 \\\hline
+AdaSyn & 0.574 & 0.135 & 0.223 & 0.258 \\\hline
+LLM-Augmented & 0.492 & 0.231 & {\bf 0.471} & {\bf 0.405} \\\hline
+\end{tabular}
+
+{\bf Table 8.} 4-class task class-wise resampling method precision\\
+
+\begin{tabular}{|l|l|l|l|l|}
+\hline
+Resampling Method & Political & Economic & Religious & Racial \\\hline
+None & 0.575 & 0.020 & 0.245 & 0.367 \\\hline
+Random Undersampling & 0.601 & 0.023 & 0.205 & 0.362 \\\hline
+SMOTE-Tomek & 0.653 & 0.018 & 0.233 & 0.378 \\\hline
+Random Oversampling & {\bf 0.661} & 0.015 & 0.227 & 0.392  \\\hline
+AdaSyn & 0.588 & 0.003 & 0.243 & 0.376 \\\hline
+LLM-Augmented & 0.570 & {\bf 0.111} & {\bf 0.360} & {\bf 0.424} \\\hline
 \end{tabular}
 \end{nscenter}
-The performance of the models trained on LLM-augmented data is roughly consistent with the performance of other models on political antisemitism, although it is slightly lower than the average, but it outperforms the other models in every other class.
-It is expected for a resampling method that the smallest classes be the most affected, as its purpose is to account for the imbalance of classes.
+In precision and recall, the model is outperformed in political antisemitism (the largest class), and in recall it is slightly outperformed by random undersampling in economic antisemitism, but it outperforms the other models in religious and racial antisemitism on recall and in all classes other than political in precision.
 
@@
 In the case of the 4-class classification task, one of the classes has only 8 observations, meaning, during undersampling, all other classes would also be reduced to 8 samples, making a very small dataset, that is perhaps inadequate for classification, providing one possible explanation for why oversampling outperformed undersampling on the 4-class and 5-class problems.
-As for why the generative LLM specifically outperformed other oversampling methods, it may simply be that generating text and then converting it into a vectorized form produces more authentic results than generating new samples from those vectorized forms with methods such as ADASYN and SMOTE.
-This would explain the augmented LLM's outperformance of other oversampling methods on the 4-class and 5-class tasks, but would leave unexplained why it was outperformed by random oversampling in the classification task.
+As for why the generative LLM specifically outperformed other oversampling methods, it may simply be that generating text and then converting it into a vectorized form produces more authentic results than generating new samples from those vectorized forms with methods such as AdaSyn and SMOTE.
+This would explain the augmented LLM's outperformance of other oversampling methods on the 4-class and 5-class tasks, but would leave unexplained why it was outperformed by random oversampling in the binary classification task.
 
+\section{Conclusions and Future Work}
 %P3: suggest further lines of inquiry (testing of this on images, tabular, testing more texts)
@@
 
-\section{Conclusions}
 This paper proposes the use of generative LLMs to oversample imbalanced text datasets in order to account for the imbalance.
@@
 
-\begin{thebibliography}{25}
+\begin{thebibliography}{27}
 \bibitem{abdelrahman}
@@
 \bibitem{bellinger}
-Bellinger, C., Japkowicz, N., Drummond, C.: Synthetic oversampling for advanced radioactive threat detection. IEEE CONFERENCE ON MACHINE LEARNING AND APPLICATIONS 2015.
+Bellinger, C., Japkowicz, N., Drummond, C.: Synthetic oversampling for advanced radioactive threat detection. IEEE Conference on Machine Learning and Applications 2015
 
@@
 Chandra, M., Pailla, D., Bhatia, H., Sanchawala, A., Gupta, M., Shrivastava, M., Kumaraguru, P.: ``Subverting the Jewtocracy'': Online Antisemitism Detection Using Multimodal Deep Learning.
-ACM WEB SCIENCE CONFERENCE 2021, pp.
-148-157.
+ACM Web Science Conference 2021, 148--157
 
 \bibitem{dai}
-Dai, W., Ng, K., Severson, K., Huan, W., Anderson, F., Stultz, C.: Generative oversampling with a contrastive variational autoencoder. IEEE INTERNATIONAL CONFERENCE ON DATA MINING 2019, pp. 101-109.
+Dai, W., Ng, K., Severson, K., Huan, W., Anderson, F., Stultz, C.: Generative oversampling with a contrastive variational autoencoder. IEEE International Conference on Data Mining 2019, 101--109
 
 \bibitem{drummond}
-Drummond, C., Holte, R. C.: C4.5, Class Imbalance, and Cost Sensitivity: Why Under-Sampling beats Over-Sampling. WORKSHOP ON LEARNING FROM IMBALANCED DATASETS 2003.
+Drummond, C., Holte, R. C.: C4.5, Class Imbalance, and Cost Sensitivity: Why Under-Sampling beats Over-Sampling. Workshop on Learning from Imbalanced Datasets 2003
+
+\bibitem{edwards}
+Edwards, A., Ushio, A., Camacho-Collados, J., De Ribaupierre, H., Preece, A.: Guiding Generative Language Models for Data Augmentation in Few-Shot Text Classification.
+\emph{arXiv preprint arXiv:2111.09064}, (2021)
 
@@
 
+\bibitem{boukouvalas}
+Japkowicz, N., Boukouvalas, Z., Shah, M.: Machine Learning Evaluation (in preparation)
+
 \bibitem{khushi}
@@
 \bibitem{martins}
-Martins, R., Gomes, M., Almeida, J. J., Novais, P., Henriques, P.: Hate speech classification in social media using emotional analysis. BRAZILIAN CONFERENCE ON INTELLIGENT SYSTEMS 2018, pp. 61-66
+Martins, R., Gomes, M., Almeida, J. J., Novais, P., Henriques, P.: Hate speech classification in social media using emotional analysis. Brazilian Conference on Intelligent Systems 2018, 61--66
 
@@
 
+\bibitem{shaikh}
+Shaikh, S., Daudpota, S.
+M., Imran, A.
+I., Kastrati, Z.: Towards Improved Classification Accuracy on Highly Imbalanced Text Dataset Using Deep Neural Language Models.
+Applied Sciences \textbf{11}(2), 869 (2021)
+
 \bibitem{shelke}
@@
 K.: A review on imbalanced data handling using undersampling and oversampling technique.
-International Journal of Recent Trends in Engineering Research \textbf{3}(4), 444--449 (2017).
+International Journal of Recent Trends in Engineering Research \textbf{3}(4), 444--449 (2017)
 
 \bibitem{siegel}
-Siegel, A.: Online Hate Speech. Social media and democracy: The state of the field, prospects for reform, 56--88 (2020).
+Siegel, A.: Online Hate Speech. Social media and democracy: The state of the field, prospects for reform, 56--88 (2020)
 
@@
 
+\bibitem{usuga}
+Usuga-Cadavid, J.
+P., Grabot, B., Lamouri, S., Fortin, A.: Artificial Data Generation with Language Models for Imbalanced Classification in Maintenance.
+Artificial data generation with language models for imbalanced classification in maintenance.
+International workshop on service orientation in holonic and multi-agent manufacturing 2021, 57--68
+
 \bibitem{wijaya}
@@
 G., Oktaria, D.: Overcoming Data Imbalance Problems in Sexual Harassment Classification with SMOTE.
-International Journal on Information and Communication Technology \textbf{8}(1), 20--29 (2022).
+International Journal on Information and Communication Technology \textbf{8}(1), 20--29 (2022)
 

### 2023-05-05 paper/paper.tex
fix mean function

@@
 While these advances in resampling have been occurring, similar advances have been made in generative LLMs, such as the GPT series of models from OpenAI, which have the ability to generate human-like text \cite{floridi} and have been applied to fields such as patent claim generation \cite{hsiang} and healthcare education \cite{sallam}.
-These LLMs have been investigated as a method for oversampling text data to account for the class imbalance problem, with relatively consistent success \cite{edwards} \cite{usuga} \cite{shaikh}, but their use is not widespread and they have not to our knowledge been applied to hate speech detection.
+These LLMs have been investigated as a method for oversampling text data to account for the class imbalance problem, with relatively consistent success over not using resampling \cite{edwards} \cite{usuga} \cite{shaikh}.
+That being said, these studies do not compare the method with other, simpler or more traditional resampling methods.
+Drummond \& Holt \cite{drummond} argue that complex methods of resampling are difficult to justify without marked improvements in performance over simpler methods.
+This provides motivation, now that the method has shown success in accounting for the class imbalance, for investigating how it compares to existing methods of resampling and under which circumstances it improves relative to those methods.
 
@@
 None & 0.667 (0\%) & 0.613 (0\%) \\\hline
-Random Undersampling & 0.688 (3.1\%) & {\bf 0.638 (4.1\%)} \\\hline
-SMOTE-Tomek & 0.667 (0\%) & 0.614 (0.2\%) \\\hline
+Random Undersampling & 0.689 (3.3\%) & {\bf 0.638 (4.1\%)} \\\hline
+SMOTE-Tomek & 0.668 (0.1\%) & 0.614 (0.2\%) \\\hline
 Random Oversampling & {\bf 0.690 (3.4\%)} & 0.623 (1.6\%) \\\hline
-AdaSyn & 0.661 (-0.9\%) & 0.608 (-0.8\%) \\\hline
+AdaSyn & 0.661 (-0.9\%) & 0.608 (0.8\%) \\\hline
 LLM-Augmented & 0.679 (1.8\%) & 0.617 (0.7\%) \\\hline
@@
 Random Undersampling & 0.321 (2.9\%) & 0.265 (0\%) \\\hline
-SMOTE-Tomek & 0.314 (0.6\%) & 0.271 (2.3\%) \\\hline
-Random Oversampling & 0.369 (18.2\%) & 0.298 (12.4\%) \\\hline
+SMOTE-Tomek & 0.289 (-7.4\%) & 0.271 (2.3\%) \\\hline
+Random Oversampling & 0.369 (18.3\%) & 0.298 (12.5\%) \\\hline
 AdaSyn & 0.316 (1.3\%) & 0.251 (-5.3\%) \\\hline
@@
 Random Oversampling & 0.319 (-39.1\%) & 0.271 (-16.6\%) \\\hline
-AdaSyn & 0.436 (-16.8\%) & 0.286 (-12\%) \\\hline
+AdaSyn & 0.436 (-16.8\%) & 0.286 (-12.0\%) \\\hline
 LLM-Augmented & 0.302 (-42.4\%) & 0.270 (-16.9\%) \\\hline
@@
 Resampling Method & Best score (\% above none) & Mean score (\% above none) \\\hline
-None &  0.382 (0\%) & 0.328 (0\%) \\\hline
-Random Undersampling & 0.414 (8.4\%) & 0.339 (3.4\%) \\\hline
-SMOTE-Tomek & 0.421 (10.2\%) & 0.366 (11.6\%) \\\hline
-Random Oversampling & 0.400 (4.7\%) & 0.358 (9.1\%) \\\hline
-AdaSyn & 0.471 (23.3\%) & 0.322 (-1.8\%) \\\hline
-LLM-Augmented &  {\bf 0.538 (40.8\%)} & {\bf 0.433 (32.0\%)} \\\hline
+None & 0.337 (0\%) & 0.328 (0\%) \\\hline
+Random Undersampling & 0.414 (22.8\%) & 0.339 (3.4\%) \\\hline
+SMOTE-Tomek & 0.420 (24.6\%) & 0.366 (11.6\%) \\\hline
+Random Oversampling & 0.400 (18.7\%) & 0.358 (9.1\%) \\\hline
+AdaSyn & 0.471 (39.8\%) & 0.322 (-1.8\%) \\\hline
+LLM-Augmented & {\bf 0.509 (51.0\%)} & {\bf 0.433 (32.0\%)} \\\hline
 \end{tabular}
@@
 None & 0.452 (0\%) & 0.372 (0\%) \\\hline
-Random Undersampling & 0.361 (-20.1\%) & 0.323 (-13.2\%) \\\hline
+Random Undersampling & 0.361 (-17.9\%) & 0.323 (-13.2\%) \\\hline
 SMOTE-Tomek & 0.381 (-15.7\%) & 0.347 (-6.7\%) \\\hline
@@
 Resampling Method & Political & Economic & Religious & Racial \\\hline
-None & 0.575 & 0.020 & 0.245 & 0.367 \\\hline
-Random Undersampling & 0.601 & 0.023 & 0.205 & 0.362 \\\hline
-SMOTE-Tomek & 0.653 & 0.018 & 0.233 & 0.378 \\\hline
+None & 0.598 & 0.029 & 0.254 & 0.390 \\\hline
+Random Undersampling & 0.601 & 0.024 & 0.205 & 0.362 \\\hline
+SMOTE-Tomek & 0.653 & 0.018 & 0.234 & 0.378 \\\hline
 Random Oversampling & {\bf 0.661} & 0.015 & 0.227 & 0.392 \\\hline
-AdaSyn & 0.588 & 0.003 & 0.243 & 0.376 \\\hline
-LLM-Augmented & 0.570 & {\bf 0.111} & {\bf 0.360} & {\bf 0.424} \\\hline
+AdaSyn & 0.602 & 0.017 & 0.243 & 0.390 \\\hline
+LLM-Augmented & 0.577 &{\bf  0.111} & {\bf 0.367} & {\bf 0.431} \\\hline
 \end{tabular}

### 2023-05-09 paper/paper.tex
update visualizations

@@
 \cite{chandra} used a combination of image and text processing and classification algorithms to classify images and text on social media sites Twitter and Gab as antisemitic or not \cite{chandra}.
-We use the text component of their dataset to further analyze the presence of hate speech on Twitter and investigate new algorithms for classification.
+In this paper, we seek a lower cost method that uses only the text component of their dataset to further analyze the presence of hate speech on Twitter and investigate new algorithms for classification.
 
@@
 First: how can automatic ML methods be improved in the domain of text classification for hate speech detection?
-Second: how can the class imbalance problem be dealt with for text data?
+Second: can generative LLMs be used to allay the class problem in text data?
 These are the questions we seek to provide answers to, with the hope that they may inform future research and hate speech detection systems.
@@
 One such group of techniques is resampling, altering the training data by adding new samples or taking existing samples away in order to improve model performance.
-Generally, there are two types of resampling: oversampling, which describes the process of adding synthetic samples to a dataset, and undersampling, which describes the process of removing existing samples from a dataset \cite{shelke}.
+Generally, there are two types of resampling: oversampling, which describes the process of adding samples to a dataset, and undersampling, which describes the process of removing existing samples from a dataset \cite{shelke}.
 Resampling techniques can improve the performance of ML models when trained on imbalanced data \cite{lee} \cite{khushi}.
@@
 One method for oversampling is the use of generative models, including generative adversarial networks (GANs) and autoencoders to generate synthetic data.
-These methods seek to match the distribution of the original dataset and create synthetic samples in accordance with that distribution \cite{hao}, with the goal of creating authentic synthetic examples for model training.
+These methods seek to match the distribution of the original dataset and create synthetic samples in accordance with that distribution \cite{hao}, with the goal of creating synthetic examples for model training that match the original distribution closely.
 They have achieved success in their applications, but primarily in computer vision and tabular data \cite{hao} \cite{engelmann} \cite{bellinger} \cite{dai}.
@@
 That being said, these studies do not compare the method with other, simpler or more traditional resampling methods.
-Drummond \& Holt \cite{drummond} argue that complex methods of resampling are difficult to justify without marked improvements in performance over simpler methods.
+Drummond \& Holte \cite{drummond} argue that complex methods of resampling are difficult to justify without marked improvements in performance over simpler methods.
 This provides motivation, now that the method has shown success in accounting for the class imbalance, for investigating how it compares to existing methods of resampling and under which circumstances it improves relative to those methods.
@@
 Our models were trained on three different tasks.
-The first was a simple binary classification task, where the models would attempt to classify the text of a tweet as either antisemitic or not antisemitic.
-The second was the 4-class type classification, where the dataset would be limited to only antisemitic samples, that would then be grouped into four classes: political, economic, religious, and racial antisemitism.
-Finally, the models were trained on a 5-class type classification task, where the dataset included both antisemitic and non-antisemitic samples, and the model would attempt to classify the samples into one of the four groups of antisemitism or classify them as not antisemitic, creating five classes total.
+The first was a simple binary classification task, where the models attempted to classify the text of a tweet as either antisemitic or not antisemitic.
+The second was the 4-class type classification, where the dataset was limited to only antisemitic samples, that was then grouped into four classes: political, economic, religious, and racial antisemitism.
+Finally, the models were trained on a 5-class type classification task, where the dataset included both antisemitic and non-antisemitic samples, and the model attempted to classify the samples into one of the four groups of antisemitism or classify them as not antisemitic, creating five classes total.
 Every model was trained and tested separately for all three tasks.
@@
 %P3: talk about the resampling methods used
-Several methods of resampling were used to reduce the impact of the class imbalance.
-One set of models was trained with no resampling, with additional sets being trained using random undersampling, random oversampling, SMOTE with Tomek Links, AdaSyn, and a final set being trained on the augmented dataset generated using the LLM.
+Several methods of resampling were used to reduce the impact of the class imbalance, which were chosen to represent a variety of resampling techniques and algorithms.
+One set of models was trained with no resampling, with additional sets being trained using random undersampling \cite{shelke}, random oversampling \cite{shelke}, SMOTE with Tomek Links\footnote{SMOTE with Tomek links was selected over only SMOTE in order to represent combined oversampling and undersampling methods, as there are two other methods already (AdaSyn and random oversampling) that are pure oversampling.} \cite{fernandez}, AdaSyn \cite{he}, and a final set being trained on the augmented dataset generated using the LLM.
 GPT-2 was used to generate the samples because it is easily available and callable programmatically with the HuggingFace API, unlike more recently released GPT models.
@@
 These models then used random undersampling on this larger dataset to account for this.
+This was done in order to avoid having to regenerate samples for each model, due to the large number of models being trained.
 
@@
 %P5: talk about evaluation (mean recall) and methods used (cochran, etc)
-For model evaluation, the two main metrics used were the means of the recall and precision scores across each of the classes the model had to classify.
+For model evaluation, the two main metrics used were the macro-averages of the recall and precision scores across each of the classes the model had to classify for each task.
 This was used in lieu of accuracy in order to account for the class imbalance in the data.
 Once the models were evaluated, their answers to the testing samples were converted to a binary matrix with each column representing a model and each row representing a sample.
-We then used the Cochran's Q-test, an extension of McNemar's test appropriate for the case of comparing multiple algorithms on a single domain, to test for significant difference in the models, then the Dunn test for post-hoc analysis.\footnote{See \cite{boukouvalas}} When testing the resampling methods against each other, the data were turned into another matrix with each column representing a resampling method and each row representing a model trained with that method, with the value in the cell being the mean recall of that model.
+We then used the Cochran's Q-test, an extension of McNemar's test appropriate for the case of comparing multiple algorithms on a single domain, to test for significant difference in the models, then the Dunn test for post-hoc analysis.\footnote{See \cite{boukouvalas}} When testing the resampling methods against each other, the data were turned into another matrix with each column representing a resampling method and each row representing a model trained with that method, with the value in the cell being the mean recall or precision of that model.
 A Friedman's $\chi^2$ test was then performed with a post-hoc Nemenyi test to analyze the results.
@@
 Overall, the models trained on the LLM-augmented data showed no improvement on the binary task, but they did show improvement on the 5-class classification and especially the 4-class classification tasks.
-Tables 1 \& 2 contain information for the binary classification task with each resampling method, as well as the best-performing model performance and mean model performance for each resampling method, with ``performance'' meaning the mean recall or precision score the model achieved.
+Tables 1 \& 2 contain information for the binary classification task with each resampling method, as well as the best-performing model performance and mean model performance for each resampling method out of the twelve combinations of each algorithm (Naïve Bayes, SVM, Decision Tree, and Extreme Gradient Boosting) and representation method (BoW, TF-IDF, and frequency), with ``performance'' meaning the mean recall or precision score the model achieved.
 The best model was the best performing model of all of the models trained with that particular resampling method, not taking into account the performance on other resampling methods.
-The data used to create Table 1 received a Friedman $\chi^2$ test statistic of 13.09 and \emph{p}-value of $2.2 \times 10^{-2}$, while Table 2 received a statistic of 19.43 and a \emph{p}-value of $1.6 \times 10^{-3}$.
+The data used to create Table 1 received a Friedman $\chi^2$ test \emph{p}-value of $2.2 \times 10^{-2}$, while Table 2 received a \emph{p}-value of $1.6 \times 10^{-3}$, allowing a rejection of the null hypothesis with $\alpha=0.05$.
 After each table, an array is provided with the results of the post-hoc Nemenyi test on the resampling methods, with \emph{p}-values displayed in the array.
-A higher \emph{p}-value and lighter color means a greater degree of similarity between two models' answers and performance, while a lower \emph{p}-value and darker color means a lower degree of similarity.
+A higher \emph{p}-value and lighter color means we cannot reject the null hypothesis, while a lower \emph{p}-value and darker color means we can.
 In these arrays, ``RU'' stands for random undersampling, ``ST'' stands for SMOTE with Tomek links, ``RO'' stands for random oversampling, ``AD'' stands for AdaSyn, ``NO'' stands for no resampling method, and ``LA'' stands for LLM-augmented.}
@@
 Resampling Method & Best score (\% over none) & Mean score (\% over none) \\\hline
-None & 0.667 (0\%) & 0.613 (0\%) \\\hline
-Random Undersampling & 0.689 (3.3\%) & {\bf 0.638 (4.1\%)} \\\hline
-SMOTE-Tomek & 0.668 (0.1\%) & 0.614 (0.2\%) \\\hline
-Random Oversampling & {\bf 0.690 (3.4\%)} & 0.623 (1.6\%) \\\hline
-AdaSyn & 0.661 (-0.9\%) & 0.608 (0.8\%) \\\hline
-LLM-Augmented & 0.679 (1.8\%) & 0.617 (0.7\%) \\\hline
+None (NO) & 0.667 (0\%) & 0.613 (0\%) \\\hline
+Random Undersampling (RU) & 0.689 (3.3\%) & {\bf 0.638 (4.1\%)} \\\hline
+SMOTE-Tomek (ST) & 0.668 (0.1\%) & 0.614 (0.2\%) \\\hline
+Random Oversampling (RO) & {\bf 0.690 (3.4\%)} & 0.623 (1.6\%) \\\hline
+AdaSyn (AD) & 0.661 (-0.9\%) & 0.608 (0.8\%) \\\hline
+LLM-Augmented (LA) & 0.679 (1.8\%) & 0.617 (0.7\%) \\\hline
 \end{tabular}
@@
 Resampling Method & Best score (\% over none) & Mean score (\% over none) \\\hline
-None & {\bf 0.734 (0\%)} & {\bf 0.644 (0\%)} \\\hline
-Random Undersampling & 0.666 (-9.3\%) & 0.621 (-3.6\%) \\\hline
-SMOTE-Tomek & 0.670 (-8.7\%) & 0.612 (-5.0\%) \\\hline
-Random Oversampling & 0.674 (-8.2\%) & 0.617 (-4.2\%) \\\hline
-AdaSyn & 0.646 (-12.0\%) & 0.600 (-6.8\%) \\\hline
-LLM-Augmented & 0.668 (-9.0\%) & 0.613 (-4.8\%) \\\hline
+None (NO) & {\bf 0.734 (0\%)} & {\bf 0.644 (0\%)} \\\hline
+Random Undersampling (RU) & 0.666 (-9.3\%) & 0.621 (-3.6\%) \\\hline
+SMOTE-Tomek (ST) & 0.670 (-8.7\%) & 0.612 (-5.0\%) \\\hline
+Random Oversampling (RO) & 0.674 (-8.2\%) & 0.617 (-4.2\%) \\\hline
+AdaSyn (AD) & 0.646 (-12.0\%) & 0.600 (-6.8\%) \\\hline
+LLM-Augmented (LA) & 0.668 (-9.0\%) & 0.613 (-4.8\%) \\\hline
 \end{tabular}
@@
 The results of the 5-class classification task are shown below.
-The data used to create Table 3 received a Friedman $\chi^2$ test statistic of 33.24 and \emph{p}-value of $3.4 \times 10^{-6}$, while Table 4 received a statistic of 11.57 and a \emph{p}-value of $4.1 \times 10^{-2}$.
+The data used to create Table 3 received a Friedman $\chi^2$ test \emph{p}-value of $3.4 \times 10^{-6}$, while Table 4 received a \emph{p}-value of $4.1 \times 10^{-2}$.
 \begin{nscenter}
@@
 Resampling Method & Best score (\% above none) & Mean score (\% above none) \\\hline
-None & 0.312 (0\%) & 0.265 (0\%) \\\hline
-Random Undersampling & 0.321 (2.9\%) & 0.265 (0\%) \\\hline
-SMOTE-Tomek & 0.289 (-7.4\%) & 0.271 (2.3\%) \\\hline
-Random Oversampling & 0.369 (18.3\%) & 0.298 (12.5\%) \\\hline
-AdaSyn & 0.316 (1.3\%) & 0.251 (-5.3\%) \\\hline
-LLM-Augmented & {\bf 0.402 (28.8\%)} & {\bf 0.347 (30.9\%)} \\\hline
+None (NO) & 0.312 (0\%) & 0.265 (0\%) \\\hline
+Random Undersampling (RU) & 0.321 (2.9\%) & 0.265 (0\%) \\\hline
+SMOTE-Tomek (ST) & 0.289 (-7.4\%) & 0.271 (2.3\%) \\\hline
+Random Oversampling (RO) & 0.369 (18.3\%) & 0.298 (12.5\%) \\\hline
+AdaSyn (AD) & 0.316 (1.3\%) & 0.251 (-5.3\%) \\\hline
+LLM-Augmented (LA) & {\bf 0.402 (28.8\%)} & {\bf 0.347 (30.9\%)} \\\hline
 \end{tabular}
@@
 Resampling Method & Best score (\% above none) & Mean score (\% above none) \\\hline
-None & {\bf 0.524 (0\%)} & {\bf 0.325 (0\%)} \\\hline
-Random Undersampling & 0.274 (-47.7\%) & 0.249 (-23.4\%) \\\hline
-SMOTE-Tomek & 0.295 (-43.7\%) & 0.264 (-18.8\%) \\\hline
-Random Oversampling & 0.319 (-39.1\%) & 0.271 (-16.6\%) \\\hline
-AdaSyn & 0.436 (-16.8\%) & 0.286 (-12.0\%) \\\hline
-LLM-Augmented & 0.302 (-42.4\%) & 0.270 (-16.9\%) \\\hline
+None (NO) & {\bf 0.524 (0\%)} & {\bf 0.325 (0\%)} \\\hline
+Random Undersampling (RU) & 0.274 (-47.7\%) & 0.249 (-23.4\%) \\\hline
+SMOTE-Tomek (ST) & 0.295 (-43.7\%) & 0.264 (-18.8\%) \\\hline
+Random Oversampling (RO) & 0.319 (-39.1\%) & 0.271 (-16.6\%) \\\hline
+AdaSyn (AD) & 0.436 (-16.8\%) & 0.286 (-12.0\%) \\\hline
+LLM-Augmented (LA) & 0.302 (-42.4\%) & 0.270 (-16.9\%) \\\hline
 \end{tabular}
@@
 The results of the 4-class classification task are shown below.
-The data used to create Table 5 received a Friedman $\chi^2$ test statistic of 21.95 and \emph{p}-value of $5.3 \times 10^{-4}$, while Table 6 received a statistic of 27.7 and a \emph{p}-value of $4.2 \times 10^{-5}$.
+The data used to create Table 5 received a Friedman $\chi^2$ test \emph{p}-value of $5.3 \times 10^{-4}$, while Table 6 received a \emph{p}-value of $4.2 \times 10^{-5}$.
 \begin{nscenter}
@@
 Resampling Method & Best score (\% above none) & Mean score (\% above none) \\\hline
-None & 0.337 (0\%) & 0.328 (0\%) \\\hline
-Random Undersampling & 0.414 (22.8\%) & 0.339 (3.4\%) \\\hline
-SMOTE-Tomek & 0.420 (24.6\%) & 0.366 (11.6\%) \\\hline
-Random Oversampling & 0.400 (18.7\%) & 0.358 (9.1\%) \\\hline
-AdaSyn & 0.471 (39.8\%) & 0.322 (-1.8\%) \\\hline
-LLM-Augmented & {\bf 0.509 (51.0\%)} & {\bf 0.433 (32.0\%)} \\\hline
+None (NO) & 0.337 (0\%) & 0.328 (0\%) \\\hline
+Random Undersampling (RU) & 0.414 (22.8\%) & 0.339 (3.4\%) \\\hline
+SMOTE-Tomek (ST) & 0.420 (24.6\%) & 0.366 (11.6\%) \\\hline
+Random Oversampling (RO) & 0.400 (18.7\%) & 0.358 (9.1\%) \\\hline
+AdaSyn (AD) & 0.471 (39.8\%) & 0.322 (-1.8\%) \\\hline
+LLM-Augmented (LA) & {\bf 0.509 (51.0\%)} & {\bf 0.433 (32.0\%)} \\\hline
 \end{tabular}
@@
 Resampling Method & Best score (\% above none) & Mean score (\% above none) \\\hline
-None & 0.452 (0\%) & 0.372 (0\%) \\\hline
-Random Undersampling & 0.361 (-17.9\%) & 0.323 (-13.2\%) \\\hline
-SMOTE-Tomek & 0.381 (-15.7\%) & 0.347 (-6.7\%) \\\hline
-Random Oversampling & 0.374 (-17.3\%) & 0.351 (-5.6\%) \\\hline
-AdaSyn & 0.378 (-16.4\%) & 0.344 (-7.5\%) \\\hline
-LLM-Augmented & {\bf 0.504 (11.5\%)} & {\bf 0.409 (9.9\%)} \\\hline
+None (NO) & 0.452 (0\%) & 0.372 (0\%) \\\hline
+Random Undersampling (RU) & 0.361 (-17.9\%) & 0.323 (-13.2\%) \\\hline
+SMOTE-Tomek (ST) & 0.381 (-15.7\%) & 0.347 (-6.7\%) \\\hline
+Random Oversampling (RO) & 0.374 (-17.3\%) & 0.351 (-5.6\%) \\\hline
+AdaSyn (AD) & 0.378 (-16.4\%) & 0.344 (-7.5\%) \\\hline
+LLM-Augmented (LA) & {\bf 0.504 (11.5\%)} & {\bf 0.409 (9.9\%)} \\\hline
 \end{tabular}
@@
 Resampling Method & Political & Economic & Religious & Racial \\\hline
-None & {\bf 0.618} & 0.106 & 0.203 & 0.283 \\\hline
-Random Undersampling & 0.379 & {\bf 0.240} & 0.434 & 0.200  \\\hline
-SMOTE-Tomek & 0.506 & 0.135 & 0.465 & 0.245 \\\hline
-Random Oversampling & 0.487 & 0.163 & 0.442 & 0.230 \\\hline
-AdaSyn & 0.574 & 0.135 & 0.223 & 0.258 \\\hline
-LLM-Augmented & 0.492 & 0.231 & {\bf 0.471} & {\bf 0.405} \\\hline
+None (NO) & {\bf 0.618} & 0.106 & 0.203 & 0.283 \\\hline
+Random Undersampling (RU) & 0.379 & {\bf 0.240} & 0.434 & 0.200  \\\hline
+SMOTE-Tomek (ST) & 0.506 & 0.135 & 0.465 & 0.245 \\\hline
+Random Oversampling (RO) & 0.487 & 0.163 & 0.442 & 0.230 \\\hline
+AdaSyn (AD) & 0.574 & 0.135 & 0.223 & 0.258 \\\hline
+LLM-Augmented (LA) & 0.492 & 0.231 & {\bf 0.471} & {\bf 0.405} \\\hline
 \end{tabular}
@@
 Resampling Method & Political & Economic & Religious & Racial \\\hline
-None & 0.598 & 0.029 & 0.254 & 0.390 \\\hline
-Random Undersampling & 0.601 & 0.024 & 0.205 & 0.362 \\\hline
-SMOTE-Tomek & 0.653 & 0.018 & 0.234 & 0.378 \\\hline
-Random Oversampling & {\bf 0.661} & 0.015 & 0.227 & 0.392 \\\hline
-AdaSyn & 0.602 & 0.017 & 0.243 & 0.390 \\\hline
-LLM-Augmented & 0.577 &{\bf  0.111} & {\bf 0.367} & {\bf 0.431} \\\hline
+None (NO) & 0.598 & 0.029 & 0.254 & 0.390 \\\hline
+Random Undersampling (RU) & 0.601 & 0.024 & 0.205 & 0.362 \\\hline
+SMOTE-Tomek (ST) & 0.653 & 0.018 & 0.234 & 0.378 \\\hline
+Random Oversampling (RO) & {\bf 0.661} & 0.015 & 0.227 & 0.392 \\\hline
+AdaSyn (AD) & 0.602 & 0.017 & 0.243 & 0.390 \\\hline
+LLM-Augmented (LA) & 0.577 &{\bf  0.111} & {\bf 0.367} & {\bf 0.431} \\\hline
 \end{tabular}
@@
 %P2: suggest reasons for this (authenticity of text samples)
-One possible reason for this is the relative performance of undersampling and oversampling methods in general.
-Generally, oversampling methods tend to, at best, perform as well as undersampling methods \cite{drummond}, explaining why the only pure undersampling method scored best on the binary classification task.
-That being said, in tasks where undersampling would leave a dataset very small, this pattern may reverse.
+One possible explanation could be the size of the classes.
 In the case of the 4-class classification task, one of the classes has only 8 observations, meaning, during undersampling, all other classes would also be reduced to 8 samples, making a very small dataset, that is perhaps inadequate for classification, providing one possible explanation for why oversampling outperformed undersampling on the 4-class and 5-class problems.
@@
 
-\begin{thebibliography}{27}
+\begin{thebibliography}{31}
 \bibitem{abdelrahman}
@@
 
+\bibitem{fernandez}
+Fernández, A., García, S., Herrera, F., Chawla, N.
+V.: SMOTE for Learning from Imbalanced Data: Progress and Challenges, Marking the 15-year Anniversary.
+Journal of Artificial Intelligence Research \textbf{61}, 863--905 (2018)
+
 \bibitem{floridi}
@@
 Hao, J., Wang, C., Zhang, H., Yang, G.: Annealing genetic GAN for minority oversampling. \emph{arXiv preprint}, (2020)
+
+\bibitem{he}
+He, H., Bai, Y., Garcia, E. A., Li, S.: ADASYN: Adaptive Synthetic Approach for Imbalanced Learning. IEEE International Joint Conference on Neural Networks 2008, 1322--1328
 
@@
 P., Grabot, B., Lamouri, S., Fortin, A.: Artificial Data Generation with Language Models for Imbalanced Classification in Maintenance.
-Artificial data generation with language models for imbalanced classification in maintenance.
-International workshop on service orientation in holonic and multi-agent manufacturing 2021, 57--68
+International Workshop on Service Orientation in Holonic and Multi-Agent Manufacturing 2021, 57--68
 

### 2023-05-13 paper/paper.tex
rewrite discussion and conclusion sections

@@
 \cite{chandra} used a combination of image and text processing and classification algorithms to classify images and text on social media sites Twitter and Gab as antisemitic or not \cite{chandra}.
-In this paper, we seek a lower cost method that uses only the text component of their dataset to further analyze the presence of hate speech on Twitter and investigate new algorithms for classification.
+In this paper, we seek a lower cost method that uses only the text component of their dataset to further analyze the presence of hate speech on Twitter and investigate new algorithms for classification.\footnote{Our dataset contains less samples than that of \cite{chandra} because Twitter removed some of the hateful posts before we collected the data.}
 
@@
 %P4: Wrap up the research question and context
-With this in mind, there are two crucial research questions this paper seeks to answer.
-First: how can automatic ML methods be improved in the domain of text classification for hate speech detection?
-Second: can generative LLMs be used to allay the class problem in text data?
-These are the questions we seek to provide answers to, with the hope that they may inform future research and hate speech detection systems.
+With this in mind, along with the recent advances in generative LLMs, we ask whether such methods can allay the class imbalance problem in text data better than traditional approaches, and to what extent that may improve hate speech detection on social media.
 
@@
 \section{Discussion}
-%P1: talk about difference between binary and 4-class tasks
-The most striking result of these trials is the difference in relative performance among the different tasks.
-Given that a class imbalance exists in all tasks, it may be expected that the generative LLM oversampling would produce similar results for each of these imbalances, but this is not the case.
-Furthermore, the difference in performance of the generative LLM-trained models to the models trained on data from other resampling methods grows from the 5-class to the 4-class task, suggesting that there is something particular to the binary classification problem that the generative LLM is not as good at accounting for as other resampling methods.
+The most striking result is the difference in relative performance of the resampling methods between the three tasks.
+The binary task sees small improvements in recall with overall larger losses in precision, the 5-class task sees large losses in precision with large gains in recall, particularly by the LLM, and the 4-class task sees large gains in recall and precision for the LLM, with other methods having smaller recall gains and performing worse than no resampling on precision.
+This can primarily be explained through two factors: the degree of imbalance of the data and the degree of appropriateness of an oversampling technique to the task at hand.
 
-%P2: suggest reasons for this (authenticity of text samples)
-One possible explanation could be the size of the classes.
-In the case of the 4-class classification task, one of the classes has only 8 observations, meaning, during undersampling, all other classes would also be reduced to 8 samples, making a very small dataset, that is perhaps inadequate for classification, providing one possible explanation for why oversampling outperformed undersampling on the 4-class and 5-class problems.
-As for why the generative LLM specifically outperformed other oversampling methods, it may simply be that generating text and then converting it into a vectorized form produces more authentic results than generating new samples from those vectorized forms with methods such as AdaSyn and SMOTE.
-This would explain the augmented LLM's outperformance of other oversampling methods on the 4-class and 5-class tasks, but would leave unexplained why it was outperformed by random oversampling in the binary classification task.
+First, the degree of imbalance can affect the relative performance of resampling methods.
+The binary task saw the least benefit from resampling methods, as its losses in precision generally outweighed its gains in recall.
+This can be explained by the fact that the binary task had relatively large amounts of data for both classes, making resampling less useful compared to other tasks, hence why the observed improvement in recall is less than that of other tasks.
+For the 5-class task, on the other hand, the classes are imbalanced to a very high degree, with sample sizes ranging from 8 to 877.
+In order to make these classes equal in size, drastic measures must be taken, either by removing large portions of the dataset or adding a quantity of synthetic or duplicated samples that outnumber the original size of that class within the dataset.
+This can, in the case of undersampling, remove critical pieces of information that could help the model \cite{marques} \cite{kraiem}, and, in the case of oversampling, introduce excessive amounts of synthetic data or make the dataset prone to overfitting \cite{marques} \cite{kraiem}.
+This could explain the large decreases in precision in this task.
+On the 4-class task, the imbalance is far more even, with sample sizes ranging from 8 to 170, certainly still a large difference, but a far cry from the 5-class task.
+This could explain why the losses in precision for the 4-class task are not as high as those in the 5-class task.
+
+This factor, however, does not explain all of the difference.
+Why, for example, do the models trained on LLM-generated data outperform other models by such a wide margin in the 4-class task, not only in recall but also precision?
+Here, it is necessary to look at the methods themselves and the algorithms they use.
+The generative LLM, being trained on large amounts of text data, has likely been trained partially on samples similar in topic, purpose, or register to the ones present in the dataset.
+Its familiarity with this type of text allows it to generate more appropriate synthetic samples than those generated purely geometrically by such algorithms as SMOTE and AdaSyn, while avoiding the overfitting that can plague random oversampling as a result of simply duplicating data \cite{marques} \cite{kraiem}.
+
+These two factors, when considered together, can explain a large portion of the observed differences in performance of the resampling methods across the three tasks.
+The binary case saw very little improvement from any method because the degree of imbalance was not very high, and there was ample data present from both classes.
+The 5-class task saw the best performance from the models trained on LLM-generated data because this method is the best-suited to the task, but the very large class imbalance negatively affected precision to a high degree.
+The 4-class task also saw the best performance from these models for the same reason, but the lower degree of class imbalance made the losses in precision lower, making all methods, but especially the generative LLM due to its high appropriateness to the task, perform better on average precision on this task than the previous one.
 
 \section{Conclusions and Future Work}
-%P3: suggest further lines of inquiry (testing of this on images, tabular, testing more texts)
-In order to reach a better understanding of the applications of generative LLM oversampling, more research should be done on the abilities of the method to account for class imbalances in text data.
-Hopefully, future research that applies the method to datasets with differing sizes, levels of severity of imbalance, and number of classes to classify will help pinpoint the exact situations where generative LLM oversampling can be most helpful, informing future systems that may use the method.
+In this paper, we propose a method of oversampling for text data involving the use of generative LLMs to generate text samples and account for the class imbalance problem.
+We apply this method to an imbalanced antisemitism classification dataset on three tasks, one binary classification and two multiclass classification.
+We find that this method does not provide benefits over other methods on the binary classification task, but provides improvements over other methods in the 5-class and especially 4-class tasks.
+Our findings suggest that generative LLM oversampling, like other resampling methods, does not provide large benefits when there is plentiful data for both classes, and will not perform as well when there is a very large degree of imbalance, but has merit over other methods in certain tasks.
+We hypothesize that this is because the generative LLMs, by basing the generation on existing text knowledge rather than geometry like other synthetic oversampling methods, are more appropriate to the task of oversampling text data.
 
-This paper proposes the use of generative LLMs to oversample imbalanced text datasets in order to account for the imbalance.
-Training models with different resampling methods, we find that results are unequal among tasks.
-The models trained on generative LLM data showed no improvement over other models in the binary classification of social media posts into an antisemitic/not antisemitic dichotomy, but they showed great improvement on the classification of tweets into types of antisemitism.
-Further research on imbalanced datasets with different characteristics is required to understand why the difference in relative performance was so stark and what factors most contribute to the utility of this method to accounting for the imbalance problem.
+For future work in the area, we believe it is important to verify the results and conclusions we have reached.
+This could be achieved through testing the methods on a wide variety of tasks, with different types of text data (different registers, topics, etc.), different numbers of classes, and different degrees of imbalance.
+These studies could help test our analyses of the results and pinpoint the situations where generative LLM resampling is most appropriate.
 
-\begin{thebibliography}{31}
+\begin{thebibliography}{33}
 \bibitem{abdelrahman}
@@
 
+\bibitem{kraiem}
+Kraiem, M.
+S., Sánchez-Hernández F., Moreno-García M.
+N.: Selecting the suitable resampling strategy for imbalanced data classification regarding dataset properties.
+an approach based on association models.
+Applied Sciences \textbf{11}(18), 8546 (2021)
+
 \bibitem{hsiang} %the main author is also named lee, this is the second author
@@
 Lee, P. H.: Resampling methods improve the predictive power of modeling in class-imbalanced datasets. International Journal of Environmental Research and Public Health \textbf{11}(9), 9776--9789.
+
+\bibitem{marques}
+Marqués, A.
+I., García, V., Salvador Sánchez, J.: On the suitability of resampling techniques for the class imbalance problem for credit scoring.
+Journal of the Operational Research Society \textbf{64}(7), 1060--1070 (2013)
 

### 2023-05-16 paper/paper.tex
change metho and results

@@
 These methods use Machine Learning (ML) to automatically detect and classify hateful speech, removing some of the work done by moderators, whose primary job is to respond reactively to user reports of hate speech \cite{ullmann}.
-Chandra et al.
 \cite{chandra} used a combination of image and text processing and classification algorithms to classify images and text on social media sites Twitter and Gab as antisemitic or not \cite{chandra}.
-In this paper, we seek a lower cost method that uses only the text component of their dataset to further analyze the presence of hate speech on Twitter and investigate new algorithms for classification.\footnote{Our dataset contains less samples than that of \cite{chandra} because Twitter removed some of the hateful posts before we collected the data.}
+In this paper, we seek a lower cost method that uses only the text component of their dataset to further analyze the presence of hate speech on Twitter and investigate new algorithms for classification.\footnote{Our dataset contains fewer samples than that of \cite{chandra} because Twitter removed some of the hateful posts before we collected the data.}
 
@@
 Antisemitism detection, as well as hate speech detection generally, has been the subject of much research.
-Martins et al.
-\cite{martins} used ML models to analyze and classify antisemitism in posts on social media sites Gab and 4chan, and González-Pizarro and Zanettou \cite{gonzalez} used large language models (LLMs) to do the same.
-Chandra et al.
+\cite{martins} used ML models to analyze and classify antisemitism in posts on social media sites Gab and 4chan, and \cite{gonzalez} used large language models (LLMs) to do the same.
 \cite{chandra} introduced a new dataset of labeled antisemitic posts, including text and images.
@@
 The class imbalance problem can severely negatively affect model predictive power, with the less common class (termed the ``minority class'') often being misclassified due to its low prevalence in the dataset relative to the larger class (the ``majority class'') \cite{abdelrahman}.
+In multiclass classification, the classes with fewer samples will similarly perform worse to those with more samples.
 This problem has affected fields as distinct as medicine, fraudulent call detection, and risk management \cite{sun}.
@@
 That being said, these studies do not compare the method with other, simpler or more traditional resampling methods.
-Drummond \& Holte \cite{drummond} argue that complex methods of resampling are difficult to justify without marked improvements in performance over simpler methods.
+\cite{drummond} argues that complex methods of resampling are difficult to justify without marked improvements in performance over simpler methods.
 This provides motivation, now that the method has shown success in accounting for the class imbalance, for investigating how it compares to existing methods of resampling and under which circumstances it improves relative to those methods.
@@
 Figure 1 shows the distribution of different classes in the dataset.
-The first and third charts are for the entire dataset, and the second, containing classifications on types of antisemitism, only contains samples that are antisemitic.\\
+The first and second charts are for the entire dataset, and the third, containing classifications on types of antisemitism, only contains samples that are antisemitic.\\
 
@@
 \includegraphics[width=.3\textwidth]{binary\_pie.pdf}
+\includegraphics[width=.3\textwidth]{type5\_pie.pdf}
 \includegraphics[width=.3\textwidth]{type\_pie.pdf}
-\includegraphics[width=.3\textwidth]{type5\_pie.pdf}
 
@@
 %P5: talk about evaluation (mean recall) and methods used (cochran, etc)
-For model evaluation, the two main metrics used were the macro-averages of the recall and precision scores across each of the classes the model had to classify for each task.
-This was used in lieu of accuracy in order to account for the class imbalance in the data.
+For model evaluation, the three main metrics used were the macro-averages of the recall, precision, and $F_1$ scores across each of the classes the model had to classify for each task.
+These were used in lieu of accuracy in order to account for the class imbalance in the data.
 Once the models were evaluated, their answers to the testing samples were converted to a binary matrix with each column representing a model and each row representing a sample.
-We then used the Cochran's Q-test, an extension of McNemar's test appropriate for the case of comparing multiple algorithms on a single domain, to test for significant difference in the models, then the Dunn test for post-hoc analysis.\footnote{See \cite{boukouvalas}} When testing the resampling methods against each other, the data were turned into another matrix with each column representing a resampling method and each row representing a model trained with that method, with the value in the cell being the mean recall or precision of that model.
+We then used the Cochran's Q-test, an extension of McNemar's test appropriate for the case of comparing multiple algorithms on a single domain, to test for significant difference in the models, then the Dunn test for post-hoc analysis.\footnote{See \cite{boukouvalas}} When testing the resampling methods against each other, the data were turned into another matrix with each column representing a resampling method and each row representing a model trained with that method, with the value in the cell being the mean recall, precision, or $F_1$ score of that model.
 A Friedman's $\chi^2$ test was then performed with a post-hoc Nemenyi test to analyze the results.
@@
 Overall, the models trained on the LLM-augmented data showed no improvement on the binary task, but they did show improvement on the 5-class classification and especially the 4-class classification tasks.
-Tables 1 \& 2 contain information for the binary classification task with each resampling method, as well as the best-performing model performance and mean model performance for each resampling method out of the twelve combinations of each algorithm (Naïve Bayes, SVM, Decision Tree, and Extreme Gradient Boosting) and representation method (BoW, TF-IDF, and frequency), with ``performance'' meaning the mean recall or precision score the model achieved.
-The best model was the best performing model of all of the models trained with that particular resampling method, not taking into account the performance on other resampling methods.
-The data used to create Table 1 received a Friedman $\chi^2$ test \emph{p}-value of $2.2 \times 10^{-2}$, while Table 2 received a \emph{p}-value of $1.6 \times 10^{-3}$, allowing a rejection of the null hypothesis with $\alpha=0.05$.
-After each table, an array is provided with the results of the post-hoc Nemenyi test on the resampling methods, with \emph{p}-values displayed in the array.
+Tables 1 contains information for the binary classification task with each resampling method, as well as the mean model performance for each resampling method out of the twelve combinations of each algorithm (Naïve Bayes, SVM, Decision Tree, and Extreme Gradient Boosting) and representation method (BoW, TF-IDF, and frequency), with ``performance'' meaning the mean recall, precision, or $F_1$ score the model achieved.
+The data used to create the $F_1$ column of Table 1 received a Friedman $\chi^2$ test \emph{p}-value of $2.2 \times 10^{-2}$, allowing a rejection of $H_0$ that each method comes from the same distribution with $\alpha=0.05$.
+After each table, an array is provided with the results of the post-hoc Nemenyi test on the resampling methods for the $F_1$ score, with \emph{p}-values displayed in the array.
 A higher \emph{p}-value and lighter color means we cannot reject the null hypothesis, while a lower \emph{p}-value and darker color means we can.
@@
 \begin{nscenter}
-{\bf Table 1.} Binary task resampling method recall\\
+{\bf Table 1.} Binary task resampling method performance\\
 
-\begin{tabular}{|l|l|l|}
+\begin{tabular}{|l|l|l|l|}
 \hline
-Resampling Method & Best score (\% over none) & Mean score (\% over none) \\\hline
-None (NO) & 0.667 (0\%) & 0.613 (0\%) \\\hline
-Random Undersampling (RU) & 0.689 (3.3\%) & {\bf 0.638 (4.1\%)} \\\hline
-SMOTE-Tomek (ST) & 0.668 (0.1\%) & 0.614 (0.2\%) \\\hline
-Random Oversampling (RO) & {\bf 0.690 (3.4\%)} & 0.623 (1.6\%) \\\hline
-AdaSyn (AD) & 0.661 (-0.9\%) & 0.608 (0.8\%) \\\hline
-LLM-Augmented (LA) & 0.679 (1.8\%) & 0.617 (0.7\%) \\\hline
+Resampling Method & Recall (\% over none) & Precision (\% over none) & $F_1$ (\% over none) \\\hline
+ None (NO) & 0.613 (0\%) & {\bf 0.644 (0\%)} & 0.627 (0\%) \\\hline
+ RandomUnder (RU) & {\bf 0.638 (4.1\%)} & 0.621 (-3.6\%) & {\bf 0.630 (0.5\%)} \\\hline
+ SMOTE-Tomek (ST) & 0.614 (0.1\%) & 0.612 (-5.0\%) & 0.613 (-2.2\%) \\\hline
+ Random Over (RO) & 0.623 (1.6\%) & 0.617 (-4.2\%) & 0.620 (-1.1\%) \\\hline
+ AdaSyn (AD) & 0.608 (-0.8\%) & 0.600 (-6.8\%) & 0.604 (-3.7\%) \\\hline
+ LLM-Augmented (LA) & 0.617 (0.6\%) & 0.613 (-4.8\%) & 0.614 (-2.1\%) \\\hline
 \end{tabular}
 
-\includegraphics[scale=0.6]{binary-recall.pdf}\\
-{\bf Figure 2.} Binary recall Nemenyi test results\\
-\vspace{\baselineskip}
-{\bf Table 2.} Binary task resampling method precision\\
-
-\begin{tabular}{|l|l|l|}
-\hline
-Resampling Method & Best score (\% over none) & Mean score (\% over none) \\\hline
-None (NO) & {\bf 0.734 (0\%)} & {\bf 0.644 (0\%)} \\\hline
-Random Undersampling (RU) & 0.666 (-9.3\%) & 0.621 (-3.6\%) \\\hline
-SMOTE-Tomek (ST) & 0.670 (-8.7\%) & 0.612 (-5.0\%) \\\hline
-Random Oversampling (RO) & 0.674 (-8.2\%) & 0.617 (-4.2\%) \\\hline
-AdaSyn (AD) & 0.646 (-12.0\%) & 0.600 (-6.8\%) \\\hline
-LLM-Augmented (LA) & 0.668 (-9.0\%) & 0.613 (-4.8\%) \\\hline
-\end{tabular}
-
-\includegraphics[scale=0.6]{binary-precision.pdf}\\
-{\bf Figure 3.} Binary precision Nemenyi test results
+\includegraphics[scale=0.6]{binary-f1.pdf}\\
+{\bf Figure 2.} Binary $F_1$ Nemenyi test results\\
 \end{nscenter}
@@
 
-The results of the 5-class classification task are shown below.
-The data used to create Table 3 received a Friedman $\chi^2$ test \emph{p}-value of $3.4 \times 10^{-6}$, while Table 4 received a \emph{p}-value of $4.1 \times 10^{-2}$.
+The results of the 5-class classification task are shown below. The data used to create the $F_1$ column of Table 2 received a Friedman $\chi^2$ test \emph{p}-value of $1.6 \times 10^{-3}$.
 \begin{nscenter}
-{\bf Table 3.} 5-class task resampling method recall\\
+{\bf Table 2.} 5-class task resampling method performance\\
 
-\begin{tabular}{|l|l|l|}
+\begin{tabular}{|l|l|l|l|}
 \hline
-Resampling Method & Best score (\% above none) & Mean score (\% above none) \\\hline
-None (NO) & 0.312 (0\%) & 0.265 (0\%) \\\hline
-Random Undersampling (RU) & 0.321 (2.9\%) & 0.265 (0\%) \\\hline
-SMOTE-Tomek (ST) & 0.289 (-7.4\%) & 0.271 (2.3\%) \\\hline
-Random Oversampling (RO) & 0.369 (18.3\%) & 0.298 (12.5\%) \\\hline
-AdaSyn (AD) & 0.316 (1.3\%) & 0.251 (-5.3\%) \\\hline
-LLM-Augmented (LA) & {\bf 0.402 (28.8\%)} & {\bf 0.347 (30.9\%)} \\\hline
+Resampling Method & Recall (\% over none) & Precision (\% over none) & $F_1$ (\% over none) \\\hline
+ None (NO) & 0.265 (0\%) & {\bf 0.325 (0\%)} & 0.286 (0\%) \\\hline
+ RandomUnder (RU) & 0.265 (0\%) & 0.249 (-23.4\%) & 0.256 (-10.4\%) \\\hline
+ SMOTE-Tomek (ST) & 0.271 (2.3\%) & 0.264 (-18.8\%) & 0.267 (-6.6\%) \\\hline
+ Random Over (RO) & 0.298 (12.4\%) & 0.271 (-16.6\%) & 0.283 (-1.0\%) \\\hline
+ AdaSyn (AD) & 0.251 (-5.3\%) & 0.286 (-12.0\%) & 0.264 (-7.7\%) \\\hline
+ LLM-Augmented (LA) & {\bf 0.347 (30.9\%)} & 0.270 (-16.9\%) & {\bf 0.303 (5.9\%)} \\\hline
 \end{tabular}
 
-\includegraphics[scale=0.6]{5type-recall.pdf}\\
-{\bf Figure 4.} 5-class recall Nemenyi test results\\
-\vspace{\baselineskip}
-{\bf Table 4.} 5-class task resampling method precision\\
+\includegraphics[scale=0.6]{5type-f1.pdf}\\
+{\bf Figure 3.} 5-class $F_1$ Nemenyi test results\\
+\end{nscenter}
+The results on the 5-type task begin to show the benefits of LLM augmentation.
+The pattern of no resampling outperforming resampling on precision while the inverse happens on recall remains, but now LLM augmentation improves over no resampling by a far wider margin than any other resampling method, making its gain in mean recall account for its loss in precision, giving it a higher $F_1$ score than the other methods.
 
-\begin{tabular}{|l|l|l|}
+The results of the 4-class classification task are shown below. The data used to create the $F_1$ column of Table 3 received a Friedman $\chi^2$ test \emph{p}-value of $7.1 \times 10^{-5}$.
+\begin{nscenter}
+{\bf Table 3.} 4-class task resampling method performance\\
+
+\begin{tabular}{|l|l|l|l|}
 \hline
-Resampling Method & Best score (\% above none) & Mean score (\% above none) \\\hline
-None (NO) & {\bf 0.524 (0\%)} & {\bf 0.325 (0\%)} \\\hline
-Random Undersampling (RU) & 0.274 (-47.7\%) & 0.249 (-23.4\%) \\\hline
-SMOTE-Tomek (ST) & 0.295 (-43.7\%) & 0.264 (-18.8\%) \\\hline
-Random Oversampling (RO) & 0.319 (-39.1\%) & 0.271 (-16.6\%) \\\hline
-AdaSyn (AD) & 0.436 (-16.8\%) & 0.286 (-12.0\%) \\\hline
-LLM-Augmented (LA) & 0.302 (-42.4\%) & 0.270 (-16.9\%) \\\hline
+Resampling Method & Recall (\% over none) & Precision (\% over none) & $F_1$ (\% over none) \\\hline
+ None (NO) & 0.328 (0\%) & 0.372 (0\%) & 0.346 (0\%) \\\hline
+ RandomUnder (RU) & 0.339 (3.3\%) & 0.323 (-13.2\%) & 0.330 (-4.6\%) \\\hline
+ SMOTE-Tomek (ST) & 0.366 (11.6\%) & 0.347 (-6.7\%) & 0.356 (2.9\%) \\\hline
+ Random Over (RO) & 0.358 (9.1\%) & 0.351 (-5.6\%) & 0.354 (2.3\%) \\\hline
+ AdaSyn (AD) & 0.322 (-1.8\%) & 0.344 (-7.5\%) & 0.331 (-4.3\%) \\\hline
+ LLM-Augmented (LA) & {\bf 0.433 (32.0\%)} & {\bf 0.409 (9.9\%)} & {\bf 0.419 (21.1\%)} \\\hline
 \end{tabular}
 
-\includegraphics[scale=0.6]{5type-precision.pdf}\\
-{\bf Figure 5.} 5-class precision Nemenyi test results
-\end{nscenter}
-The results on the 5-type task begin to show the benefits of LLM augmentation.
-The pattern of no resampling outperforming resampling on precision while the inverse happens on recall remains, but now LLM augmentation improves over no resampling by a far wider margin than any other resampling method, while retaining similar levels of drops in precision to other methods.
-Now, its gain in mean recall is larger than its loss in mean precision, which is not the case for any other resampling method in the 5-class task, and is only true for random undersampling in the binary task.
-
-The results of the 4-class classification task are shown below.
-The data used to create Table 5 received a Friedman $\chi^2$ test \emph{p}-value of $5.3 \times 10^{-4}$, while Table 6 received a \emph{p}-value of $4.2 \times 10^{-5}$.
-\begin{nscenter}
-{\bf Table 5.} 4-class task resampling method recall\\
-
-\begin{tabular}{|l|l|l|}
-\hline
-Resampling Method & Best score (\% above none) & Mean score (\% above none) \\\hline
-None (NO) & 0.337 (0\%) & 0.328 (0\%) \\\hline
-Random Undersampling (RU) & 0.414 (22.8\%) & 0.339 (3.4\%) \\\hline
-SMOTE-Tomek (ST) & 0.420 (24.6\%) & 0.366 (11.6\%) \\\hline
-Random Oversampling (RO) & 0.400 (18.7\%) & 0.358 (9.1\%) \\\hline
-AdaSyn (AD) & 0.471 (39.8\%) & 0.322 (-1.8\%) \\\hline
-LLM-Augmented (LA) & {\bf 0.509 (51.0\%)} & {\bf 0.433 (32.0\%)} \\\hline
-\end{tabular}
-
-\includegraphics[scale=0.6]{4type-recall.pdf}\\
-{\bf Figure 6.} 4-class recall Nemenyi test results\\
-\vspace{\baselineskip}
-{\bf Table 6.} 4-class task resampling method precision\\
-
-\begin{tabular}{|l|l|l|}
-\hline
-Resampling Method & Best score (\% above none) & Mean score (\% above none) \\\hline
-None (NO) & 0.452 (0\%) & 0.372 (0\%) \\\hline
-Random Undersampling (RU) & 0.361 (-17.9\%) & 0.323 (-13.2\%) \\\hline
-SMOTE-Tomek (ST) & 0.381 (-15.7\%) & 0.347 (-6.7\%) \\\hline
-Random Oversampling (RO) & 0.374 (-17.3\%) & 0.351 (-5.6\%) \\\hline
-AdaSyn (AD) & 0.378 (-16.4\%) & 0.344 (-7.5\%) \\\hline
-LLM-Augmented (LA) & {\bf 0.504 (11.5\%)} & {\bf 0.409 (9.9\%)} \\\hline
-\end{tabular}
-
-\includegraphics[scale=0.6]{4type-precision.pdf}\\
-{\bf Figure 5.} 4-class precision Nemenyi test results
+\includegraphics[scale=0.6]{4type-f1.pdf}\\
+{\bf Figure 4.} 4-class $F_1$ Nemenyi test results\\
 \end{nscenter}
@@
 Looking further at the data, however, while some classes were more affected than others, overall, the models trained on LLM-augmented data show improvements in multiple classes.
-On the 4-type classification task, Tables 7 \& 8 show the mean recall and precision of each method on each particular class, respectively.
+On the 4-type classification task, Table 4 shows the mean $F_1$ of each method on each particular class, respectively.
 \begin{nscenter}
-{\bf Table 7.} 4-class task class-wise resampling method recall\\
+{\bf Table 4.} 4-class task class-wise resampling method $F_1$ score\\
 
@@
 Resampling Method & Political & Economic & Religious & Racial \\\hline
-None (NO) & {\bf 0.618} & 0.106 & 0.203 & 0.283 \\\hline
-Random Undersampling (RU) & 0.379 & {\bf 0.240} & 0.434 & 0.200  \\\hline
-SMOTE-Tomek (ST) & 0.506 & 0.135 & 0.465 & 0.245 \\\hline
-Random Oversampling (RO) & 0.487 & 0.163 & 0.442 & 0.230 \\\hline
-AdaSyn (AD) & 0.574 & 0.135 & 0.223 & 0.258 \\\hline
-LLM-Augmented (LA) & 0.492 & 0.231 & {\bf 0.471} & {\bf 0.405} \\\hline
-\end{tabular}
-
-{\bf Table 8.} 4-class task class-wise resampling method precision\\
-
-\begin{tabular}{|l|l|l|l|l|}
-\hline
-Resampling Method & Political & Economic & Religious & Racial \\\hline
-None (NO) & 0.598 & 0.029 & 0.254 & 0.390 \\\hline
-Random Undersampling (RU) & 0.601 & 0.024 & 0.205 & 0.362 \\\hline
-SMOTE-Tomek (ST) & 0.653 & 0.018 & 0.234 & 0.378 \\\hline
-Random Oversampling (RO) & {\bf 0.661} & 0.015 & 0.227 & 0.392 \\\hline
-AdaSyn (AD) & 0.602 & 0.017 & 0.243 & 0.390 \\\hline
-LLM-Augmented (LA) & 0.577 &{\bf  0.111} & {\bf 0.367} & {\bf 0.431} \\\hline
+None (NO) & {\bf 0.597} & 0.059 & 0.233 & 0.315 \\\hline
+Random Undersampling (RU) & 0.463 & 0.050 & 0.276 & 0.253 \\\hline
+SMOTE-Tomek (ST) & 0.568 & 0.056 & 0.304 & 0.288 \\\hline
+Random Oversampling (RO) & 0.558 & 0.038 & 0.294 & 0.280 \\\hline
+AdaSyn (AD) & 0.581 & 0.046 & 0.228 & 0.301 \\\hline
+LLM-Augmented (LA) & 0.526 & {\bf 0.161} & {\bf 0.403} & {\bf 0.415} \\\hline
 \end{tabular}
 \end{nscenter}
-In precision and recall, the model is outperformed in political antisemitism (the largest class), and in recall it is slightly outperformed by random undersampling in economic antisemitism, but it outperforms the other models in religious and racial antisemitism on recall and in all classes other than political in precision.
+In $F_1$ score, the model is outperformed in political antisemitism (the largest class), but it outperforms the other models in all other classes.
 

### 2023-05-17 paper/paper.tex
add appendix

@@
 \end{thebibliography}
+
+\appendix
+\section{Algorithm-specific data}
+\subsection{Extreme Gradient Boosting}
+\begin{nscenter}
+{\bf Table 5.} XGBoost-Bag of Words $F_1$ scores\\
+
+\begin{tabular}{|l|l|l|l|}
+\hline
+Resampling Method & Binary & 5-class & 4-class \\\hline
+None (NO) & 0.656 & 0.294  & 0.338 \\\hline
+Random Undersampling (RU) & {\bf 0.666} & 0.238 & 0.333 \\\hline
+SMOTE-Tomek (ST) & 0.660 & 0.286 & 0.372 \\\hline
+Random Oversampling (RO) & 0.663 & {\bf 0.319} & 0.386 \\\hline
+AdaSyn (AD) & 0.629 & 0.251 & 0.342 \\\hline
+LLM-Augmented (LA) & 0.660 & 0.300 & {\bf 0.432} \\\hline
+\end{tabular}
+\end{nscenter}
+
+\begin{nscenter}
+{\bf Table 6.} XGBoost-TF-IDF $F_1$ scores\\
+
+\begin{tabular}{|l|l|l|l|}
+\hline
+Resampling Method & Binary & 5-class & 4-class \\\hline
+None (NO) & 0.630 & 0.293 & 0.320 \\\hline
+Random Undersampling (RU) & 0.629 & 0.296 & 0.306 \\\hline
+SMOTE-Tomek (ST) & 0.642 & 0.273 & 0.327 \\\hline
+Random Oversampling (RO) & 0.627 & 0.274 & 0.318 \\\hline
+AdaSyn (AD) & {\bf 0.644} & 0.254 & 0.307 \\\hline
+LLM-Augmented (LA) & 0.636 & {\bf 0.313} & {\bf 0.520} \\\hline
+\end{tabular}
+\end{nscenter}
+
+\begin{nscenter}
+{\bf Table 7.} XGBoost-Frequency $F_1$ scores\\
+
+\begin{tabular}{|l|l|l|l|}
+\hline
+Resampling Method & Binary & 5-class & 4-class \\\hline
+None (NO) & 0.635 & 0.272 & 0.342 \\\hline
+Random Undersampling (RU) & 0.630 & 0.269 & 0.309 \\\hline
+SMOTE-Tomek (ST) & {\bf 0.644} & 0.287 & 0.357 \\\hline
+Random Oversampling (RO) & 0.636 & 0.272 & 0.363 \\\hline
+AdaSyn (AD) & 0.630 & 0.255 & 0.309 \\\hline
+LLM-Augmented (LA) & 0.642 & {\bf 0.329} & {\bf 0.460} \\\hline
+\end{tabular}
+\end{nscenter}
+
+\subsection{Support Vector Machine}
+
+\begin{nscenter}
+{\bf Table 8.} SVM-Bag of Words $F_1$ scores\\
+
+\begin{tabular}{|l|l|l|l|}
+\hline
+Resampling Method & Binary & 5-class & 4-class \\\hline
+None (NO) & {\bf 0.699} & {\bf 0.363} & 0.362 \\\hline
+Random Undersampling (RU) & 0.674 & 0.249 & 0.356 \\\hline
+SMOTE-Tomek (ST) & 0.665 & 0.281 & 0.400 \\\hline
+Random Oversampling (RO) & 0.682 & 0.298 & 0.380 \\\hline
+AdaSyn (AD) & 0.646 & 0.302 & 0.338 \\\hline
+LLM-Augmented (LA) & 0.673 & 0.332 & {\bf 0.413} \\\hline
+\end{tabular}
+\end{nscenter}
+
+\begin{nscenter}
+{\bf Table 9.} SVM-TF-IDF $F_1$ scores\\
+
+\begin{tabular}{|l|l|l|l|}
+\hline
+Resampling Method & Binary & 5-class & 4-class \\\hline
+None (NO) & {\bf 0.679} & {\bf 0.340} & 0.332 \\\hline
+Random Undersampling (RU) & 0.649 & 0.254 & 0.379 \\\hline
+SMOTE-Tomek (ST) & 0.667 & 0.287 & 0.377 \\\hline
+Random Oversampling (RO) & 0.659 & 0.329 & 0.378 \\\hline
+AdaSyn (AD) & 0.648 & 0.310 & 0.314 \\\hline
+LLM-Augmented (LA) & 0.658 & 0.295 & {\bf 0.403} \\\hline
+\end{tabular}
+\end{nscenter}
+
+\begin{nscenter}
+{\bf Table 10.} SVM-Frequency $F_1$ scores\\
+
+\begin{tabular}{|l|l|l|l|}
+\hline
+Resampling Method & Binary & 5-class & 4-class \\\hline
+None (NO) & 0.663 & 0.307 & 0.395 \\\hline
+Random Undersampling (RU) & {\bf 0.668} & 0.236 & 0.327 \\\hline
+SMOTE-Tomek (ST) & 0.654 & 0.298 & 0.340 \\\hline
+Random Oversampling (RO) & 0.656 & {\bf 0.316} & 0.370 \\\hline
+AdaSyn (AD) & 0.653 & 0.289 & 0.298 \\\hline
+LLM-Augmented (LA) & 0.634 & 0.312 & {\bf 0.402} \\\hline
+\end{tabular}
+\end{nscenter}
+
+\subsection{Decision Tree}
+
+\begin{nscenter}
+{\bf Table 11.} Decision Tree-Bag of Words $F_1$ scores\\
+
+\begin{tabular}{|l|l|l|l|}
+\hline
+Resampling Method & Binary & 5-class & 4-class \\\hline
+None (NO) & 0.619 & 0.283 & 0.392 \\\hline
+Random Undersampling (RU) & {\bf 0.650} & 0.257 & 0.343 \\\hline
+SMOTE-Tomek (ST) & 0.638 & 0.267 & 0.384 \\\hline
+Random Oversampling (RO) & 0.635 & 0.292 & 0.384 \\\hline
+AdaSyn (AD) & 0.602 & 0.234 & 0.323 \\\hline
+LLM-Augmented (LA) & 0.577 & {\bf 0.313} & {\bf 0.453} \\\hline
+\end{tabular}
+\end{nscenter}
+
+\begin{nscenter}
+{\bf Table 12.} Decision Tree-TF-IDF $F_1$ scores\\
+
+\begin{tabular}{|l|l|l|l|}
+\hline
+Resampling Method & Binary & 5-class & 4-class \\\hline
+None (NO) & 0.603 & 0.250 & 0.316 \\\hline
+Random Undersampling (RU) & {\bf 0.613} & 0.274 & 0.353 \\\hline
+SMOTE-Tomek (ST) & 0.590 & 0.246 & 0.334 \\\hline
+Random Oversampling (RO) & 0.587 & 0.254 & 0.305 \\\hline
+AdaSyn (AD) & 0.598 & 0.240 & 0.336 \\\hline
+LLM-Augmented (LA) & 0.599 & {\bf 0.312} & {\bf 0.436} \\\hline
+\end{tabular}
+\end{nscenter}
+
+\begin{nscenter}
+{\bf Table 13.} Decision Tree-Frequency $F_1$ scores\\
+
+\begin{tabular}{|l|l|l|l|}
+\hline
+Resampling Method & Binary & 5-class & 4-class \\\hline
+None (NO) & 0.601 & 0.251 & 0.318 \\\hline
+Random Undersampling (RU) & {\bf 0.622} & 0.283 & 0.301 \\\hline
+SMOTE-Tomek (ST) & 0.589 & 0.243 & 0.328 \\\hline
+Random Oversampling (RO) & 0.596 & 0.256 & 0.337 \\\hline
+AdaSyn (AD) & 0.602 & 0.244 & 0.302 \\\hline
+LLM-Augmented (LA) & 0.581 & {\bf 0.325} & {\bf 0.478} \\\hline
+\end{tabular}
+\end{nscenter}
+
+\subsection{Naïve Bayes}
+
+\begin{nscenter}
+{\bf Table 14.} NB-Bag of Words $F_1$ scores\\
+
+\begin{tabular}{|l|l|l|l|}
+\hline
+Resampling Method & Binary & 5-class & 4-class \\\hline
+None (NO) & {\bf 0.625} & 0.287 & 0.352 \\\hline
+Random Undersampling (RU) & 0.594 & 0.212 & 0.312 \\\hline
+SMOTE-Tomek (ST) & 0.539 & 0.253 & 0.375 \\\hline
+Random Oversampling (RO) & 0.608 & 0.282 & 0.353 \\\hline
+AdaSyn (AD) & 0.538 & 0.287 & {\bf 0.418} \\\hline
+LLM-Augmented (LA) & 0.565 & {\bf 0.303} & 0.335 \\\hline
+\end{tabular}
+\end{nscenter}
+
+\begin{nscenter}
+{\bf Table 15.} NB-TF-IDF $F_1$ scores\\
+
+\begin{tabular}{|l|l|l|l|}
+\hline
+Resampling Method & Binary & 5-class & 4-class \\\hline
+None (NO) & 0.559 & 0.253 & 0.347 \\\hline
+Random Undersampling (RU) & 0.575 & 0.255 & 0.343 \\\hline
+SMOTE-Tomek (ST) & 0.535 & 0.254 & 0.341 \\\hline
+Random Oversampling (RO) & 0.540 & 0.255 & 0.343 \\\hline
+AdaSyn (AD) & 0.528 & 0.255 & 0.346 \\\hline
+LLM-Augmented (LA) & {\bf 0.578} & {\bf 0.267} & {\bf 0.356} \\\hline
+\end{tabular}
+\end{nscenter}
+
+\begin{nscenter}
+{\bf Table 16.} NB-Frequency $F_1$ scores\\
+
+\begin{tabular}{|l|l|l|l|}
+\hline
+Resampling Method & Binary & 5-class & 4-class \\\hline
+None (NO) & 0.558 & 0.243 & 0.336 \\\hline
+Random Undersampling (RU) & {\bf 0.584} & 0.250 & 0.297 \\\hline
+SMOTE-Tomek (ST) & 0.535 & 0.250 & 0.336 \\\hline
+Random Oversampling (RO) & 0.548 & {\bf 0.255} & 0.334 \\\hline
+AdaSyn (AD) & 0.528 & 0.249 & 0.336 \\\hline
+LLM-Augmented (LA) & 0.576 & 0.245 & {\bf 0.342} \\\hline
+\end{tabular}
+\end{nscenter}
+
 \end{document}

### 2023-05-17 paper/paper.tex
final revisions

@@
 In these arrays, ``RU'' stands for random undersampling, ``ST'' stands for SMOTE with Tomek links, ``RO'' stands for random oversampling, ``AD'' stands for AdaSyn, ``NO'' stands for no resampling method, and ``LA'' stands for LLM-augmented.}
+\clearpage
 \begin{nscenter}
@@
 \section{Algorithm-specific data}
+In order to get a more precise understanding of what happens for each combination of text representation and classification algorithm, we report the $F_1$ results obtained in each of the 12 cases considered.
+Space limitation prevented us from reporting the precision and recall results, but they are available upon request.
+The results show that for all but one combination models resampled with LLM augmentation fared better than all the other models in the 4-class case.
+The results are mixed in the 5-class case with XGBoost, decision trees, and Naïve Bayes sometimes benefitting from LLM-Augmented resampling.
+Overall, the best method in the 4-class classification is XGBoost with TF-IDF and LLM resampling, while in the other two tasks it is SVM with BoW and no resampling.
+
 \subsection{Extreme Gradient Boosting}
@@
 
+\clearpage
 \begin{nscenter}
